### 2023-09-08 Tulai.qmd
test

---
title: "Tulai"
author: ""
date: now
format: 
  html:
    self-contained: true
execute: 
  warning: false
  fig-dpi: 600
title-block-published: "last updated"
editor: visual
---

```{r}
# library
library(tidyverse)
library(ggplot2)
library(dplyr)
library(magrittr)
library(stringr)

```

Cleaning Data:

```{r}
# Read the data
tl <- read_csv("Tulai_Lithic_Assamblage.csv")

# Create a new data frame with renamed columns
tl1 <- tl
names(tl1)[1:2] <- c("lithic_id", 
                     "TA")

# modify TA column
tl1 <- tl1 %>%
  # Remove everything within brackets
  mutate(TA = str_remove_all(TA, 
                             "\\(.*\\)")) %>%

  # Add slash between TA and numbers
  mutate(TA = str_replace_all(TA, 
                              "TA([0-9]+)", 
                              "TA/\\1")) %>%

  # Add slash between numbers
  mutate(TA = str_replace_all(TA, 
                              "([0-9]+)/([0-9]+)", "\\1/\\2")) %>% 
 # fill NA values in TA column
  tidyr::fill(TA, .direction = "down") %>%
  # Prepend 'TA/' to numbers
  mutate(TA = if_else(str_detect(TA, 
                                 "^[0-9]+$"), 
                      paste0("TA/",
                             TA), 
                      TA)) %>% 
  # Remove all non-alphanumeric characters
  mutate(TA = str_replace_all(TA, 
                              "[^[:alnum:] /]", 
                              "")) %>% 
  # Make all text lowercase
  mutate(TA = tolower(TA)) %>% 
  # Remove white space
  mutate(TA = str_replace_all(TA, " ", ""))  %>%

  # Separate 'TA' column into 'area' and 'depth' columns
  separate(TA, into = c("area", 
                        "depth"), 
           sep = "/",
           remove = FALSE,
           convert = TRUE) %>%
  # Convert 'depth' to numeric
  mutate(depth = as.numeric(depth)) %>%

  # Create depth ranges
  mutate(depth = case_when(
             depth == 3 ~ "0-30 cm",
             depth == 34 ~ "30-40 cm",
             depth == 45 ~ "40-50 cm",
             depth == 56 ~ "50-60 cm",
             depth == 67 ~ "60-70 cm",
             depth == 72 ~ "70-120 cm",
             depth == 125 ~ "120-150 cm",
             depth == 158 ~ "150-180 cm",
             depth == 189 ~ "180-190 cm",
             depth == 190 ~ "190-200 cm",
             depth == 201 ~ "200-210 cm",
             depth == 212 ~ "210-220 cm",
             depth == 223 ~ "220-230 cm",
             depth == 234 ~ "230-240 cm",
             depth == 235 ~ "230-250 cm",
             depth == 256 ~ "250-260 cm",
             depth == 267 ~ "260-270 cm",
             depth == 72 ~ "70-120 cm",
             depth == 1 ~ "0-10 cm",
             depth == 2 ~ "0-20 cm", 
             depth == 13 ~ "10-30 cm",
             depth == 19 ~ "10-90 cm",
             depth == 23 ~ "20-30 cm",
             depth == 34 ~ "30-40 cm",
             depth == 45 ~ "40-50 cm",
             depth == 51 ~ "50-100 cm",
             depth == 56 ~ "50-60 cm",
             depth == 57 ~ "50-70 cm",
             depth == 89 ~ "80-90 cm",
             depth == 91 ~ "90-100 cm",
             depth == 101 ~ "100-110 cm",
             depth == 112 ~ "110-120 cm",
             depth == 123 ~ "120-130 cm",
             depth == 124 ~ "120-140 cm",
             depth == 23 ~ "20-30 cm",
             depth == 46 ~ "40-60 cm",
             depth == 78 ~ "70-80 cm",
             depth == 235 ~ "230-250 cm",
             depth == 12 ~ "10-20 cm",
             depth == 90 ~ "90-100 cm",

             TRUE ~ as.character(depth)
  )) %>%

  separate(depth, 
           into = c("upper",
                    "lower"),
           sep = "-",
           remove = FALSE,
           convert = TRUE) %>% 
  mutate(lower = parse_number(lower)) %>% 
  rowwise() %>% 
  mutate(midpoint = mean(c(upper, lower))) %>%

# create 'layer' column
  mutate(layer = case_when(
      area == "tp1" & depth == "0-10 cm" ~ "c",
      area == "tp1" & depth == "0-20 cm" ~ "c",
      area == "tp1" & depth == "10-30 cm" ~ "c",
      area == "tp1" & depth == "10-90 cm" ~ "c",
    area == "tp1" & depth == "20-30 cm" ~ "c",
    area == "tp1" & depth == "30-40 cm" ~ "c",
    area == "tp1" & depth == "40-50 cm" ~ "c",
    area == "tp1" & depth == "50-100 cm" ~ "c",
    area == "tp1" & depth == "50-60 cm" ~ "c",
    area == "tp1" & depth == "50-70 cm" ~ "c",
    area == "tp1" & depth == "80-90 cm" ~ "c",
    area == "tp1" & depth == "90-100 cm" ~ "c",
    area == "tp1" & depth == "100-110 cm" ~ "c",
    area == "tp1" & depth == "110-120 cm" ~ "c",
    area == "tp1" & depth == "120-130 cm" ~ "c",
    area == "tp1" & depth == "120-140 cm" ~ "c",
    area == "d1" & depth == "0-20 cm" ~ "c",
    area == "d1" & depth == "20-30 cm" ~ "c",
    area == "d1" & depth == "40-60 cm" ~ "a",
    area == "d1" & depth == "50-60 cm" ~ "a",
    area == "d1" & depth == "60-70 cm" ~ "a",
    area == "d1" & depth == "70-80 cm" ~ "a",
    area == "ta" & depth == "0-30 cm" ~ "e",
    area == "ta" & depth == "30-40 cm" ~ "e",
    area == "ta" & depth == "40-50 cm" ~ "e",
    area == "ta" & depth == "50-60 cm" ~ "e",
    area == "ta" & depth == "60-70 cm" ~ "e",
    area == "ta" & depth == "70-120 cm" ~ "d",
    area == "ta" & depth == "120-150 cm" ~ "d",
    area == "ta" & depth == "150-180 cm" ~ "d",
    area == "ta" & depth == "180-190 cm" ~ "d",
    area == "ta" & depth == "190-200 cm" ~ "d",
    area == "ta" & depth == "200-210 cm" ~ "d",
    area == "ta" & depth == "210-220 cm" ~ "d",
    area == "ta" & depth == "220-230 cm" ~ "d",
    area == "ta" & depth == "230-240 cm" ~ "d",
    area == "ta" & depth == "230-250 cm" ~ "d",
    area == "ta" & depth == "250-260 cm" ~ "d",
    area == "ta" & depth == "260-270 cm" ~ "d",

      TRUE ~ NA_character_
  ))
# Remove specific columns
tl1 %>% 
  select(-c("AREA", 
            "DEPTH (cm)", 
            "Raw Material", 
            "Colour/Grain/Opacity", 
            "Pattern"))

```

```{r}
#exploring Data

tl1 %>% 
  group_by(midpoint) %>% 
  tally() %>% 
  drop_na(midpoint) %>% 
ggplot() + 
  aes(midpoint, n) +
  geom_col()
```

```{r}
#exploring Data

tl1 %>% 
  group_by(layer) %>% 
  tally() %>% 
  drop_na(layer) %>% 
ggplot() + 
  aes(layer, n) +
  geom_col()
```

```{r}
#exploring Data

tl2 <- tl1 %>%
  mutate(
    Blank = case_when(
      Breakage %in% c("0", "1", NA_character_) ~ Blank,

      TRUE ~ Breakage  
    )
  )

```

```{r}
# Cleaning Blank column
library(dplyr)
library(stringr)
library(tidyr)

tl2 <- tl2 %>%
  mutate(
    Blank = str_to_lower(Blank),
    Blank = str_trim(Blank),
    Blank = str_replace_all(Blank, "[./]", "-"),
    Blank = str_replace_all(Blank, "\\s*-\\s*", "-")
  ) %>%
  mutate(Blank = case_when(
             Blank == "peo" ~ "bladelet-pro",
             Blank == "pro" ~ "bladelet-pro",
             Blank == "bladelet-dis" ~ "bladelet-dis",
             Blank == "microblade" ~ "bladelet-complete",
             Blank == "microblade-pro" ~ "bladelet-pro",
             Blank == "microblade-med" ~ "bladelet-med",
             Blank == "micrpblade-dis" ~ "bladelet-dis",
             Blank == "dis" ~ "bladelet-dis",
             Blank == "bladelrt-med" ~ "bladelet-med",
             Blank == "bladelet-mes" ~ "bladelet-med",
             Blank == "bladelet-,ed" ~ "bladelet-med",
             Blank == "bladelet-bladelet-med" ~ "bladelet-med",
             Blank == "medial" ~ "bladelet-med",
             Blank == "bladelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bladelert-pro" ~ "bladelet-pro",
             Blank == "microblade-medial" ~ "bladelet-med",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "microblade-?" ~ "bladelet-complete",
             Blank == "med" ~ "bladelet-med",
             Blank == "bladeler-pro" ~ "bladelet-pro",
             Blank == "microblade-dis" ~ "bladelet-dis",
             Blank == "microblde-dis" ~ "bladelet-dis",
             Blank == "bladlet-pro" ~ "bladelet-pro",
             Blank == "indistinct" ~ "bladelet-med",
             Blank == "flake" ~ "flake-complete",
             Blank == "dladelet-pro" ~ "bladelet-pro",
             Blank == "dladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-pro" ~ "bladelet-pro",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "microblade-?" ~ "bladelet-med",
             Blank == "bladelt-pro" ~ "bladelet-pro",
             Blank == "bladeket-med" ~ "bladelet-med",
             Blank == "nicroblade-pro" ~ "bladelet-pro",
             Blank == "blaelet" ~ "bladelet-complete",
             Blank == "bladele-med" ~ "bladelet-med",
             Blank == "bladelet-nearly complete" ~ "bladelet-complete",
             Blank == "blade-nearly complete" ~ "blade-complete",
             Blank == "bladelet-?" ~ "bladelet-complete",
             Blank == "flake?" ~ "flake-complete",
             Blank == "micrpblade-dis" ~ "bladelet-dis",
             Blank == "blaelet-med" ~ "bladelet-med",
             Blank == "blaedlet-pro" ~ "bladelet-pro",  
             Blank == "bladelt-pro" ~ "bladelet-pro",   
             Blank == "bladelet-nearly complete" ~ "bladelet-complete", 
             Blank == "microblade-med" ~ "bladelet-med", 
             Blank == "bldelet-bladelet-pro" ~ "bladelet-pro",
             Blank == "bldelet-flake" ~ "blade-complete",
             Blank == "mwdial" ~ "bladelet-med",
             Blank == "blade" ~ "blade-complete",
             Blank == "bladelt" ~ "bladelet-complete",
             Blank == "bladlet" ~ "bladelet-complete",
             Blank == "bladlet-med" ~ "bladelet-med",
             Blank == "bladelt-med" ~ "bladelet-med",
             Blank == "flaje" ~ "flake",
             Blank == "flke" ~ "flake",
             Blank == "bladelet" ~ "bladelet-complete",
             Blank == "NA" ~ "bladelet-",
             Blank == "thick flake" ~ "flake",
             Blank == "flke-mid" ~ "flake-med",
             Blank == "flake-blade" ~ "flake",
             Blank == "angular flake?" ~ "flake",
             Blank == "flake-prox" ~ "flake-pro",
             Blank == "fklae-dis" ~ "flake-dis",
             Blank == "bladelete" ~ "bladelet",
             Blank == "NA" ~ NA_character_,
    TRUE ~ Blank
  )) %>%
  mutate(Blank = na_if(Blank, ""))

unique(tl2$Blank)

tl3 <- tl2 %>%
  separate(Blank, 
           into = c("Blank2", "Blank_part"),
           sep = "-",
           remove = FALSE,
           convert = TRUE)

na_rows <- which(is.na(tl2$Blank))
print(na_rows)

zero_rows <- which(tl2$Blank == "0")
print(zero_rows)

one_rows <- which(tl2$Blank == "1")
print(one_rows)

print(unique
      (tl3$Blank2))

```

```{r}

tl3 <- tl3 %>%
  mutate(
    Blank2 = case_when(
      Blank2 == "flake" ~ "flake",  # Keep "flake" as it is
      (`Core Typology` %in% c("0", "NA", "na", "-", ".") | 
       is.na(`Core Typology`) | 
       `Core Typology` == "") & `Width (mm)` < 12 ~ "bladelet",
      (`Core Typology` %in% c("0", "NA", "na", "-", ".") | 
       is.na(`Core Typology`) | 
       `Core Typology` == "") & `Width (mm)` >= 12 ~ "blade",
      TRUE ~ Blank2
    )
  )

```

```{r}
tl3 <- tl3 %>%
  mutate(
    Blank2 = case_when(
      Blank2 %in% c("blade", "bladelet", "flake") ~ Blank2,
      TRUE ~ NA_character_
    )
  )

```

```{r}
# Remove rows based on row names
tl3 <- tl3[!(rownames(tl3) %in% c("1030", "644", "543")), ]
```

```{r}
tl4 <- tl3 %>%
  mutate(
    Utilization = if_else(Retouch == "1", NA_character_, as.character(Utilization))
  )

```

```{r}
library(dplyr)
library(stringr)

tl4 <- tl4 %>%
  mutate_all(~ ifelse(is.na(.x) | .x == "", NA, .x)) %>% 
  mutate_all(~ str_replace(., "(?<=\\D)\\*(?=\\D)|\\s+|[*?!]+", "")) %>%
  mutate(
    SHINE = str_trim(SHINE) %>% 
            str_to_title(),
    shine2 = SHINE
  ) %>%
  mutate(
    shine2 = case_when(
      SHINE %in% c("Subparallel", "Parallel", "Sub-Parallel", "Parallel-Subprarallel",
                   "Scaled", "Sub-Paralel", "Sub-Parallell", "Sub-Paallel",
                   "Sub-Paralell", "Sub-Prallel", "L:semiparallel/R: Scaled",
                   "Semiparallel", "Semi-Parallel", "Scaled/Sub-Parallel",
                   "Seb-Parallel", "Semi-Abrupt") ~ NA_character_,
      TRUE ~ SHINE
    ),
    Utilization = if_else(Utilization == "0", NA_character_, Utilization),
    Retouch = if_else(Retouch == "0", NA_character_, Retouch),
    SHINE = if_else(SHINE == "0", NA_character_, SHINE),
    shine2 = if_else(shine2 == "0", NA_character_, shine2),
    Blank2 = if_else(Blank2 == "0", NA_character_, Blank2)
  ) %>% 
  mutate(
    SHINE = na_if(SHINE, "")
  )

print(unique(tl4$SHINE))

```

```{r}
library(dplyr)
library(stringr)
# Convert 'NA' and 'na' strings to actual NA
tl4$`Core Typology` <- na_if(tl4$`Core Typology`, "NA")
tl4$`Core Typology` <- na_if(tl4$`Core Typology`, "na")

tl4 <- tl4 %>%
  mutate(
    `Core Typology` = str_to_lower(`Core Typology`),
    `Core Typology` = str_trim(`Core Typology`),
    `Core Typology` = str_replace_all(`Core Typology`, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%

  mutate(`Core Typology` = case_when(
    `Core Typology` == "flatcore" ~ "flat-pressure",
             `Core Typology` == "rejuvention" ~ "rejuvenation piece-NA",
             `Core Typology` == "pressureprymidal core" ~ "pyramid-pressure",
             `Core Typology` == "pressureflat core" ~ "flat-pressure",
             `Core Typology` == "pressurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prussurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prismaticcore" ~ "prismatic-percussion",
             `Core Typology` == "pyramidalpressure core" ~ "pyramid-pressure",
             `Core Typology` == "flatcore with one debitage surface/pressure" ~ "flat-pressure",
             `Core Typology` == "corefragment" ~ "core fragment-NA",

             `Core Typology` == "cilandrical/pressure" ~ "cylinder-pressure",
             `Core Typology` == "bullet/pressure" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/pressure" ~ "pyramid-pressure",
             `Core Typology` == "percussion/pyramidal" ~ "pyramid-pressure",
             `Core Typology` == "flat/pressure" ~ "flat-pressure",
             `Core Typology` == "pyramidal/ percussion??" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ peressure" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ pressure" ~ "pyramid-pressure",
             `Core Typology` == "pressure" ~ "pyramid-pressure",
             `Core Typology` == "multidirectional/percussion" ~ "shapeless-percussion",
             `Core Typology` == "prismatic/ percussion" ~ "pyramid-pressure",
             `Core Typology` == "flatcore/ unidirectional/pressure" ~ "flat-pressure",
             `Core Typology` == "heavilyused" ~ "NA",
             `Core Typology` == "burnt" ~ "NA",
             `Core Typology` == "patinated" ~ "NA",
             `Core Typology` == "pyramidal(bullet)/pressure" ~ "bullet-pressure",
             `Core Typology` == "cylandrycal/ bidirectional pressure core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure bullet core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "bladelet" ~ "bullet-pressure",
             `Core Typology` == "lip" ~ "NA",
             `Core Typology` == "bladlet" ~ "bullet-pressure",
             `Core Typology` == "bladelet(30.56)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(28.55.6)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(16.28.1)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(40 5.7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(23.2 2.3)" ~ "bullet-pressure",
             `Core Typology` == "bladelet-flakeblade(41 16)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(3512.6/7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(27.307.5)" ~ "bullet-pressure",
    `Core Typology` == "pressure/pyramidal" ~ "pyramid-pressure",
    `Core Typology` == "bladeletcore" ~ "bullet-pressure",
    `Core Typology` == "pyramidal? / pressure" ~ "pyramid-pressure",
    `Core Typology` == "prismatic" ~ "pyramid-pressure",

    `Core Typology` == "bullet" ~ "bullet-pressure",

    TRUE ~ `Core Typology`
  ))

# Separate 'Core Typology' into 'core-typology' and 'core-technology'
tl4 <- tl4 %>%
  separate(`Core Typology`, 
           into = c("core-typology", "core-technology"),
           sep = "-",
           remove = FALSE,
           convert = TRUE)

# Convert 'NA' and 'na' in new columns to actual NA
tl4$`core-typology` <- na_if(tl4$`core-typology`, "na")
tl4$`core-technology` <- na_if(tl4$`core-technology`, "na")

# Check the unique values again
unique(tl4$`core-typology`)
unique(tl4$`core-technology`)
```

```{r}

library(dplyr)
library(stringr)

tl4 <- tl4 %>%
  mutate(
    Typology = str_to_lower(Typology),
    Typology = str_trim(Typology),
    Typology = str_replace_all(Typology, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(Typology = case_when(

             `Typology` == "notch, truncated" ~ "tool/notch;truncated piece",

             `Typology` == "retouchpiece" ~ "tool/retouched piece",

             `Typology` == "endscraper" ~ "tool/end scraper",

             `Typology` == "notch" ~ "tool/notch",

             `Typology` == "sidescraper" ~ "tool/side scraper",

             `Typology` == "doubleside scraper" ~ "tool/double side scraper",

             `Typology` == "convergentscraper (?)" ~ "tool/convergent scraper",

             `Typology` == "doublenotch on retouched piece" ~ "tool/double notch",

             `Typology` == "retouchpieces" ~ "tool/retouched piece",

             `Typology` == "burin(?)/ notch/scraper" ~ "tool/burin-notch-scraper",

             `Typology` == "borer" ~ "tool/perforator-borer",

             `Typology` == "retouchedpiece" ~ "tool/retouched piece",

             `Typology` == "notch/side scraper" ~ "tool/notch-side scraper",

             `Typology` == "corefragment" ~ "core/core fragment",

             `Typology` == "backed" ~ "tool/backed knife",

             `Typology` == "doubleside scraper ?" ~ "tool/double side scraper",

             `Typology` == "inversedenticulate" ~ "tool/denticulate",

             `Typology` == "retouchpieces/ side scraper?" ~ "tool/side scraper",

             `Typology` == "sidescraper ?" ~ "tool/side scraper",

             `Typology` == "notch-denticulate" ~ "tool/notch-denticulate",

             `Typology` == "borer/drill" ~ "tool/perforator-drill",

             `Typology` == "retouchedpiece (side scraper?)" ~ "tool/side scraper",

             `Typology` == "retouchepiece (side scraper?)" ~ "tool/side scraper",

             `Typology` == "retouchedpiece (double side scraper)" ~ "tool/double side scraper",

             `Typology` == "retouchedpiece (double side scraper?)" ~ "tool/double side scraper",

             `Typology` == "retouchepiece" ~ "tool/retouched piece",

             `Typology` == "inversenotch" ~ "tool/notch",

             `Typology` == "retouchepiece (double side scaraper?)" ~ "tool/double side scraper",

             `Typology` == "retouchedpiece( double side scraper?)" ~ "tool/double side scraper",

             `Typology` == "notchon retouche piece" ~ "tool/notch",

             `Typology` == "retouchedpiece (souble side scraper)" ~ "tool/double side scraper",

             `Typology` == "point" ~ "tool/perforator-point",

             `Typology` == "chunk" ~ "NA/NA",

             `Typology` == "coretablet" ~ "core/core tablet",

             `Typology` == "core" ~ "core/shapeless",

             `Typology` == "denticulate" ~ "tool/denticulate",

             `Typology` == "doublenotch (haft)" ~ "tool/double notch",

             `Typology` == "retouchedpiece (fine retouch)" ~ "tool/retouched piece",

             `Typology` == "endscraper on retouched piece" ~ "tool/end scraper",

             `Typology` == "retouchedpiece (serrated scraper)" ~ "tool/serrated scraper",

             `Typology` == "truncation" ~ "tool/truncated piece",

             `Typology` == "rejuventionpiece" ~ "core/rejuvenation piece",

             `Typology` == "notch/denticulate" ~ "tool/notch-denticulate",

             `Typology` == "retoucheduse on breakage" ~ "tool/retouched piece",

             `Typology` == "inversenotch/endscraper" ~ "tool/notch-end scraper",

             `Typology` == "alternatingside scraper" ~ "tool/side scraper",

             `Typology` == "alternatedouble side scraper" ~ "tool/double side scraper",

             `Typology` == "truncation/sidescaraper" ~ "tool/truncated piece",

             `Typology` == "doublenotch" ~ "tool/double notch",

             `Typology` == "borer/awl" ~ "tool/perforator-awl",

             `Typology` == "roundscraper?" ~ "tool/round scraper",

             `Typology` == "doublenotch scraper" ~ "tool/double notch scraper",

             `Typology` == "corerejuvention (core tablet?)" ~ "core/core tablet",

             `Typology` == "truncationاریبب" ~ "tool/truncated piece",

             `Typology` == "retouchedpiece (side scraper)" ~ "tool/side scraper",

             `Typology` == "retochedpice" ~ "tool/retouched piece",

             `Typology` == "truncation/notch" ~ "tool/truncated piece-notch",

             `Typology` == "doublealternate scraper" ~ "tool/double side scraper",

             `Typology` == "awl/ inverse notch" ~ "tool/notch",

             `Typology` == "retouchedpiece (alternate scraper?)" ~ "tool/double side scraper",

             `Typology` == "inverseside scraper" ~ "tool/side scraper",

             `Typology` == "usedcore rejuvention" ~ "core/rejuvenation piece",

             `Typology` == "coreside rejuvention" ~ "core/rejuvenation piece",

             `Typology` == "coreplatform rejuvention flake" ~ "core/rejuvenation piece",

             `Typology` == "corerejuvention platform" ~ "core/rejuvenation piece",

             `Typology` == "coreside rejuvention blade" ~ "core/rejuvenation piece",

             `Typology` == "coreside rejuvention blade??" ~ "core/rejuvenation piece",

             `Typology` == "coreedge rejuvention" ~ "core/rejuvenation piece",

             `Typology` == "coreedge rejivention" ~ "core/rejuvenation piece",

             `Typology` == "retouchedpice (side scraper?)" ~ "tool/side scraper" ,

             `Typology` == "scraper" ~ "tool/side scraper",

             `Typology` == "sickleblade" ~ "tool/sickle shine",

             `Typology` == "coretool (scraper)" ~ "tool/scraper-on core piece",

             `Typology` == "retouchedpiece (scraper)" ~ "tool/side scraper",

             `Typology` == "inversedenticulate?" ~ "tool/denticulate",

             `Typology` == "retouchedpiece (scraper?)" ~ "tool/side scraper",

             `Typology` == "coreside rejuvention flake" ~ "core/rejuvenation piece",

             `Typology` == "truncation/endscraper" ~ "tool/endscraper on truncated piece",

             `Typology` == "retouchedpice(double side scraper?)" ~ "tool/double side scraper",

             `Typology` == "doublescraper" ~ "tool/double side scraper",

             `Typology` == "dendiculate" ~ "tool/denticulate",

             `Typology` == "double-scraper" ~ "tool/double side scraper",

             `Typology` == "pyramidalcore" ~ "core/pyramid",

             `Typology` == "bulletcore" ~ "core/bullet",

             `Typology` == "corerejuvention" ~ "core/rejuvenation piece",

             `Typology` == "pyrmidalcore" ~ "core/pyramid",

             `Typology` == "corepreperation" ~ "core/core preparation piece",

             `Typology` == "crested" ~ "core/crested bladelet",

             `Typology` == "preperationblade" ~ "core/core preperation",

             `Typology` == "burin" ~ "tool/burin",

             `Typology` == "serrateddenticulate" ~ "tool/serrated scraper",

             `Typology` == "saw:serrated denticulate" ~ "tool/serrated scraper",

             `Typology` == "serratedused" ~ "tool/serrated scraper",

             `Typology` == "corepreeration" ~ "tool/core preparation piece",

             `Typology` == "coreprepearation" ~ "tool/core preparation piece",

             `Typology` == "awl" ~ "tool/perforator-awl",

             `Typology` == "notch.inverse" ~ "tool/notch",

             `Typology` == "point.notch" ~ "tool/perforator-point;notch",

             `Typology` == "inversescraper" ~ "tool/side scraper",

             `Typology` == "roundendscraper" ~ "tool/round scraper",

             `Typology` == "alternate.scraper" ~ "tool/double side scraper",

             `Typology` == "convergentscraper/disc scraper" ~ "tool/scraper-convergent scraper",

             `Typology` == "serratedside scraper" ~ "tool/serrated scraper",

             `Typology` == "debri" ~ "NA/NA",

             `Typology` == "distalpart of a drill" ~ "tool/perforator-drill",

             `Typology` == "corepreperation?" ~ "core/core preparation piece",

             `Typology` == "preformend of a borer" ~ "tool/perforator-borer; preform",

             `Typology` == "used" ~ "tool/used",

             `Typology` == "distalend of a inverse sidescraper" ~ "tool/side scraper",

             `Typology` == "retouchedblade" ~ "tool/retouched piece",

             `Typology` == "drill" ~ "tool/perforator-drill",

             `Typology` == "corerejuvention tablet" ~ "core/core tablet",

             `Typology` == "bladeletcore" ~ "core/pyramid",

             `Typology` == "flake-bladecore" ~ "core/mixed",

             `Typology` == "bladelet-corerejuvention" ~ "core/rejuvenation piece",

             `Typology` == "bladelet-corefragment" ~ "core/core fragment",

             `Typology` == "pointedused" ~ "tool/perforator-point",

             `Typology` == "microburin" ~ "tool/microburin",

             `Typology` == "bladeletcore fragment" ~ "core/core fragment",

             `Typology` == "rejuventionface flake" ~ "core/rejuvenation piece",

             `Typology` == "corepreperation flake" ~ "core/core preparation piece",

             `Typology` == "corerejuvention flake"  ~ "core/rejuvenation piece",

             `Typology` == "unfinished  pyramid core" ~ "core/pyramid",

             `Typology` == "bladeletcoretablet" ~ "core/core tablet",

             `Typology` == "bladeletcore rejuvention" ~ "core/rejuvenation piece",

             `Typology` == "debitage" ~ "NA/NA",

             `Typology` == "bladeletcore  rejuvention" ~ "core/rejuvenation piece",

             `Typology` == "inverseserrated scraper" ~ "tool/serrated scraper",

             `Typology` == "backedknife" ~ "tool/backed knife",

             `Typology` == "corepreparation" ~ "core/core preparation piece",

             `Typology` == "primaryflake" ~ "core/primary flake",

             `Typology` == "pointedconvergent scraper on bladelet" ~ "tool/perforator-point",

             `Typology` == "trie" ~ "NA/NA",

             `Typology` == "part of core tablet" ~ "core/core tablet",

             `Typology` == "zaviedar" ~ "NA/NA",

             `Typology` == "convergentscraper" ~ "tool/convergent scraper",

             `Typology` == "preparationflake" ~ "core/core preparation piece",

             `Typology` == "atypicalborer" ~ "tool/perforator-borer",

             `Typology` == "naturalybacked knife" ~ "tool/backed knife-naturally backed knife",

             `Typology` == "pyramidbladelet core" ~ "core/pyramid",

             `Typology` == "flatbladelet core" ~ "core/flat",

             `Typology` == "bladecore" ~ "core/pyramid",

             `Typology` == "bulletbladelet core" ~ "core/bullet",

             `Typology` == "coerejuvention flake" ~ "core/rejuvenation piece",

             `Typology` == "finishedbladelet core" ~ "core/bullet",

             `Typology` == "failed pyramid core?" ~ "core/pyramid",

             `Typology` == "pyramidplain bladelet core" ~ "core/pyramid",

             `Typology` == "coeon flake" ~ "core/core on flake",

             `Typology` == "coeejuvention flake" ~ "core/rejuvenation piece",

             `Typology` == "bladeletcoe rejuvention tablet?" ~ "core/core tablet",

             `Typology` == "pyramidcore" ~ "core/pyramid",

             `Typology` == "bladeletcore tablet" ~ "core/core tablet",

             `Typology` == "pyramidbladeleet core" ~ "core/pyramid",

             `Typology` == "flake" ~ "NA/NA",

             `Typology` == "serrated" ~ "tool/serrated scraper",

             `Typology` == "curvedretouched piece" ~ "tool/retouched piece",

             `Typology` == "pyramidalbladelet core" ~ "core/pyramid",

             `Typology` == "multidirectionalcore fragment" ~ "core/multidirectional core fragment",

             `Typology` == "unidirectionalblade? core fragent" ~ "core/pyramid",

             `Typology` == "backed/corerejuvention" ~ "tool/backed knife",

             `Typology` == "unidirectionalbladelet core" ~ "core/pyramid",

             `Typology` == "unifacialbalde core fragment?" ~ "core/pyramid",

             `Typology` == "flakedetached from a bladelet core" ~ "core/core preparation piece",

             `Typology` == "corepreparation tablet" ~ "core/core tablet",

             `Typology` == "bladeletcore fragent" ~ "core/core fragment-pyramid",

             `Typology` == "alternatescraper??" ~ "tool/double side scraper",

             `Typology` == "endscraper.onrejuvention" ~ "core/rejuvenation piece",

             `Typology` == "inverseserrated" ~ "tool/serrated scraper",

             `Typology` == "patination.omitted" ~ "NA/NA",

             `Typology` == "trapze" ~ "tool/geometric-triangle",

             `Typology` == "trapzoid" ~ "tool/geometric-triangle",

             `Typology` == "borer.drill" ~ "tool/perforator-drill",

             `Typology` == "serraed" ~ "tool/serrated scraper",

             `Typology` == "bladeletcore on a flake" ~ "core/core on flake",

             `Typology` == "geofact" ~ "NA/NA",

             `Typology` == "denticulateborer?" ~ "tool/perforator-borer;denticulate",

             `Typology` == "sideborer-denticulate" ~ "tool/perforator-borer;denticulate",

             `Typology` == "borer.awl" ~ "tool/perforator-awl",

             `Typology` == "roundscraper" ~ "tool/round scraper",

             `Typology` == "truncationon a bladelet core" ~ "tool/truncated piece",

             `Typology` == "truncation.notch" ~ "tool/truncated piece;notch",

             `Typology` == "backed-denticulate" ~ "tool/backed;denticulate",

             `Typology` == "brokendrill" ~ "tool/perforator-drill",

             `Typology` == "drillbroken" ~ "tool/perforator-drill",

             `Typology` == "point/broken drill" ~ "tool/perforator-point",

             `Typology` == "brokendrill?" ~ "tool/perforator-drill",

             `Typology` == "partialyserrated" ~ "tool/serrated scraper",

             `Typology` == "naturallybacked" ~ "tool/backed knife-naturally backed knife",

             `Typology` == "transversescraper" ~ "tool/transverse scraper",

             `Typology` == "naturalybacked" ~ "tool/backed knife-naturally backed knife",

             `Typology` == "awl-notch" ~ "tool/perforator-awl",

             `Typology` == "psedulevalois" ~ "tool/psedulevalois flake",

             `Typology` == "alternatingdenticulate" ~ "tool/denticulate",

             `Typology` == "usedcore fragment" ~ "tool/core tool-used core fragment",

             `Typology` == "multidirectionalmicroflake core" ~ "core/multidirectional flake core",

             `Typology` == "mixedpyramidal core" ~ "core/pyramid",

             `Typology` == "usedbladecore fragment" ~ "tool/core tool-used core fragment",

             `Typology` == "multidirectionalbladelet core fragment" ~ "core/multidirectional bladelet core",

             `Typology` == "unidirectionalbladelet core.cylinder" ~ "core/prismatic",

             `Typology` == "pyrymedalmixed core" ~ "core/pyramid",

             `Typology` == "pyramedial/bullet core" ~ "core/bullet",

             `Typology` == "corefragent" ~ "core/core fragment",

             `Typology` == "pyramedialbladelet core fragment" ~ "core/pyramid",

             `Typology` == "unidirectionalbladelet core fragment" ~ "core/pyramid",

             `Typology` == "unifacialunidirectional bladelet core" ~ "core/pyramid",

             `Typology` == "unidirectionalblade core" ~ "core/pyramid",

             `Typology` == "pyramidalbladelet core fragment" ~ "core/pyramid",

             `Typology` == "corerejuvention?" ~ "core/rejuvenation piece",

             `Typology` == "flatburin? point?" ~ "tool/flat burin",

             `Typology` == "primaryblade" ~ "core/primary blade",

             `Typology` == "trihedralgeometric" ~ "tool/geometric-triangle",

             `Typology` == "lunate" ~ "tool/geometric-lunate",

             `Typology` == "corepaltform rejuvention flake" ~ "core/rejuvenation piece",

             `Typology` == "pressurebladelet core /semi flat" ~ "core/pyramid",

             `Typology` == "failedpercussion blade core/semi pyramedal" ~ "core/pyramid",

             `Typology` == "percussionbladelet core/unidirectional" ~ "core/pyramid-percussion",

             `Typology` == "percussion.unidirectional.blade-flakecore" ~ "core/mixed-percussion",

             `Typology` == "obliqueretouched bladelet" ~ "tool/retouched piece",

             `Typology` == "percussionunidirectional pyramidal  bladelet core" ~ "core/pyramid",

             `Typology` == "lunategeometric" ~ "tool/geometric-lunate",

             `Typology` == "awl on a core rejuvention" ~ "tool/perforator-awl on core rejuvenation piece",

             `Typology` == "truncated" ~ "tool/truncated piece",

             `Typology` == "multiplenotch" ~ "tool/notch",

             `Typology` == "dishedconvergent bladelet" ~ "tool/convergent scraper",

             `Typology` == "bipolarpercussion blade-bladelet core" ~ "core/bipolar-percussion",

             `Typology` == "bladeletbullet core" ~ "core/bullet",

             `Typology` == "alternatingserrated" ~ "tool/serrated scraper",

             `Typology` == "shaplesscore" ~ "core/shapeless",

             `Typology` == "corepreparation?" ~ "core/core preparation piece",

             `Typology` == "obliqueretouched" ~ "tool/retouched piece",

             `Typology` == "corepreparation flake" ~ "core/core preparation piece",

             `Typology` == "backeddenticulate" ~ "tool/backed;denticulate",

             `Typology` == "awlnotch" ~ "tool/perforator-awl;notch",

             `Typology` == "scraperburin?" ~ "tool/burin;scraper",

             `Typology` == "notchround scraper" ~ "tool/round scraper;notch",

             `Typology` == "corerejuvention tablet round scraper" ~ "tool/round scraper on core tablet",

             `Typology` == "alternatescraper" ~ "tool/double side scraper",

             `Typology` == "coreplatform preperation flake" ~ "core/rejuvenation piece",

             `Typology` == "coreplatform rejuvention tablet" ~ "core/core tablet",

             `Typology` == "brokenpyramidal core bladelet" ~ "core/pyramid",

             `Typology` == "pyramidalbladelet core/semi bullet?" ~ "core/bullet",

             `Typology` == "coreside rejuvention flake?" ~ "core/rejuvenation piece-side",

             `Typology` == "unidirectionalpressure bladelet core/unifacila" ~ "core/pyramid",

             `Typology` == "coreplatform rejuvention tablet/or multidirectional core" ~ "core/rejuvenation piece",

             `Typology` == "bulletbladlet core" ~ "core/bullet",

             `Typology` == "cylinderbipolar bladelet core" ~ "core/prismatic",

             `Typology` == "backedknife/core platform rejuvenation" ~ "tool/backed knife",

             `Typology` == "flatunifacial bladelet core" ~ "core/flat",

             `Typology` == "coreprepration flake" ~ "core/core preparation piece",

             `Typology` == "coreplatform preperation" ~ "core/core preparation piece-platform",

             `Typology` == "borerpreform" ~ "tool/perforator-borer",

             `Typology` == "rejuvention" ~ "core/rejuvenation piece",

             `Typology` == "borerpreform?" ~ "tool/perforator-borer",

             `Typology` == "inversnotch" ~ "tool/notch",

             `Typology` == "alternateconvergent" ~ "tool/convergent scraper",

             `Typology` == "unidirectionalbladelet/flake-blade core" ~ "core/mixed",

             `Typology` == "unidirectionalmixed core" ~ "core/mixed",

             `Typology` == "unidirectionalflat bladelet core" ~ "core/flat",

             `Typology` == "" ~ "NA-NA",
             `Typology` == "na" ~ "NA-NA",

    TRUE ~ `Typology`
  ))

tl4 <- tl4 %>%
  mutate(
    typology_split = str_split(Typology, "/", simplify = TRUE),
    `typology-tool-core` = typology_split[, 1],
    tool_or_core = ifelse(str_starts(`typology-tool-core`, "tool"), "tool", "core"),
    `tool-typology` = ifelse(tool_or_core == "tool", typology_split[, 2], NA),
    `core-fragment` = ifelse(tool_or_core == "core", typology_split[, 2], NA)
  ) %>%
  # remove temporary columns
  select(-typology_split, -tool_or_core)

unique(tl4$`tool-typology`)
unique(tl4$`core-fragment`)
```

```{r}
# Exploring datra

tl4 %>% 
  group_by(midpoint) %>% 
  tally() %>% 
  drop_na(midpoint) %>% 
ggplot() + 
  aes(midpoint, n) +
  geom_col()

```

```{r}
# convert all columns to character type (I DONT KNOW WHY I DID THIS!)
tl4 <- tl4 %>%
  mutate(across(everything(), as.character))

# replace 'NA', 'na', or '0' with R's NA
tl4 <- tl4 %>%
  mutate(across(everything(), ~na_if(., "NA"))) %>%
  mutate(across(everything(), ~na_if(., "na"))) %>%
  mutate(across(everything(), ~na_if(., "0")))

print(tl4)

```

```{r}
#refine the table

library(dplyr)

tl_final <- tl4 %>%

  rename(
    blank = Blank2,
    `sickle shine` = shine2,
    length = `Length (mm)` ,
    width = `Width (mm)`,
    thickness = `Thickness (mm)`, 
    typology = Typology,
    utilization = Utilization,
    retouch = Retouch,
    blank = Blank2, 
    `blank part` = `Blank_part`,
    `bulb thickness` = `Bulb Thickness (mm)`, 
    `platform thickness` = `Platform Thickness (mm)`, 
    `platform length` = `platform Length (mm)`, 
    eraillure = Eraillure , 
    cortex = `Cortex (%)`)

```

```{r}
tl_final <- tl_final %>%
  mutate(al = paste0(toupper(area), "_", layer)) %>%
  relocate(al, .after = layer)

```

```{r}
tl_final <- tl_final  %>%

 select(lithic_id, 
         TA, 
         area, 
         layer,
        al,
        cortex,
         depth, upper, lower, midpoint, 
         `length`, `width`, `thickness`, 
         typology, 
        utilization, 
        `sickle shine`,
        retouch, 
        `typology-tool-core`,
        `tool-typology`,
         blank, `blank part`,

         `core-typology`, `core-fragment`,`core-technology`,

        cortex)
```

```{r}
#set the value of utilization to NA only for the rows where `sickle shine` == 1. 
#All other rows should remain unchanged

#unsuccessful, should try again later

```

```{r}
# Update the 'typology-tool-core' column based on 'Retouch'
tl_final <- tl_final %>% 
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",

      TRUE ~ `typology-tool-core`      
    )
  )

print(tl_final)

```

```{r}

tl_final <- tl_final %>% 
  mutate(
    `tool-typology` = case_when(
      utilization == 1 ~ "utilized tool",
      `sickle shine` == 1 ~ "sickle shine",
      TRUE ~ `tool-typology`               
    )
  )

print(tl_final)

```

```{r}
tl_final <- tl_final %>% 
  mutate(
    `tool-typology` = case_when(

     # If Retouch is 1 and tool-typology is NA
       retouch == 1 & is.na(`tool-typology`) ~ "retouched piece",
       `sickle shine` == 1 & is.na(`tool-typology`) ~ "sickle shine",

      # For all other cases 
      TRUE ~ `tool-typology`                                       
    )
  )

# Show updated data frame
unique(tl_final$`tool-typology`)

```

```{r}

# Creating tooltypegroup

tl_final <- tl_final %>%
  mutate(
    tooltypegroup = case_when(
      `tool-typology` %in% 
        c("end scraper", "side scraper", "double side scraper", "convergent scraper", "round scraper", "perforator-convergent scraper", "perforator-borer; preform", "scraper-on core piece", "endscraper on truncated piece", "round scraper", "transverse scraper", "core tool-used core fragment", "round scraper on core tablet") ~ "scraper",
      `tool-typology` %in% 
        c("perforator-borer", "perforator-drill", "perforator-point", "perforator-awl", "perforator-point;notch", "perforator-awl on core rejuvenation piece", "perforator-awl;notch", "perforator-borer;denticulate") ~ "Perforator",
      `tool-typology` %in% 
        c("backed knife", "backed knife-naturally backed knife", "backed;denticulate") ~ "backed pieces",
      `tool-typology` %in% c("double notch", "denticulate", " denticulate", "notch-denticulate", "notch") ~ "denticulate-notch",
      `tool-typology` %in% c("notch-side scraper", "notch-end scraper", "double notch scraper", "round scraper;notch") ~ "scraper-notch",
      `tool-typology` %in% c("geometric-triangle", "geometric-lunate") ~ "geometric",
      `tool-typology` == "serrated scraper" ~ "serrated scraper",
      `tool-typology` %in% c("truncated piece", "truncated piece-notch", "truncated piece;notch", "notch;truncated piece") ~ "truncated pieces",

      # Represents NA group
      `tool-typology` %in% c("core preparation piece", "used", "psedulevalois flake") ~ NA_character_, 
      `tool-typology` %in% c("burin", "flat burin", "burin;scraper", "burin-notch-scraper") ~ "burin",
      `tool-typology` == "microburin" ~ "microburin",
      `tool-typology` == "retouched piece" ~ "retouched piece",
      `tool-typology` == "utilized tool" ~ "utilized tool",
      `tool-typology` == "sickle shine" ~ "sickle shine",

      #classify any remaining values as NA
      TRUE ~ NA_character_ 
    )
  )

# Checking the unique values of tooltypegroup
unique(tl_final$tooltypegroup)

```

```{r}
tl_final <- tl_final %>%
  mutate(
    `core-fragment` = if_else(
      !is.na(`core-typology`),  
      NA_character_,            
      `core-fragment`           
    )
  )

print(tl_final)

```

```{r}

library(dplyr)
library(stringr)

tl_final <- tl_final %>%
  mutate(
    `core-fragment` = if_else(
      !is.na(`core-typology`),
      NA_character_,
      `core-fragment`

  ) 
  ) %>%

  mutate(`core-fragment` = case_when(
    `core-fragment` == "rejuvenation piece" ~ "rejuvenation piece",

             `core-fragment` == "rejuvention piece-side" ~ "rejuvenation piece",

             `core-fragment` == "core preparation piece-platform" ~ "core tablet",

             `core-fragment` == "core preparation" ~ "rejuvenation piece",

             `core-fragment` == "core preparation piece" ~ "rejuvenation piece",

             `core-fragment` == "core preperation" ~ "rejuvenation piece",

             `core-fragment` == "pyramid" ~ "pyramid",

             `core-fragment` == "pyramid core" ~ "pyramid",

             `core-fragment` == "pyramid-unidirectional" ~ "pyramid",

             `core-fragment` == "core fragment-pyramid" ~ "pyramid",

             `core-fragment` == "cylinder" ~ "cylinder/prismatic",

             `core-fragment` == "prismatic" ~ "cylinder/prismatic",

             `core-fragment` == "shapeless" ~ "cylinder/prismatic",

             `core-fragment` == "multidirectional core fragment" ~ "shapeless",

             `core-fragment` == "multidirectional flake core" ~ "shapeless",

             `core-fragment` == "multidirectional bladelet core" ~ "shapeless",

             `core-fragment` == "mixed core" ~ "shapeless",

             `core-fragment` == "mixed-percussion" ~ "shapeless",

             `core-fragment` == "mixed" ~ "shapeless",

             `core-fragment` == ("bipolar-percussion") ~ "bipolar",

             `core-fragment` == "flat" ~ "flat",

             `core-fragment` == "core on flake" ~ "core on flake",

             `core-fragment` == "bullet" ~ "bullet",

    TRUE ~ `core-fragment`

  ))

```

```{r}

tl_final <- tl_final %>%
  mutate(
    core = case_when(
      `core-fragment` %in% c("pyramid", "pyramid core", "pyramid-unidirectional", "core fragment-pyramid", "cylinder/prismatic", "shapeless", "multidirectional core fragment", "multidirectional flake core", "multidirectional bladelet core", "mixed core", "mixed-percussion", "mixed", "bipolar-percussion", "flat", "core on flake", "bullet") ~ `core-fragment`,
      TRUE ~ NA_character_
    ),
    corefrag = case_when(
      `core-fragment` %in% c("rejuvenation piece", "rejuvention piece-side", "core preparation piece-platform", "core preparation", "core preparation piece", "core preperation", "crested bladelet" , "primary flake" , "primary blade" ) ~ `core-fragment`,
      TRUE ~ NA_character_
    )
  )

unique(`tl_final`$`core`)
unique(`tl_final`$`corefrag`)
```

```{r}

library(dplyr)

tl_final <- tl_final %>%
  # Convert empty strings to NA in `core-fragment` column
  mutate(`core-fragment` = if_else(`core-fragment` == "", NA_character_, `core-fragment`)) %>%
  # Convert "pyramid-percussion" to "pyramid" in `core-fragment` column
  mutate(`core-fragment` = 
           if_else(
             `core-fragment` == "pyramid-percussion", "pyramid", `core-fragment`)) %>%
  # Update existing `core` column with selected values from `core-fragment` column
  mutate(core = if_else(`core-fragment` %in% 
                          c("pyramid", 
                            "cylinder/prismatic", 
                            "flat", 
                            "bullet", 
                            "core on flake", 
                            "pyramid"), 
                        `core-fragment`, 
                        core))

unique(tl_final$`core-fragment`)
unique(tl_final$`corefrag`)
unique(tl_final$`core`)
```

```{r}

library(dplyr)

tl_final <- tl_final %>%
  # Add 'core tablet' and 'core fragment' from `core-fragment` to `corefrag`
  mutate(corefrag = case_when(
    `core-fragment` == "core tablet" ~ "core tablet",
    `core-fragment` == "core fragment" ~ "core fragment",
    TRUE ~ corefrag
  ))

unique(tl_final$`core-fragment`)
unique(tl_final$`corefrag`)
unique(tl_final$`core`)
```

```{r}

```

```{r}

library(dplyr)

tl_final <- tl_final %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(corefrag = case_when(
    `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
    `core-typology` == "core fragment" ~ "core fragment",
    TRUE ~ corefrag
  )) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology` 
  mutate(`core-typology` = if_else(`core-typology` %in% c("rejuvenation piece", "core fragment"), NA_character_, `core-typology`))

unique(tl_final$`core-typology`)
unique(tl_final$`corefrag`)
unique(tl_final$`core-fragment`)

```

```{r}

library(dplyr)

# Update the 'blank' column based on conditions in the 'core-fragment' column
tl_final <- tl_final %>%
  mutate(blank = if_else(`core-fragment` %in% c(
    "core fragment", "core tablet", "rejuvenation piece", "pyramid", "crested bladelet", 
    "shapeless", "primary flake", "flat", "bullet", "core on flake", "cylinder/prismatic",
    "primary blade", "rejuvenation piece-side", "core platform rejuvention"),
    NA_character_, 
    blank))

print(tl_final)
unique(tl_final$`core-fragment`)
unique(tl_final$`blank`)

```

Summary tables: 1.AREA

```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
    `Cores (n)` = sum(!is.na(core) & core != ""),
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) & corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

print(summary_table_area)

```

```{r}
library(kableExtra)

kable(summary_table_area, 
      caption = "Table: Summary by Area", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

2.LAYER

```{r}
# Create a summary table
summary_table_layer <- tl_final %>%
  group_by(layer) %>%  # Note that 'area' is replaced with 'layer'
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```

```{r}
# Generate the table
kable(summary_table_layer, 
      caption = "Table: Summary by Layer", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

AREA_LAYER:

```{r}
# Create a summary table based on the 'al' column
summary_table_al <- tl_final %>%
  group_by(al) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```

```{r}
# Generate the table
library(kableExtra)

kable(summary_table_al, 
      caption = "Table: Summary by Al", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

DEPTH:

```{r}
tl_final <- tl_final %>%
  mutate(depth = gsub("cm", "", depth))
```

```{r}
# Create a summary table based on the 'depth' column
summary_table_depth <- tl_final %>%
  group_by(depth) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```

```{r}
# Generate the table
library(kableExtra)

kable(summary_table_depth, 
      caption = "Table: Summary by Depth", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

```{r}
# Create a summary table based on the 'depth', 'area', and 'layer' columns
summary_table_depth_area_layer <- tl_final %>%
  group_by(depth, area, layer, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

# Generate the table
library(kableExtra)

kable(summary_table_depth_area_layer, 
      caption = "Table: Summary by Depth, Area, and Layer", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

Techno-Typo:

```{r}
tl_final_area <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1"))

```

```{r}

library(dplyr)

# Prepare data for Summary table-area
summary_table_area <- tl_final_area %>% 
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(area) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
summary_table_area <- summary_table_area %>% 
  add_row(
    area = "Total",
    blade_count = sum(summary_table_area$blade_count, na.rm = TRUE),
    flake_count = sum(summary_table_area$flake_count, na.rm = TRUE),
    bladelet_count = sum(summary_table_area$bladelet_count, na.rm = TRUE),
    tool_count = sum(summary_table_area$tool_count, na.rm = TRUE),
    core_count = sum(summary_table_area$core_count, na.rm = TRUE),
    core_rejuvenation_count = sum(summary_table_area$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count = sum(summary_table_area$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count = sum(summary_table_area$utilized_tool_count, na.rm = TRUE),
    shine_tool_count = sum(summary_table_area$shine_tool_count, na.rm = TRUE),
    total = sum(summary_table_area$total, na.rm = TRUE)  # Explicitly setting the total
  )

```

```{r}
# Prepare data for Summary table-al
summary_table_al <- tl_final_area %>% 
  group_by(al) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
summary_table_al <- summary_table_al %>% 
  add_row(
    al = "Total",
    blade_count = sum(summary_table_al$blade_count, na.rm = TRUE),
    flake_count = sum(summary_table_al$flake_count, na.rm = TRUE),
    bladelet_count = sum(summary_table_al$bladelet_count, na.rm = TRUE),
    tool_count = sum(summary_table_al$tool_count, na.rm = TRUE),
    core_count = sum(summary_table_al$core_count, na.rm = TRUE),
    core_rejuvenation_count = sum(summary_table_al$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count = sum(summary_table_al$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count = sum(summary_table_al$utilized_tool_count, na.rm = TRUE),
    shine_tool_count = sum(summary_table_al$shine_tool_count, na.rm = TRUE),
    total = sum(summary_table_al$total, na.rm = TRUE)  # Explicitly setting the total
  )

# View the summary table
print(summary_table_al)

```

```{r}
summary_table_al <- summary_table_al %>%
  filter(!(al %in% c("D1_NA", "TP1_NA")))

library(kableExtra)

kable(summary_table_al, 
      col.names = c("Area-Layer", "Blade", "Flake", "Bladelet", "Tools", "Cores", "Core Rejuvenations", "Retouched Tools", "Utilized Tools", "Sickle Shine", "Total"), 
      caption = "Table: Summary ", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

```{r}
#summary table-area

library(knitr)
library(dplyr)
# Rename columns
summary_table_area <- summary_table_area %>%
  rename(
    "Area" = area,
    "Blade" = blade_count,
    "Flake" = flake_count,
    "Bladelet" = bladelet_count,
    "Retouched Tools" = retouched_tool_count,
    "Tools" = tool_count,
    "Utilized Tools" = utilized_tool_count,
    "Sickle Shine" = shine_tool_count,
    "Cores" = core_count,
    "Core rejuvenations pieces" = core_rejuvenation_count,
    "Total" = total
  ) %>%
  mutate(
    Area = if_else(row_number() == n() & is.na(Area), "Total", Area)
  ) %>%
  mutate(
    Area = recode(Area, "ta" = "TA", "d1" = "D1", "tp1" = "TP1", "Total" = "Total")
  )
library(kableExtra)

kable(summary_table_area, 
      caption = "Table 1: Count of debitage type and tools in three different main areas.", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 12) %>%

  row_spec(0, font_size = 11)

```

```{r}
# summary table-core and core rejuvenation
```

```{r}
# summary table for tools: retouched-utelized-sickle

```

```{r}

library(dplyr)
library(tidyr)

# Create a summary table
summary_table_tool_area <- tl_final_area %>%
  filter(!is.na(tooltypegroup), tooltypegroup != "") %>%  # Exclude rows where 'tooltypegroup' is NA or empty
  group_by(area, tooltypegroup) %>%  # Group the data by area and tooltypegroup
  tally() %>%  # Count the number of rows in each group
  arrange(area, tooltypegroup) %>%  # Arrange by area and tooltypegroup
  spread(key = area, value = n, fill = 0)  # Spread the 'area' column into multiple columns

# Add a "Total" column that sums each row
summary_table_tool_area <- summary_table_tool_area %>%
  mutate(Total = rowSums(select(., -tooltypegroup)))

total_row <- summary_table_tool_area %>%
  select(-tooltypegroup) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  mutate(tooltypegroup = "Total")

summary_table_tool_area <- bind_rows(summary_table_tool_area, total_row)

print(summary_table_tool_area)

```

```{r}
library(kableExtra)

kable(summary_table_tool_area, 
      col.names = c("Tools", "D1", "TA", "TP1", "Total"),
      caption = "Table: Tool Typology in three different main areas.", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  column_spec(1, extra_css = "text-align:left;") %>%  
  row_spec(0, bold = TRUE, font_size = 14)

```

```{r}
# Create the summary table based on layer
summary_table1_layer <- tl_final_area %>%
  group_by(layer) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    numb_cores = sum(!is.na(`core-typology`), na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

```

volume for each area:

```{r}
# Define the volumes for each area 
volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375) # in m^3
)

# Count the number of artifacts by area
artifact_counts <- tl_final_area %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Join artifact_counts with volumes by area
lithic_density_df <- left_join(artifact_counts, volumes, by = "area")

# Compute lithic density for each area and round it to 2 decimal places
lithic_density_df <- lithic_density_df %>%
  mutate(lithic_density = round(n / volume, 2)) # number of artifacts / volume

# Create the summary table
summary_table2_area <- tl_final_area %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

# Merge the density data with the summary table
summary_table2_area <- left_join(summary_table2_area, lithic_density_df[, c("area", "lithic_density")], by = "area") %>%
  mutate(area = recode(area, 'ta' = 'TA', 'tp1' = 'TP1', 'd1' = 'D1'),
         lithic_volumetric_density = lithic_density)

# Remove the 'lithic_density' column
summary_table2_area <- summary_table2_area %>% 
  select(-lithic_density)

# Print out the resulting data frame
print(summary_table2_area)

```

```{r}

                      #########DO NOT RUN0

# Define the volumes for each area based on the given information
volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375) # in m^3
)

# Count the number of artifacts by area
artifact_counts <- tl_final_area %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Calculate retouch frequency with explicit NA check
retouch_freq <- tl_final_area %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  summarise(retouchFreq = ifelse(all(is.na(retouch)), NA, mean(!is.na(retouch))), .groups = 'drop')

# Compute lithic density for each area and round it to 2 decimal places
lithic_density_df <- left_join(artifact_counts, volumes, by = "area") %>%
  mutate(lithic_density = round(n / volume, 2))

# Create the summary table
summary_table2_area <- tl_final_area %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

# Merge all the calculated data
summary_table2_area <- left_join(summary_table2_area, lithic_density_df[, c("area", "lithic_density")], by = "area")
summary_table2_area <- left_join(summary_table2_area, retouch_freq, by = "area")

# Final modifications and print
summary_table2_area <- summary_table2_area %>%
  mutate(
    area = recode(area, 'ta' = 'TA', 'tp1' = 'TP1', 'd1' = 'D1'),
    lithic_volumetric_density = lithic_density
  ) %>%
  select(-lithic_density)  # Remove the unwanted column

print(summary_table2_area)

```

```{r}
# generate table 2 - area

library(kableExtra)

kable(summary_table2_area, 
      caption = "Table 2. Lithic counts for each of the areas in Tula'i", 
      align = 'c',
      col.names = c("Area", "Number of Retouched Lithics", "Total Number of Lithics", "Retouch frequency (%)", "Lithic Volumetric Density(pcs/(m^3)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 11) %>%
  column_spec(1, width = "1cm") %>% 
  column_spec(2, width = "1cm") %>%

  column_spec(3, width = "1cm") %>% 
  column_spec(4, width = "1cm") %>% 
  column_spec(5, width = "1cm")

```

```{r}

                                            ##### DO NOT RUN ######

# layer-Create the summary table
summary_table2_layer <- tl_final_area %>%
  filter(layer %in% c(c("a", "b", "c" , "d", "e"))) %>%  
  group_by(layer) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    numb_cores = sum(!is.na(`core-typology`), na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

# Create a data frame for volumetric density values
density_data_layer <- data.frame(
  layer = c("a", "b", "c" , "d", "e"),
  `lithic_volumetric_density` = round(c(1.575, 6.12, 2.115, 47, 13), 2)
)

summary_table2_layer <- merge(summary_table2_layer, density_data_layer, by = "layer")

```

WABI_AL Ta-e → 13 m3 Ta-d → 45 m3 D1-c → 0.675 m3 Tp-c → 1.44 m3 D1-a → 1.575 m3

```{r}
# Create a tibble 
summary_table2_area <- tibble(
  area = c("D1_a", "D1_c", "TA_d", "TA_e", "TP1_c"),
  lithic_volumetric_density = c(51.2, 110.9, 45.1, 100.2, 88.3),
  percent_retouch = c(4.9, 9.3, 15.3, 18.0, 9.8)
)

```

```{r}

library(tidyverse)
library(ggpmisc)

# Create the data frame
summary_table2_area <- tibble(
  area = c("D1_a", "D1_c", "TA_d", "TA_e", "TP1_c"),
  lithic_volumetric_density = c(51.43, 111.11, 45.13, 100.08, 212.50),
  retouch_frequency = c(0.05, 0.09, 0.07, 0.07, 0.10)
)

# Create the ggplot
WABI_AL <- ggplot(summary_table2_area, 
            aes(x = lithic_volumetric_density,
                                     y = retouch_frequency)) +

  # Add points
  geom_point(color = 
               "black", 
             size = 2) +

  # line with log scale
  stat_smooth(method = "lm", 
              formula = y ~ log(x), 
              se = TRUE, aes(group = 1), 
              color = "blue", 
              size = 0.5) +

  # Add text labels for each point
  geom_text(aes(label = area), 
            vjust = -1, 
            hjust = -0.5, 
            size = 2.5) +

  # Add regression equation, R-squared, and p-value
  stat_poly_eq(aes(label = paste(stat(eq.label), stat(rr.label), stat(p.value.label), sep = "~~~")),
               formula = y ~ log(x), 
               parse = TRUE, 
               size = 2) +

  # Add titles and labels
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  scale_x_log10() +

  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

print(WABI_AL)

```

```{r}

library(googlesheets4)

sum_depth_sheet <- "https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0"

sum_depth_sheet <- read_sheet(sum_depth_sheet)

sum_depth_sheet <- sum_depth_sheet %>%
  mutate(area_layer_level = paste0(toupper(area), "_", layer, "_", level)) %>%
  relocate(area_layer_level, .after = level)

sum_depth_sheet <- sum_depth_sheet %>%
  mutate(retouch_freq = round(retouch_tool / total_lithics, 2),
         lithic_dens = round(total_lithics / volume, 2)
  )

write_sheet(sum_depth_sheet, ss = sheet_url)

```

```{r}
# Create the ggplot
WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  geom_point(color = "black", size = 2) +

  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              size = 0.5) +

  geom_text(aes(label = area_layer_level), 
            vjust = 1.5, 
            hjust = 0.5, 
            size = 2,
            check_overlap = TRUE) +

  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                 stat(rr.label), 
                                 stat(p.value.label), 
                                 sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +

  # Add titles and labels
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

print(WABI_Level)

```

```{r}
# Create the ggplot with y-axis on a log scale
WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  # Add points with different shapes based on area
  geom_point(aes(shape = area), size = 2) +

  # Add smoothing line
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              size = 0.5) +

  # Add regression equation, R-squared, and p-value
  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                 stat(rr.label), 
                                 stat(p.value.label), 
                                 sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +

  # Add titles and labels
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  # Apply log scale to y-axis
  scale_y_log10() +

  # Add shape legend
  scale_shape_manual(values = c(16, 17, 25)) +

  # Customize theme
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

# Print the plot
print(WABI_Level)

```

```{r}
# log-scale

WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  # Add points
  geom_point(color = "black", size = 2) +

  # Add smoothing line
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              size = 0.5) +

  # Add text labels for each point
  geom_text(aes(label = area_layer_level), 
            vjust = 1.5, 
            hjust = 0.5, 
            size = 2,
            check_overlap = TRUE) +

  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                 stat(rr.label), 
                                 stat(p.value.label), 
                                 sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +

  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  scale_y_log10() +

  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

print(WABI_Level)

```

```{r}

# Create the ggplot
WABI_Level <- ggplot(sum_depth_sheet, 
            aes(x = lithic_dens,
                y = retouch_freq)) +

  # Add points
  geom_point(color = 
               "black", 
             size = 2) +

  # line with log scale

stat_smooth(method = "lm", 
            formula = y ~ log(x), 
            se = TRUE, aes(group = 1), 
            color = "blue", 
            size = 0.5)

  # Add text labels for each point
  geom_text(aes(label = area_layer_level), 
            vjust = -1, 
            hjust = -0.5, 
            size = 2.5) +

  # Add regression equation, R-squared, and p-value
  stat_poly_eq(aes(label = paste(stat(eq.label), stat(rr.label), stat(p.value.label), sep = "~~~")),
               formula = y ~ log(x), 
               parse = TRUE, 
               size = 2) +

  # Add titles and labels
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  scale_x_log10() +

  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

print(WABI_Level)

```

WABI_AREA

```{r}
# regression plot-area
library(ggplot2)
library(ggpmisc)

# Plotting
p_area <- ggplot(summary_table2_area, aes(x = lithic_volumetric_density, y = percent_retouch)) +

  # Blue regression line with shaded area
  geom_smooth(method = "lm", 
              se = TRUE, 
              aes(group = 1), 
              color = "blue") +

  # Points colored by area
  geom_point(color = "black") +

  # Labels points with area names
  geom_text(aes(label = area),
            vjust = -1, 
            hjust = -0.5) +

  # Regression equation, R-squared, and p-value
  stat_poly_eq(aes(label = paste(stat(eq.label), stat(rr.label), stat(p.value.label), sep = "~~~")),
               formula = y ~ x,
               parse = TRUE,
               size = 3) +

  # Titles and labels
  labs(title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
       x = "Lithic Volumetric Density",
       y = "Retouch Frequency (%)")

print(p_area)

```

```{r}
# Fit the linear model
linear_model <- lm(percent_retouch ~ lithic_volumetric_density, data = summary_table2_area)

# Display the summary of the model to get coefficients, R-squared, and p-value
summary_linear_model <- summary(linear_model)

# Extract the coefficients (intercept and slope)
coefficients <- summary_linear_model$coefficients

# Extract the R-squared value
r_squared <- summary_linear_model$r.squared

# Extract the p-value for the slope (lithic_volumetric_density)
p_value_slope <- coefficients["lithic_volumetric_density", "Pr(>|t|)"]

cat("Coefficients:\n")
print(coefficients)

cat("\nR-squared value:", r_squared)

cat("\nP-value for slope:", p_value_slope)

```

```{r}
#whole assemblage

  # Summarize data by Area
summary_table_tech <- tl_final_area %>%
  group_by(area) %>%

  summarise(
   blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),

    bullet_core_count = sum(`core-typology` == "bullet", na.rm = TRUE),
    flat_core_count = sum(`core-typology` == "flat", na.rm = TRUE),
    pyramid_core_count = sum(`core-typology` == "pyramid", na.rm = TRUE),
    cylinder_core_count = sum(`core-typology` %in% c("prismatic", "cylinder"), na.rm = TRUE),
    shapeless_core_count = sum(`core-typology` == "shapeless", na.rm = TRUE),

    core_fragment_count = sum(`corefrag` == "core fragment", na.rm = TRUE),
    core_rejuvention_count = sum(`corefrag` == "rejuvenation piece", na.rm = TRUE),
    core_tablet_count = sum(`corefrag` == "core tablet", na.rm = TRUE),
    core_crested_count = sum(`corefrag` == "crested bladelet", na.rm = TRUE),
    core_primarypiece_count = sum(`corefrag`  %in% c("primary flake", "primary blade"), na.rm = TRUE),

   total = n()
  )
```

```{r}
#plot_blank_type

library(dplyr)
library(ggplot2)
library(ggbeeswarm)

# Data filtering and plotting
tl_final_area %>% 
  filter(blank %in% c('bladelet', 'blade', 'flake')) %>%
  ggplot(aes(x = area, fill = blank)) +
  geom_bar(position = "dodge") +
  labs(
    x = "Area",
    y = "Number",
    fill = "Blank Type"
  ) +
  theme_minimal()

```

```{r}

```

```{r}

```

```{r}
# Load required packages
library(ggplot2)
library(cowplot)

# Filter the data to exclude the value 0 and keep only blade, bladelet, and flake
tl4_filtered <- tl4_tool_Blank %>%
  filter(Blank2 %in% c("blade", "bladelet", "flake"))

# Plot Utilization by Blank2
plot_utilization <- ggplot(tl4_filtered, aes(Blank2, Utilization)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Utilization by Blank",
       x = "Blank2",
       y = "Utilization") +
  theme_minimal()

# Plot Retouch by Blank
plot_retouch <- ggplot(tl4_filtered, aes(Blank2, Retouch)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Retouch by Blank",
       x = "Blank2",
       y = "Retouch") +
  theme_minimal()

# Plot shine by Blank
plot_shine <- ggplot(tl4_filtered, aes(x = Blank2, y = shine2)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "shine by Blank",
       x = "Blank2",
       y = "shine2") +
  theme_minimal()

# Combine the plots using cowplot or patchwork
plot_combined <- cowplot::plot_grid(plot_utilization, plot_retouch, plot_shine, ncol = 3)

print(plot_combined)

```

```{r}

#|label: fig_
#|fig-cap:  vs Debitage
#|fig-height: 5

# Filter data
tl4__blank %>% 
  filter(Blank2 %in% c('tp1', 'ta', 'd1')) %>% 
  select(Blank2, 
         `Width (mm)`, 
         `Thickness (mm)`, 
         `Length (mm)`) %>%
mutate(`Thickness (mm)` = as.numeric(`Thickness (mm)`))

# Pivot data to longer format
long_data <- tl3_dimention_blank %>% 
  pivot_longer(cols = c(`Width (mm)`, `Thickness (mm)`, `Length (mm)`), 
               names_to = "Dimension", 
               values_to = "Value")

# Create boxplot
ggplot(long_data, aes(x = area, y = Value)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.01) +
  facet_wrap(~Dimension, scales = "free_y") +
  scale_y_log10() +
  theme_minimal()

```

```{r}
library(ggbeeswarm)
tl3_dimention_blank <- tl3 %>% 
  filter(area %in% tl3_area_tally_greater_than_n$area) %>% 
  pivot_longer(-area) %>% 
  ggplot() +
  aes(area,
      value) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.01) +
  facet_wrap( ~ name,
              scales = "free_y") +
  scale_y_log10() +
  theme_minimal()

```

```{r}
library(ggbeeswarm)
tl3 %>% 
  filter(area %in% tl3_area_tally_greater_than_n$area) %>% 
  pivot_longer(-area) %>% 
  ggplot() +
  aes(area,
      value) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.01) +
  facet_wrap( ~ name,
              scales = "free_y") +
  scale_y_log10() +
  theme_minimal()

```

```{r}
#| label: fig-Width-of-Artifacts-Area-A-by-depth
#| fig-cap: "Width of Artifacts by depth (Area A)"
#| fig-height: 5
tl %>%
  filter(AREA == "A") %>%
  select(
         `Width (mm)`,
         `DEPTH (cm)`) %>%
  ggplot() +
  aes(x = `DEPTH (cm)`,
      y = `Width (mm)`) +
  geom_boxplot() +
  theme_minimal()
```

```{r}
#| label: Width of Artifacts (Area A)
#| fig-cap: "Width of Artifacts (Area A)"
#| fig-height: 5

tl1 %>%
  filter(AREA == "TP1") %>%
  select(AREA, `Width (mm)`) %>%
  ggplot() +
  aes(x = `Width (mm)`) +
  geom_histogram(binwidth = 2) +
  labs(x = "Width of Debitage (mm)", 
       y = "Frequency", 
       title = "Histogram of Width for Area TP1" )+
  theme_minimal()
```

WABI:

```{r}
# regression plot-layer
library(ggplot2)
library(ggpmisc)

# Plotting
p <- ggplot(summary_table2_layer, 
            aes(lithic_volumetric_density, 
                percent_retouch)) +

  geom_smooth(method = "lm", 
              se = TRUE, 
              aes(group = 1), 
              color = "blue") +

  # Points colored by layer
  geom_point(color = "black") +

  # Labels points with layer names
  geom_text(aes(label = layer),
            vjust = -1, 
            hjust = -0.5) +

  # Regression equation, R-squared, and p-value
  stat_poly_eq(aes(label = paste(stat(eq.label), stat(rr.label), stat(p.value.label), sep = "~~~")),
               formula = y ~ x,
               parse = TRUE,
               size = 3) +

  # Titles and labels
  labs(title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
       x = "Lithic Volumetric Density",
       y = "Retouch Frequency (%)")

print(p)
```

```{r}
#barplot_layer

tl_final_area %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d")) %>%
  filter(!is.na(corefrag)) %>% 
  select(layer, corefrag) %>%
  ggplot() +
  aes(x = corefrag) +
  geom_bar() +  
  labs(x = "Core and Core fragments", y = "Frequency") +
  facet_wrap(~layer, ncol = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
#Core-barplot_area

tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  filter(!is.na(corefrag)) %>% # Exclude rows where corefrag is NA
  select(area, corefrag) %>%
  ggplot() +
  aes(x = corefrag) +
  geom_bar() +
  labs(x = "Core and Core fragments",
       y = "Frequency") +
  facet_wrap(~area, ncol = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}

library(ggplot2)
library(dplyr)
filtered_data <- tl_final %>%
  filter(!is.na(`tool-typology`) & `tool-typology` != "" & `tool-typology` != "NA")
ggplot(data = filtered_data) + 
  aes(y = `tool-typology`) + 
  geom_bar() +
  labs(x = "Frequency",
       y = "Tool Type") +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 20, hjust = 1, vjust = 0.5))

```

```{r}

filtered_data <- tl_final %>%
  filter(!is.na(``) & `` != "")

# Create the plot using the filtered data
ggplot(data = filtered_data) + 
  aes(y = ``) + 
  geom_bar() +
  labs(y = "Tool Type",
       x = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}

```

PCA-LAYER

```{r}
#prepare the data for PCA-layer

# Calculating Blank Frequency
blank_freq <- tl_final_area %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(blank)) %>%
  group_by(layer) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Calculating Core Frequency
core_freq <- tl_final_area %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(corefrag)) %>%
  group_by(layer) %>%
  summarise(coreFreq = sum(corefrag %in% c("core fragment", "bullet", "flat", "pyramid", "cylinder/prismatic", "core rejuvenation piece", "shapeless", "core on flake", "bipolar")) / sum(!is.na(corefrag)), .groups = 'drop')

# Calculating Tool Diversity using Menhinick's Index
tool_div <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(tooltypegroup)) %>%
  group_by(layer) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(sum(!is.na(tooltypegroup))), .groups = 'drop')

# Combining all the calculated frequencies into one tibble
final_table <- blank_freq %>%
  inner_join(core_freq, by = "layer") %>%
  inner_join(tool_div, by = "layer")

# Printing the final table
print(final_table)

```

```{r}
# Blank Frequency
blank_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(blank)) %>%
  group_by(layer) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Calculating Core Frequency
core_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "e or d")) %>%
  group_by(layer) %>%
  summarise(coreFreq = sum(corefrag %in% c("core fragment", "bullet", "flat", "pyramid", "cylinder/prismatic", "core rejuvenation piece", "shapeless", "core on flake", "bipolar")) / n(), .groups = 'drop')

# Calculating Tool Diversity using Menhinick's Index
tool_div <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "e or d"), !is.na(tooltypegroup)) %>%
  group_by(layer) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(sum(!is.na(tooltypegroup))), .groups = 'drop')

# Calculating Retouch Frequency
retouch_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d")) %>%
  group_by(layer) %>%
  summarise(retFreq = sum( == "retouched piece", na.rm = TRUE) / n(), .groups = 'drop')

# Combining all the calculated frequencies, including retouch frequency, into one tibble
final_table <- blank_freq %>%
  inner_join(core_freq, by = "layer") %>%
  inner_join(tool_div, by = "layer") %>%
  inner_join(retouch_freq, by = "layer")

# Printing the final table
print(final_table)

```

```{r}
# Joining calculated frequencies
final_table <- blank_freq %>%
  inner_join(core_freq, by = "layer") %>%
  inner_join(tool_div, by = "layer") %>%
  inner_join(retouch_freq, by = "layer")

# Converting layer to row names
final_table_rownames <- final_table %>%
  column_to_rownames(var = "layer")

```

```{r}
# Compute the PCA
library(FactoMineR)

res_pca <-
  PCA(final_table_rownames,
      graph = FALSE)

```

```{r}
library(factoextra)

# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca,
             axes = 1,   
             choice = "var")

```

```{r}
library(factoextra)

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca,
             axes = 2,   
             choice = "var")

```

```{r}
library(factoextra)

# Creating a biplot
fviz_pca_biplot(res_pca)
```

```{r}
library(factoextra)
fviz_screeplot(res_pca)
```

```{r}
fviz_pca_var(res_pca)
```

```{r}

library(FactoMineR)
library(factoextra)

# Run PCA
res_pca <- PCA(final_table_rownames, graph = FALSE)

# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca, axes = 1, choice = "var")

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca, axes = 2, choice = "var")

# Visualize eigenvalues and explained variance
fviz_screeplot(res_pca)

# Create a biplot of the PCA
fviz_pca_biplot(res_pca)

# Visualize the variables on the principal components
fviz_pca_var(res_pca)

```

```{r}

library(FactoMineR)
library(factoextra)

# Run PCA
res_pca <- PCA(final_table_rownames, graph = FALSE)

# Get the loadings (coefficients) of original variables for each principal component
loadings_matrix <- res_pca$var$coord

# Print the loadings matrix
print(loadings_matrix)

```

```{r}
#Table

# Eigenvalues
eigenvalues <- res_pca$eig[1:2, 1]

# Variance percent
variance_percent <- res_pca$eig[1:2, 2]

# Cumulative variance percent
cumulative_variance_percent <- res_pca$eig[1:2, 3]

# Create the table
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)

rownames(df) <- c("Dim1", "Dim2")

# Print the table
print(df)

```

```{r}
#Table

library(gt)

# Create data frame
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)
rownames(df) <- c("Dim1", "Dim2")

# Create the table using gt
gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = vars(Eigenvalue, Variance_Percent, Cumulative_Variance_Percent),
    decimals = 2
  )

```

calculating PCA based on AREA: TA/ TP1/ D1:

```{r}
#4lithic density

# A→ V=4×5×2.6 = 52 m3
# TP → V=2×1.8×1.6 = 5.76 m 3
# D1 → 1.5×1.5×0.95 = 2.1375 m3

library(dplyr)

# Define the volumes for each area based on the given information
volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375) # in m^3
)

# Count the number of artifacts by area
artifact_counts <- tl_final_area %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Join artifact_counts with volumes by area
lithic_density_df <- left_join(artifact_counts, volumes, by = "area")

# Compute lithic density for each area
lithic_density_df <- lithic_density_df %>%
  mutate(lithic_density = n / volume) # number of artifacts / volume

# Print the resulting table
print(lithic_density_df[, c("area", "lithic_density")])

```

```{r}

#2 Calculate the frequency of cores and core fragments by area

library(dplyr)
# Calculate the core frequency
core_frequency_by_area <- left_join(number_of_cores_by_area, 
                                    total_lithics_by_area, 
                                    by = "area") %>%
  mutate(core_frequency = (number_of_cores / total_lithics) * 100)

# Display the result
print(core_frequency_by_area)

```

```{r}
# 1number of Blank

# Calculate the total number of lithic artifacts
total_lithics <- tl_final_area %>%
  summarise(total_lithics = n()) %>%
  pull(total_lithics)

# Display the total number of lithic artifacts
print(paste("The total number of lithic artifacts is:", total_lithics))

```

```{r}
# 5Calculate the retouched frequency for each area
library(dplyr)

total_lithics_by_area <- tl_final_area %>%
  group_by(area) %>%
  summarise(total_lithics = n())

number_of_retouched_by_area <- tl_final_area %>%
  mutate(retouch = as.numeric(`retouch`)) %>% 
  filter(!is.na(retouch), retouch == 1) %>%  
  group_by(area) %>%  
  summarise(number_of_retouched = sum(retouch))

retouched_frequency_by_area <- left_join(number_of_retouched_by_area, total_lithics_by_area, by = "area") %>%
  mutate(retouched_frequency = (number_of_retouched / total_lithics) * 100)

# Display the retouched frequency by area
print(retouched_frequency_by_area)

print(number_of_retouched_by_area)

```

```{r}
# 3Calculating Tool Diversity using Menhinick's Index
tool_div <- tl_final_area %>%
  filter(area %in% c("ta", "d1", "tp1"), 
         !is.na(tooltypegroup)) %>%
  group_by(area) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(sum(!is.na(tooltypegroup))), .groups = 'drop')
```

feature frequency: d1 1 ta 1 tp1 0)

estimated volumes (area) A→ 52 m3 TP → 5.76 m 3 D1 → 2.1375 m3

```{r}
library(dplyr)
library(FactoMineR)
library(tibble)

# Define estimated area and feature frequency data
area_data <- tibble(
  area = c("d1", "ta", "tp1"),
  estimated_area = c(2.1375, 52, 5.76),
  feature_freq = c(1, 1, 0)
)

# Step 1: Number of artifacts by area
artifact_counts <- tl_final_area %>%
  group_by(area) %>%
  tally()

# Step 2: Calculate each variable

# Retouch Frequency 
retouch_freq <- tl_final_area %>%
  group_by(area) %>%
  summarise(retouchFreq = mean(!is.na(retouch)), .groups = 'drop')

# Blank Frequency
blank_freq <- tl_final_area %>%
  group_by(area) %>%
  summarise(blankFreq = mean(!is.na(blank)), .groups = 'drop')

# Core Frequency
core_freq <- tl_final_area %>%
  group_by(area) %>%
  summarise(coreFreq = mean(!is.na(`core-typology`)),.groups = 'drop')

# Lithic density (Artifact counts / estimated area)
lithic_density <- artifact_counts %>%
  left_join(area_data, by = "area") %>%
  mutate(lithic_density = n / estimated_area)

# Tool Diversity using Menhinick's Index
tool_div <- tl_final_area %>%
  group_by(area) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(n()), .groups = 'drop')

# Step 3: Combining all variables

pca_final_table_area <- left_join(area_data, retouch_freq, by = "area") %>%
  left_join(blank_freq, by = "area") %>%
  left_join(core_freq, by = "area") %>%
  left_join(lithic_density, by = "area") %>%
  left_join(tool_div, by = "area") %>%
  select(area, estimated_area, feature_freq, retouchFreq, blankFreq, coreFreq, lithic_density, toolDiv)

```

```{r}
pca_final_table_area <- pca_final_table_area %>%
    select(
        area, 
        estimated_area = estimated_area.x, 
        feature_freq = feature_freq.x, 
        retouchFreq, 
        blankFreq, 
        coreFreq, 
        n, 
        lithic_density, 
        toolDiv
    )

```

```{r}
# Step 4: Perform PCA
res_pca_area <- PCA(pca_final_table_area %>% select(-area, -n), graph = FALSE)

# Show the summary of PCA
print(res_pca_area)

print(res_pca_area$eig)
print(res_pca_area$var$coord)
```

```{r}
# TABLE: Eigenvalues and percentage of variance for each dimension of PCA 
library(kableExtra)

# Define the values manually (you would get these from your PCA results)
eigenvalues <- c(4.411234, 2.588766)
prop_var_explained <- c(63.02, 36.98)
cumulative_var_explained <- c(63.02, 100.00)

# Create the data frame
results_df <- data.frame(
  Dimension = c("Dim. 1", "Dim. 2"),  # Only keep the dimensions you want to display
  Eigenvalue = eigenvalues,
  `Variance / percent` = prop_var_explained,
  `Cumulative / variance percent` = cumulative_var_explained
)

# Generate the table
kable(results_df,
      caption = "Table 3. Eigenvalues and percentage of variance for each dimension of PCA",
      align = 'c',
      col.names = c("Dimension", "Eigenvalue", "Variance (%)", "Cumulative Variance (%)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 11)

```

```{r}
#plotting Dim-1 and Dim-2
library(FactoMineR)
library(factoextra)

# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca_area, axes = 1, choice = "var")

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca_area, axes = 2, choice = "var")

```

```{r}
#|fig-cap: Contribution of variables for each of the tow relevant PCA dimensions

# having Dim1 and Dim2 plots next to each other

library(gridExtra)

# Create the first plot and assign it to p1
p1 <- fviz_contrib(res_pca_area, axes = 1, choice = "var")

# Create the second plot and assign it to p2
p2 <- fviz_contrib(res_pca_area, axes = 2, choice = "var")

# Use grid.arrange to place them next to each other
grid.arrange(p1, p2, ncol = 2)

```

```{r}

loadings <- res_pca_area$var$coord

# 
eigenvalues <- res_pca_area$eig

#  Dim.1
contrib_dim1 <- (loadings[,1]^2 / eigenvalues[1, 1]) * 100

#  Dim.2
contrib_dim2 <- (loadings[,2]^2 / eigenvalues[2, 1]) * 100

contrib_df <- data.frame(
  Variable = rownames(loadings),
  Contrib_Dim1 = contrib_dim1,
  Contrib_Dim2 = contrib_dim2
)

print(contrib_df)

```

```{r}
# screeplot

library(FactoMineR)
library(factoextra)

fviz_eig(res_pca_area)

```

```{r}
# Correlation plots of variables for each of the tow relevant PCA dimensions

fviz_pca_var(res_pca_area,

             repel = TRUE) + 
  theme(text = element_text(size = 10))

```

```{r}
# Eigenvalues and percentage of variance
eig.val <- res_pca_area$eig
print(eig.val)

# Variable coordinates (correlations with principal components)
var.coord <- res_pca_area$var$coord
print(var.coord)

```

```{r}
#PCA biplot for Dimensions 1 and 2

fviz_pca_biplot(res_pca_area, labelsize = 3, repel = TRUE)

```

```{r}
# Create the initial biplot without individual labels
p <- fviz_pca_biplot(res_pca_area, label = "none", 
                     repel = TRUE, labelsize = 4)

# Add the text labels for the areas using their coordinates from the PCA result
p + geom_text(aes(x = res_pca_area$ind$coord[, 1], 
                  y = res_pca_area$ind$coord[, 2], 
                  label = pca_final_table_area$area), 
             size = 3, nudge_x = 0.1, nudge_y = 0.1)

```

```{r}
individual_coordinates = res_pca_area$ind$coord
variable_coordinates = res_pca_area$var$coord

```

```{r}
# For variables
fviz_cos2(res_pca_area, choice = "var")

# For individuals
fviz_cos2(res_pca_area, choice = "ind")

```

DONE WITH PCA

```{r}

individual_coordinates <- res_pca_area$ind$coord

individual_coordinates_df <- as.data.frame(individual_coordinates)
individual_coordinates_df$area <- unique(area_data$area)

print(individual_coordinates_df)

```

```{r}
library(kableExtra)

# Extract individual coordinates
individual_coordinates <- res_pca_area$ind$coord

# Convert to a data frame for better visibility and add the 'area' column back for context
individual_coordinates_df <- as.data.frame(individual_coordinates)
individual_coordinates_df$Area <- unique(area_data$area)  # Make sure the order matches

# Use kable to create a beautiful table
kable(individual_coordinates_df, 
      col.names = c("Area", "Principal Component 1", "Principal Component 2"), 
      caption = "PCA Coordinates for the Archaeological Areas excavation",
      align = c('l', 'c', 'c'),
      digits = 3) %>%
  kable_styling("striped", full_width = F)

```

```{r}
# Get the loadings (coefficients) of original variables for each principal component
loadings_matrix <- res_pca$var$coord

```

```{r}
# Eigenvalues
eigenvalues <- res_pca$eig[1:2, 1]
variance_percent <- res_pca$eig[1:2, 2]
cumulative_variance_percent <- res_pca$eig[1:2, 3]

# Create the table
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)

rownames(df) <- c("Dim1", "Dim2")

# Print the table
print(df)

```

```{r}
library(gt)

gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

```

```{r}

artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Print the resulting table
print(artifact_counts)
```

```{r}

blank_freq <- tl_final_area %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(blank)) %>%
  group_by(area) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Calculating Core Frequency
core_freq <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(corefrag)) %>%
  group_by(area) %>%
  summarise(coreFreq = sum(corefrag %in% c("core fragment", "bullet", "flat", "pyramid", "cylinder/prismatic", "core rejuvenation piece", "shapeless", "core on flake", "bipolar")) / sum(!is.na(corefrag)), .groups = 'drop')

# Calculating Tool Diversity using Menhinick's Index
tool_div <- tl_final_area %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(tooltypegroup)) %>%
  group_by(area) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(sum(!is.na(tooltypegroup))), .groups = 'drop')

# Combining all the calculated frequencies into one tibble
final_table <- blank_freq %>%
  inner_join(core_freq, by = "area") %>%
  inner_join(tool_div, by = "area")

# Printing the final table
print(final_table)

```

```{r}
# Joining calculated frequencies
final_table <- blank_freq %>%
  inner_join(core_freq, by = "area") %>%
  inner_join(tool_div, by = "area") %>%
  inner_join(retouch_freq, by = "area")

# Converting area to row names
final_table_rownames <- final_table %>%
  column_to_rownames(var = "area")

```

```{r}

library(FactoMineR)
library(factoextra)

# Run PCA
res_pca <- PCA(final_table_rownames, graph = FALSE)

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca, axes = 2, choice = "var")

```

```{r}
# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca, axes = 1, choice = "var") 
```

```{r}
# Visualize eigenvalues and explained variance
fviz_screeplot(res_pca)
```

```{r}
fviz_pca_biplot(res_pca, labelsize = 3, repel = TRUE)

```

```{r}
# Visualize the variables on the principal components
fviz_pca_var(res_pca)
```

```{r}
# Get the loadings (coefficients) of original variables for each principal component
loadings_matrix <- res_pca$var$coord

# Print the loadings matrix
print(loadings_matrix)

```

```{r}
# Eigenvalues
eigenvalues <- res_pca$eig[1:2, 1]
variance_percent <- res_pca$eig[1:2, 2]
cumulative_variance_percent <- res_pca$eig[1:2, 3]

# Create the table
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)

rownames(df) <- c("Dim1", "Dim2")

# Print the table
print(df)

```

```{r}
library(gt)

# Create the table 
gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

```

```{r}
# Count the number of artifacts by area
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Print the resulting table
print(artifact_counts)

```

```{r}

volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375)
)

# Count the number of artifacts by area
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Join artifact_counts with volumes by area
lithic_density_df <- left_join(artifact_counts, volumes, by = "area")

# Compute lithic density for each area
lithic_density_df <- lithic_density_df %>%
  mutate(lithic_density = n / volume)

# Print the resulting table
print(lithic_density_df[, c("area", "lithic_density")])

```

```{r}
# Counting the number of cores by area
core_counts <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1") & !is.na(`core and core fragments`)) %>%
  group_by(area) %>%
  tally()

print(core_counts)

```

OK

```{r}
#| label: fig-1-layer
#| fig-cap: "Type of Removals"
#| fig-height: 5

#Type of blank based on Layers

library(ggbeeswarm)

tl4 %>% 
  filter(!is.na(layer) & Blank2 %in% c('bladelet', 
                                       'blade', 
                                       'flake')) %>%

  # Create bar plot
  ggplot() +
  aes(x = layer, 
      fill = Blank2) +
  geom_bar(position = "dodge") +
  labs(x = "Area", 
       y = "Number", 
       fill = "Blank Type") +
  theme_minimal()

```

```{r}
#Type of core based on Layers

library(ggbeeswarm)

tl4 %>% 
  filter(!is.na(layer) & 
           `core-typology` %in% c('bullet','prismatic','pyramid')) %>%

  # Create bar plot
  ggplot() +
  aes(x = layer, 
      fill = `core-typology`) +
  geom_bar(position = "dodge") +
  labs(x = "Area", 
       y = "Number", 
       fill = "Blank Type") +
  theme_minimal()
```

consistency in dimention:

```{r}
#|label: fig_2
#|fig-cap: "Boxplot for each Dimention"
#|fig-height: 5

library(ggplot2)
library(dplyr)
library(tidyr)

# Set the value for n
n <- 100

# Tally by area
tl_final_area_tally <- tl_final_area %>%
  group_by(area) %>%
  tally()

# Filter areas with more than n rows
tl_final_area_tally_greater_than_n <- tl_final_area_tally %>%
  filter(n >= n)

# Filter data based on areas with more than n rows
tl_final_area_filtered <- tl_final_area %>%
  filter(area %in% tl_final_area_tally_greater_than_n$area) %>%
  filter(!is.na(blank) & blank != "")

# Select and mutate dimensions
tl_final_area_selected <- tl_final_area_filtered %>%
  select(area, width, thickness, length) %>%
  mutate(
    width = as.numeric(width),
    thickness = as.numeric(thickness),
    length = as.numeric(length)
  )

# Pivot data to longer format
tl_final_area_long <- tl_final_area_selected %>%
  pivot_longer(cols = c(width, thickness, length), 
               names_to = "Dimension", 
               values_to = "Value")

# Create the box plot
ggplot(tl_final_area_long, aes(x = area, y = Value)) +
  geom_boxplot() +
  facet_wrap(~Dimension, scales = "free") +
  labs(x = "Area", y = "Value") +
  theme_minimal()

```

```{r}
#|label: fig_3
#|fig-cap: Dimention By Area
#|ffig-height: 5

library(ggplot2)
library(dplyr)
library(tidyr)
library(ggbeeswarm)

# Create boxplot
ggplot(tl_final_area_long, 
       aes(x = area, y = Value)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.01) +
  facet_wrap(~Dimension, scales = "free_y") +
  scale_y_log10() +
  theme_minimal()

```

```{r}

```

```{r}
#|label: fig_2-layer
#|fig-cap: "Boxplot for each Dimention"
#|fig-height: 5

n <- 100

# Tally count of each layer
tl_final_layer_tally <- tl_final_area %>% 
  group_by(layer) %>% 
  tally()

# Filter layers with more than n rows
tl_final_layer_tally_greater_than_n <- tl_final_layer_tally %>% 
  filter(n >= n)

# Filter data for layers with more than n rows
tl_final_layer_filtered <- tl_final_area %>% 
  filter(layer %in% tl_final_layer_tally_greater_than_n$layer)

# Select and mutate
tl_final_layer_selected <- tl_final_layer_filtered %>%
  select(layer, 
         `width`, 
         `thickness`, 
         `length`) %>%
  mutate(across(c
                (
           `width`, 
           `thickness`, 
           `length`), 
           as.numeric)
         )

# Pivot longer
tl_final_layer_long <- tl_final_layer_selected %>%
  pivot_longer(cols = 
                 c(width, 
                   thickness, 
                   length), 
               names_to = 
                 "Dimension", 
               values_to = 
                 "Value") %>%
  drop_na(Value)

# Create box plot
ggplot(tl_final_layer_long, 
       aes(layer, 
           Value)) +

  geom_boxplot() +
  facet_wrap(~Dimension, 
             scales = "free") +
  labs(x = "Layer", 
       y = "Value") +

   theme_minimal()

```

```{r}
#|label: fig_3-layer
#|fig-cap: Dimension By Layer
#|fig-height: 5

# Create boxplot
ggplot(tl_final_layer_long, 
       aes(x = layer, y = Value)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.01) +
  facet_wrap(~Dimension, scales = "free_y") +
  scale_y_log10() +
  theme_minimal()

```

```{r}
#|label: fig_4-width
#|fig-cap: "Width of Artifacts (Area A)"
#| fig-height: 5
tl_final_area %>%
 filter(area == "ta") %>%
  select(area, 
         width) %>%
  ggplot() +
  aes(x = width) +
  geom_histogram(binwidth = 2) +
  labs(x = "Width of Debitage (mm)", 
       y = "Frequency" )+
  theme_minimal()
```

```{r}
#|label: fig_5-width
#|fig-cap: "Width of Artifacts (Area A, D1, TP1)"
#|fig-height: 5

tl_final_area %>%
  filter(area %in% 
           c("ta", "d1", "tp1")) %>%
  mutate(width = as.numeric(width)) %>%  # Explicitly convert to numeric

   select(area, 
          `width`) %>%
  ggplot() +

  aes(x = width) +  # Make sure the aesthetic mapping is correct

  geom_histogram(binwidth = 2) +

  labs(x = "Width of Debitage (mm)", 
       y = "Frequency" ) +

  facet_wrap(~area, 
             ncol = 1) +

  theme_minimal()

```

```{r}
#|label: fig_5-width-layer
#|fig-cap: "Width of Artifacts (Area A, D1, TP1)"
#|fig-height: 5

tl_final_area %>%
  filter(layer %in% c("a", 
                      "c", 
                      "e or d", 
                      "e", 
                      "d")) %>%
  mutate(width = as.numeric(width)) %>%  # Convert to numeric
  select(layer, width) %>%
  ggplot() +
  aes(x = width) +
  geom_histogram(binwidth = 2) +
  labs(x = "Width of Debitage (mm)", y = "Frequency") +
  facet_wrap(~layer, ncol = 1) +
  theme_minimal()

```

```{r}

```

```{r}
#|label: fig_6-thickness-layer
#|fig-cap: "Thickness of Artifacts (Layers A, C, E or D)"
#|fig-height: 5

tl_final_area %>%
 filter(layer %in% c("a", 
                     "c", 
                     "e or d", 
                     "d", 
                     "e")) %>%
 mutate(thickness = as.numeric(thickness)) %>% 
  select(layer, 
         thickness) %>%
  ggplot() +
  aes(x = thickness) +
  geom_histogram(binwidth = 2) +
  labs(x = "Thickness of Debitage (mm)", 
       y = "Frequency" )+
  facet_wrap(~layer , ncol = 1) +
  theme_minimal()

```

```{r}
#|label: fig_7
#|fig-cap: "Length of Artifacts (Area A, D1, TP1)"
#|fig-height: 5

tl3 %>%
  filter(Blank_part == "complete", area %in% c("ta", "tp1", "d1")) %>%
  mutate(`Length (mm)` = as.numeric(`Length (mm)`)) %>%
  select(area, 
         `Length (mm)`) %>%
  ggplot() +
  aes(x = `Length (mm)`) +
  geom_histogram(binwidth = 2) +
  labs(x = "Length of Debitage (mm)", 
       y = "Frequency") +
  facet_wrap(~area, ncol = 1) +
  theme_minimal()

```

```{r}
#| label: fig-7-length-layer
#| fig-cap: Length of Artefact based on layers
#| fig-height: 5
tl3 %>% 
  filter(layer %in% c("layer a" , "layer d" , "layer e" , "layer c" , "layer e or d")) %>% 
  mutate(`Length (mm)` = as.numeric(`Length (mm)`)) %>% 
  select(layer, `Length (mm)`) %>% 
  ggplot(aes(x = `Length (mm)`)) + 
  geom_histogram(binwidth = 2) +
  labs(x = "Length of Debitage (mm)" ,
       y = "Frequency") + 
  facet_wrap(~layer, ncol = 1) + 
  theme_minimal()

```

```{r}
#|label: NA 
#|fig-cap: "tool typology"
#|fig-height: 5

```

```{r}
library(ggplot2)

# Calculate the counts for debitage and 
counts <- tl4_tool_Blank %>%
  group_by(debitage, ) %>%
  summarise(count = n())

# Create the bar plot
ggplot(counts, 
       aes(x = factor(debitage), 
           y = count, 
           fill = factor(debitage))) +

  geom_bar(stat = "identity", 
           position = "dodge") +

  # Move the labels slightly above the bars
  geom_text(aes(label = count), 
            vjust = -0.5) +

  geom_hline(data = counts %>% 
               filter( == 1), 
             aes(yintercept = count),
             linetype = "dashed", 
             color = "red") +
  labs(title = "Debitage vs. ",
       x = "",
       y = "Numbers") +
  theme_minimal()

```

```{r}
# Calculate the counts for debitage and 
counts <- tl4_tool_Blank %>%
  group_by(Retouch = 1,  =1) %>%
  summarise(count = n())

# Create the bar plot
ggplot(counts, 
       aes(x = factor( = 1), 
           y = count, 
           fill = factor( = 1))) +

  geom_bar(stat = "identity", 
           position = "dodge") +

  # Move the labels slightly above the bars
  geom_text(aes(label = count), 
            vjust = -0.5) +

  geom_hline(data = counts %>% 
               filter( == 1), 
             aes(yintercept = count),
             linetype = "dashed", 
             color = "red") +
  labs(title = " vs. formal ",
       x = "",
       y = "Numbers") +
  theme_minimal()
```

```{r}
#|label: fig_
#|fig-cap: ""
#| fig-height: 5
tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", "blade", "flake")) %>%
  select(Blank2, shine2 = 1) %>%
  ggplot() +
  aes(x = shine2) +
  geom_bar() +
  labs(x = "shine", 
       y = "Frequency" )+
  theme_minimal()

tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", "blade", "flake")) %>%
  select(Blank2, Retouch = 1) %>%
  ggplot() +
  aes(x = Retouch) +
  geom_bar() +
  labs(x = "Retouch", 
       y = "Frequency" )+
  theme_minimal()

tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", "blade", "flake")) %>%
  select(Blank2, Utilization = 1) %>%
  ggplot() +
  aes(x = Utilization) +
  geom_bar() +
  labs(x = "Utilization", 
       y = "Frequency" )+
  theme_minimal()

```

```{r}
library(ggplot2)
library(patchwork)

plot_shine <- tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", 
                       "blade", 
                       "flake")) %>%
  select(Blank2, shine2 = 1) %>%
  ggplot() +
  aes(x = shine2) +
  geom_bar() +
  labs(x = "shine", y = "Frequency") +
  theme_minimal()

plot_retouch <- tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", 
                       "blade", 
                       "flake")) %>%
  select(Blank2, 
         Retouch = 1) %>%
  ggplot() +
  aes(x = Retouch) +
  geom_bar() +
  labs(x = "Retouch", 
       y = "Frequency") +
  theme_minimal()

plot_utilization <- tl4_tool_Blank %>%
  filter(Blank2 %in% c("bladelet", "blade", "flake")) %>%
  select(Blank2, Utilization = 1) %>%
  ggplot() +
  aes(x = Utilization) +
  geom_bar() +
  labs(x = "Utilization", y = "Frequency") +
  theme_minimal()

# Combine the plots using patchwork
plot_combined <- plot_shine + plot_retouch + plot_utilization
plot_combined

```