### 2023-09-12 paper/paper_tulai.qmd
edited most parts-made main figures for our purpose

---
title: ""
author: ""
date: now
format: 
  html:
    self-contained: true
execute: 
  warning: false
  fig-dpi: 600
title-block-published: "last updated"
editor: visual
---

```{r}
# load libraries
library(tidyverse)
library(here)
library(ggpmisc)
library(ggrepel)
library(ggbeeswarm)
```

Cleaning Data:

```{r}
# Read the data
tl <- read_csv(here("data/Tulai_Lithic_Assamblage.csv"))

# Create a new data frame with renamed columns
tl1 <- tl
names(tl1)[1:2] <- c("lithic_id", 
                     "TA")

# modify TA column
tl1 <- tl1 %>%
  # Remove everything within brackets
  mutate(TA = str_remove_all(TA, 
                             "\\(.*\\)")) %>%

  # Add slash between TA and numbers
  mutate(TA = str_replace_all(TA, 
                              "TA([0-9]+)", 
                              "TA/\\1")) %>% 
  # Add slash between numbers
  mutate(TA = str_replace_all(TA, 
                              "([0-9]+)/([0-9]+)", "\\1/\\2")) %>% 
 # fill NA values in TA column
  tidyr::fill(TA, .direction = "down") %>%
  # Prepend 'TA/' to numbers
  mutate(TA = if_else(str_detect(TA, 
                                 "^[0-9]+$"), 
                      paste0("TA/",
                             TA), 
                      TA)) %>% 
  # Remove all non-alphanumeric characters
  mutate(TA = str_replace_all(TA, 
                              "[^[:alnum:] /]", 
                              "")) %>% 
  # Make all text lowercase
  mutate(TA = tolower(TA)) %>% 
  # Remove white space
  mutate(TA = str_replace_all(TA, " ", ""))  %>%
  # Separate 'TA' column into 'area' and 'depth' columns
  separate(TA, into = c("area", 
                        "depth"), 
           sep = "/",
           remove = FALSE,
           convert = TRUE) %>%
  # Convert 'depth' to numeric
  mutate(depth = as.numeric(depth)) %>% 
  # Create depth ranges
  mutate(depth = case_when(
             depth == 3 ~ "0-30 cm",
             depth == 34 ~ "30-40 cm",
             depth == 45 ~ "40-50 cm",
             depth == 56 ~ "50-60 cm",
             depth == 67 ~ "60-70 cm",
             depth == 72 ~ "70-120 cm",
             depth == 125 ~ "120-150 cm",
             depth == 158 ~ "150-180 cm",
             depth == 189 ~ "180-190 cm",
             depth == 190 ~ "190-200 cm",
             depth == 201 ~ "200-210 cm",
             depth == 212 ~ "210-220 cm",
             depth == 223 ~ "220-230 cm",
             depth == 234 ~ "230-240 cm",
             depth == 235 ~ "230-250 cm",
             depth == 256 ~ "250-260 cm",
             depth == 267 ~ "260-270 cm",
             depth == 72 ~ "70-120 cm",
             depth == 1 ~ "0-10 cm",
             depth == 2 ~ "0-20 cm", 
             depth == 13 ~ "10-30 cm",
             depth == 19 ~ "10-90 cm",
             depth == 23 ~ "20-30 cm",
             depth == 34 ~ "30-40 cm",
             depth == 45 ~ "40-50 cm",
             depth == 51 ~ "50-100 cm",
             depth == 56 ~ "50-60 cm",
             depth == 57 ~ "50-70 cm",
             depth == 89 ~ "80-90 cm",
             depth == 91 ~ "90-100 cm",
             depth == 101 ~ "100-110 cm",
             depth == 112 ~ "110-120 cm",
             depth == 123 ~ "120-130 cm",
             depth == 124 ~ "120-140 cm",
             depth == 23 ~ "20-30 cm",
             depth == 46 ~ "40-60 cm",
             depth == 78 ~ "70-80 cm",
             depth == 235 ~ "230-250 cm",
             depth == 12 ~ "10-20 cm",
             depth == 90 ~ "90-100 cm",

             TRUE ~ as.character(depth)
  )) %>%

  separate(depth, 
           into = c("upper",
                    "lower"),
           sep = "-",
           remove = FALSE,
           convert = TRUE) %>% 
  mutate(lower = parse_number(lower)) %>% 
  rowwise() %>% 
  mutate(midpoint = mean(c(upper, lower))) %>%
# create 'layer' column
  mutate(layer = case_when(
      area == "tp1" & depth == "0-10 cm" ~ "c",
      area == "tp1" & depth == "0-20 cm" ~ "c",
      area == "tp1" & depth == "10-30 cm" ~ "c",
      area == "tp1" & depth == "10-90 cm" ~ "c",
      area == "tp1" & depth == "20-30 cm" ~ "c",
      area == "tp1" & depth == "30-40 cm" ~ "c",
      area == "tp1" & depth == "40-50 cm" ~ "c",
      area == "tp1" & depth == "50-100 cm" ~ "c",
      area == "tp1" & depth == "50-60 cm" ~ "c",
      area == "tp1" & depth == "50-70 cm" ~ "c",
      area == "tp1" & depth == "80-90 cm" ~ "c",
      area == "tp1" & depth == "90-100 cm" ~ "c",
      area == "tp1" & depth == "100-110 cm" ~ "c",
      area == "tp1" & depth == "110-120 cm" ~ "c",
      area == "tp1" & depth == "120-130 cm" ~ "c",
      area == "tp1" & depth == "120-140 cm" ~ "c",
      area == "d1" & depth == "0-20 cm" ~ "c",
      area == "d1" & depth == "20-30 cm" ~ "c",
      area == "d1" & depth == "40-60 cm" ~ "a",
      area == "d1" & depth == "50-60 cm" ~ "a",
      area == "d1" & depth == "60-70 cm" ~ "a",
      area == "d1" & depth == "70-80 cm" ~ "a",
      area == "ta" & depth == "0-30 cm" ~ "e",
      area == "ta" & depth == "30-40 cm" ~ "e",
      area == "ta" & depth == "40-50 cm" ~ "e",
      area == "ta" & depth == "50-60 cm" ~ "e",
      area == "ta" & depth == "60-70 cm" ~ "e",
      area == "ta" & depth == "70-120 cm" ~ "d",
      area == "ta" & depth == "120-150 cm" ~ "d",
      area == "ta" & depth == "150-180 cm" ~ "d",
      area == "ta" & depth == "180-190 cm" ~ "d",
      area == "ta" & depth == "190-200 cm" ~ "d",
      area == "ta" & depth == "200-210 cm" ~ "d",
      area == "ta" & depth == "210-220 cm" ~ "d",
      area == "ta" & depth == "220-230 cm" ~ "d",
      area == "ta" & depth == "230-240 cm" ~ "d",
      area == "ta" & depth == "230-250 cm" ~ "d",
      area == "ta" & depth == "250-260 cm" ~ "d",
      area == "ta" & depth == "260-270 cm" ~ "d",

      TRUE ~ NA_character_
  ))

# Remove specific columns
tl1 %>% 
  select(-c("AREA", 
            "DEPTH (cm)", 
            "Raw Material", 
            "Colour/Grain/Opacity", 
            "Pattern"))
```

```{r}
# exploring Data

tl1 %>% 
  group_by(midpoint) %>% 
  tally() %>% 
  drop_na(midpoint) %>% 
ggplot() + 
  aes(midpoint, n) +
  geom_col()
```

```{r}
# exploring Data

tl1 %>% 
  group_by(layer) %>% 
  tally() %>% 
  drop_na(layer) %>% 
ggplot() + 
  aes(layer, n) +
  geom_col()
```

```{r}
# exploring Data

tl2 <- tl1 %>%
  mutate(
    Blank = case_when(
      Breakage %in% c("0", "1", NA_character_) ~ Blank,

      TRUE ~ Breakage  
    )
  )

```

```{r}
# Cleaning Blank column

tl2 <- tl2 %>%
  mutate(
    Blank = str_to_lower(Blank),
    Blank = str_trim(Blank),
    Blank = str_replace_all(Blank, "[./]", "-"),
    Blank = str_replace_all(Blank, "\\s*-\\s*", "-")
  ) %>%
  mutate(Blank = case_when(
             Blank == "peo" ~                      "bladelet-pro",
             Blank == "pro" ~                      "bladelet-pro",
             Blank == "bladelet-dis" ~             "bladelet-dis",
             Blank == "microblade" ~               "bladelet-complete",
             Blank == "microblade-pro" ~           "bladelet-pro",
             Blank == "microblade-med" ~           "bladelet-med",
             Blank == "micrpblade-dis" ~           "bladelet-dis",
             Blank == "dis" ~                      "bladelet-dis",
             Blank == "bladelrt-med" ~             "bladelet-med",
             Blank == "bladelet-mes" ~             "bladelet-med",
             Blank == "bladelet-,ed" ~             "bladelet-med",
             Blank == "bladelet-bladelet-med" ~    "bladelet-med",
             Blank == "medial" ~                   "bladelet-med",
             Blank == "bladelet-bladelet-pro" ~    "bladelet-pro",
             Blank == "bladelert-pro" ~            "bladelet-pro",
             Blank == "microblade-medial" ~        "bladelet-med",
             Blank == "nicroblade-pro" ~           "bladelet-pro",
             Blank == "microblade-?" ~             "bladelet-complete",
             Blank == "med" ~                      "bladelet-med",
             Blank == "bladeler-pro" ~             "bladelet-pro",
             Blank == "microblade-dis" ~           "bladelet-dis",
             Blank == "microblde-dis" ~            "bladelet-dis",
             Blank == "bladlet-pro" ~              "bladelet-pro",
             Blank == "indistinct" ~               "bladelet-med",
             Blank == "flake" ~                    "flake-complete",
             Blank == "dladelet-pro" ~             "bladelet-pro",
             Blank == "dladelet-pro" ~             "bladelet-pro",
             Blank == "bldelet-bladelet-pro" ~     "bladelet-pro",
             Blank == "bldelet-pro" ~              "bladelet-pro",
             Blank == "nicroblade-pro" ~           "bladelet-pro",
             Blank == "microblade-?" ~             "bladelet-med",
             Blank == "bladelt-pro" ~              "bladelet-pro",
             Blank == "bladeket-med" ~             "bladelet-med",
             Blank == "nicroblade-pro" ~           "bladelet-pro",
             Blank == "blaelet" ~                  "bladelet-complete",
             Blank == "bladele-med" ~              "bladelet-med",
             Blank == "bladelet-nearly complete" ~ "bladelet-complete",
             Blank == "blade-nearly complete" ~    "blade-complete",
             Blank == "bladelet-?" ~               "bladelet-complete",
             Blank == "flake?" ~                   "flake-complete",
             Blank == "micrpblade-dis" ~           "bladelet-dis",
             Blank == "blaelet-med" ~              "bladelet-med",
             Blank == "blaedlet-pro" ~             "bladelet-pro",  
             Blank == "bladelt-pro" ~              "bladelet-pro",   
             Blank == "bladelet-nearly complete"~  "bladelet-complete", 
             Blank == "microblade-med" ~           "bladelet-med", 
             Blank == "bldelet-bladelet-pro" ~     "bladelet-pro",
             Blank == "bldelet-flake" ~            "blade-complete",
             Blank == "mwdial" ~                   "bladelet-med",
             Blank == "blade" ~                    "blade-complete",
             Blank == "bladelt" ~                  "bladelet-complete",
             Blank == "bladlet" ~                  "bladelet-complete",
             Blank == "bladlet-med" ~              "bladelet-med",
             Blank == "bladelt-med" ~              "bladelet-med",
             Blank == "flaje" ~                    "flake",
             Blank == "flke" ~                     "flake",
             Blank == "bladelet" ~                 "bladelet-complete",
             Blank == "NA" ~                       "bladelet-",
             Blank == "thick flake" ~              "flake",
             Blank == "flke-mid" ~                 "flake-med",
             Blank == "flake-blade" ~              "flake",
             Blank == "angular flake?" ~           "flake",
             Blank == "flake-prox" ~               "flake-pro",
             Blank == "fklae-dis" ~                "flake-dis",
             Blank == "bladelete" ~                "bladelet",
             Blank == "NA" ~                    NA_character_,
    TRUE ~ Blank
  )) %>%
  mutate(Blank = na_if(Blank, ""))

rev(sort(table(tl2$Blank)))

# BM: still many typos and unusual categories in here, why is that? Can they be fixed?

tl3 <- tl2 %>%
  separate(Blank, 
           into = c("Blank2", "Blank_part"),
           sep = "-",
           remove = FALSE,
           convert = TRUE)

na_rows <- which(is.na(tl2$Blank))
print(na_rows)

zero_rows <- which(tl2$Blank == "0")
print(zero_rows)

one_rows <- which(tl2$Blank == "1")
print(one_rows)

rev(sort(table(tl3$Blank2))) # BM: still a lot of typos on the categories here, can you fix them?

```

```{r}
# BM: this is an unusual coding pattern, I'm curious about where you got it from?

tl3 <- tl3 %>%
  mutate(
    Blank2 = case_when(
      Blank2 == "flake" ~ "flake",  # Keep "flake" as it is
      (`Core Typology` %in% c("0", "NA", "na", "-", ".") | 
       is.na(`Core Typology`) | 
       `Core Typology` == "") & `Width (mm)` < 12 ~ "bladelet",
      (`Core Typology` %in% c("0", "NA", "na", "-", ".") | 
       is.na(`Core Typology`) | 
       `Core Typology` == "") & `Width (mm)` >= 12 ~ "blade",
      TRUE ~ Blank2
    )
  )  %>%
  mutate(
    Blank2 = case_when(
      Blank2 %in% c("blade", "bladelet", "flake") ~ Blank2,
      TRUE ~ NA_character_
    )
  )

```

```{r}
# Remove rows based on row names
# tl3 <- tl3[!(rownames(tl3) %in% c("1030", "644", "543")), ]

# BM: why are we removing these rows? A comment here would be good to explain. A simpler way, by referencing the row numbers directly:
# SS: they are natural rocks, and don't want to count them as lithic

tl3 <- tl3[-c(1030, 644, 543), ]
```

```{r}
tl4 <- tl3 %>%
  mutate(
    Utilization = if_else(Retouch == "1", 
                          NA_character_, 
                          as.character(Utilization))
  )

```

```{r}

tl4 <- tl4 %>%
  # BM: the problem with this combination of mutate_all and ifelse that you use below is that they 
  # coerce all columns to be the same data type, so you'll see that all the numeric
  # columns are now character columns after you run these lines. So you need to
  # coerce those number columns back to numeric type so you can plot and compute
  # with them
  mutate_all(~ ifelse(is.na(.x) | .x == "", NA, .x)) %>% 
  # BM: what does this next line do?
  #SS: I wanted to remove any asterisks, whitespace, or *, ?, or !. is that wrong?

  mutate_all(~ str_replace(., "(?<=\\D)\\*(?=\\D)|\\s+|[*?!]+", "")) %>%
  # BM: what is this SHINE variable? I'm not familiar with it

  ##SS: we meant sickle shine, this is one of the criteria based on which Hole claimed the site is not related to agriculture (similar to ALi Kosh located in Dehluran)
  mutate(
    SHINE = str_trim(SHINE) %>% 
            str_to_title(),
    shine2 = SHINE
  ) %>%
  mutate(
    shine2 = case_when(
      SHINE %in% c("Subparallel", 
                   "Parallel", 
                   "Sub-Parallel", 
                   "Parallel-Subprarallel",
                   "Scaled", 
                   "Sub-Paralel", 
                   "Sub-Parallell", 
                   "Sub-Paallel",
                   "Sub-Paralell", 
                   "Sub-Prallel", 
                   "L:semiparallel/R:scaled",
                   "Semiparallel", 
                   "Semi-Parallel", 
                   "Scaled/Sub-Parallel",
                   "Seb-Parallel", 
                   "Semi-Abrupt") ~ NA_character_,
      TRUE ~ SHINE
    ),
    Utilization = if_else(Utilization == "0", NA_character_, Utilization),
    Retouch =     if_else(Retouch == "0", NA_character_, Retouch),
    SHINE =       if_else(SHINE == "0", NA_character_, SHINE),
    shine2 =      if_else(shine2 == "0", NA_character_, shine2),
    Blank2 =      if_else(Blank2 == "0", NA_character_, Blank2)
  ) %>% 
  mutate(
    SHINE = na_if(SHINE, "")
  )

# BM: I think this is a better way to inspect:
rev(sort(table(tl4$SHINE)))
rev(sort(table(tl4$shine2)))

```

```{r}

# Convert 'NA' and 'na' strings to actual NA
tl4$`Core Typology` <- na_if(tl4$`Core Typology`, "NA")
tl4$`Core Typology` <- na_if(tl4$`Core Typology`, "na")

tl4 <- tl4 %>%
  mutate(
    `Core Typology` = str_to_lower(`Core Typology`),
    `Core Typology` = str_trim(`Core Typology`),
    # BM: what does this regex do in the next line?
    `Core Typology` = str_replace_all(`Core Typology`, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(`Core Typology` = case_when(
    `Core Typology` == "flatcore" ~ "flat-pressure",
             `Core Typology` == "rejuvention" ~ "rejuvenation piece-NA",
             `Core Typology` == "pressureprymidal core" ~ "pyramid-pressure",
             `Core Typology` == "pressureflat core" ~ "flat-pressure",
             `Core Typology` == "pressurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prussurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prismaticcore" ~ "prismatic-percussion",
             `Core Typology` == "pyramidalpressure core" ~ "pyramid-pressure",
             `Core Typology` == "flatcore with one debitage surface/pressure" ~ "flat-pressure",
             `Core Typology` == "corefragment" ~ "core fragment-NA",
             `Core Typology` == "cilandrical/pressure" ~ "cylinder-pressure",
             `Core Typology` == "bullet/pressure" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/pressure" ~ "pyramid-pressure",
             `Core Typology` == "percussion/pyramidal" ~ "pyramid-pressure",
             `Core Typology` == "flat/pressure" ~ "flat-pressure",
             `Core Typology` == "pyramidal/ percussion??" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ peressure" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ pressure" ~ "pyramid-pressure",
             `Core Typology` == "pressure" ~ "pyramid-pressure",
             `Core Typology` == "multidirectional/percussion" ~ "shapeless-percussion",
             `Core Typology` == "prismatic/ percussion" ~ "pyramid-pressure",
             `Core Typology` == "flatcore/ unidirectional/pressure" ~ "flat-pressure",
             `Core Typology` == "heavilyused" ~ "NA",
             `Core Typology` == "burnt" ~ "NA",
             `Core Typology` == "patinated" ~ "NA",
             `Core Typology` == "pyramidal(bullet)/pressure" ~ "bullet-pressure",
             `Core Typology` == "cylandrycal/ bidirectional pressure core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure bullet core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "bladelet" ~ "bullet-pressure",
             `Core Typology` == "lip" ~ "NA",
             `Core Typology` == "bladlet" ~ "bullet-pressure",
             `Core Typology` == "bladelet(30.56)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(28.55.6)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(16.28.1)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(40 5.7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(23.2 2.3)" ~ "bullet-pressure",
             `Core Typology` == "bladelet-flakeblade(41 16)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(3512.6/7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(27.307.5)" ~ "bullet-pressure",
    `Core Typology` == "pressure/pyramidal" ~ "pyramid-pressure",
    `Core Typology` == "bladeletcore" ~ "bullet-pressure",
    `Core Typology` == "pyramidal? / pressure" ~ "pyramid-pressure",
    `Core Typology` == "prismatic" ~ "pyramid-pressure",

    `Core Typology` == "bullet" ~ "bullet-pressure",

    TRUE ~ `Core Typology`
  )) %>%
# Separate 'Core Typology' into 'core-typology' and 'core-technology'
  separate(`Core Typology`, 
           into = c("core-typology", "core-technology"),
           sep = "-",
           remove = FALSE,
           convert = TRUE)

# Convert 'NA' and 'na' in new columns to actual NA
tl4$`core-typology` <- na_if(tl4$`core-typology`, "na")
tl4$`core-technology` <- na_if(tl4$`core-technology`, "na")

# Check the unique values again
# BM: still many typos, are they important?
rev(sort(table((tl4$`core-typology`))))
rev(sort(table((tl4$`core-technology`))))

```

```{r}

tl4 <- tl4 %>%
  mutate(
    Typology = str_to_lower(Typology),
    Typology = str_trim(Typology),
    Typology = str_replace_all(Typology, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(
    # BM: wow, seems like almost each artefact is its own type! 
    # BM: That is not ideal for statistical analysis
    # BM: This looks like it was very time-consuming to simplify. 
    Typology = case_when(
      `Typology` == "notch, truncated" ~ "tool/notch;truncated piece",
      `Typology` == "retouchpiece" ~ "tool/retouched piece",
      `Typology` == "endscraper" ~ "tool/end scraper",
      `Typology` == "notch" ~ "tool/notch",
      `Typology` == "sidescraper" ~ "tool/side scraper",
      `Typology` == "doubleside scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper (?)" ~ "tool/convergent scraper",
      `Typology` == "doublenotch on retouched piece" ~ "tool/double notch",
      `Typology` == "retouchpieces" ~ "tool/retouched piece",
      `Typology` == "burin(?)/ notch/scraper" ~ "tool/burin-notch-scraper",
      `Typology` == "borer" ~ "tool/perforator-borer",
      `Typology` == "retouchedpiece" ~ "tool/retouched piece",
      `Typology` == "notch/side scraper" ~ "tool/notch-side scraper",
      `Typology` == "corefragment" ~ "core/core fragment",
      `Typology` == "backed" ~ "tool/backed knife",
      `Typology` == "doubleside scraper ?" ~ "tool/double side scraper",
      `Typology` == "inversedenticulate" ~ "tool/denticulate",
      `Typology` == "retouchpieces/ side scraper?" ~ "tool/side scraper",
      `Typology` == "sidescraper ?" ~ "tool/side scraper",
      `Typology` == "notch-denticulate" ~ "tool/notch-denticulate",
      `Typology` == "borer/drill" ~ "tool/perforator-drill",
      `Typology` == "retouchedpiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchepiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchedpiece (double side scraper)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece (double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchepiece" ~ "tool/retouched piece",
      `Typology` == "inversenotch" ~ "tool/notch",
      `Typology` == "retouchepiece (double side scaraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece( double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "notchon retouche piece" ~ "tool/notch",
      `Typology` == "retouchedpiece (souble side scraper)" ~ "tool/double side scraper",
      `Typology` == "point" ~ "tool/perforator-point",
      `Typology` == "chunk" ~ "NA/NA",
      `Typology` == "coretablet" ~ "core/core tablet",
      `Typology` == "core" ~ "core/shapeless",
      `Typology` == "denticulate" ~ "tool/denticulate",
      `Typology` == "doublenotch (haft)" ~ "tool/double notch",
      `Typology` == "retouchedpiece (fine retouch)" ~ "tool/retouched piece",
      `Typology` == "endscraper on retouched piece" ~ "tool/end scraper",
      `Typology` == "retouchedpiece (serrated scraper)" ~ "tool/serrated scraper",
      `Typology` == "truncation" ~ "tool/truncated piece",
      `Typology` == "rejuventionpiece" ~ "core/rejuvenation piece",
      `Typology` == "notch/denticulate" ~ "tool/notch-denticulate",
      `Typology` == "retoucheduse on breakage" ~ "tool/retouched piece",
      `Typology` == "inversenotch/endscraper" ~ "tool/notch-end scraper",
      `Typology` == "alternatingside scraper" ~ "tool/side scraper",
      `Typology` == "alternatedouble side scraper" ~ "tool/double side scraper",
      `Typology` == "truncation/sidescaraper" ~ "tool/truncated piece",
      `Typology` == "doublenotch" ~ "tool/double notch",
      `Typology` == "borer/awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper?" ~ "tool/round scraper",
      `Typology` == "doublenotch scraper" ~ "tool/double notch scraper",
      `Typology` == "corerejuvention (core tablet?)" ~ "core/core tablet",
      `Typology` == "truncationاریبب" ~ "tool/truncated piece",
      `Typology` == "retouchedpiece (side scraper)" ~ "tool/side scraper",
      `Typology` == "retochedpice" ~ "tool/retouched piece",
      `Typology` == "truncation/notch" ~ "tool/truncated piece-notch",
      `Typology` == "doublealternate scraper" ~ "tool/double side scraper",
      `Typology` == "awl/ inverse notch" ~ "tool/notch",
      `Typology` == "retouchedpiece (alternate scraper?)" ~ "tool/double side scraper",
      `Typology` == "inverseside scraper" ~ "tool/side scraper",
      `Typology` == "usedcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "corerejuvention platform" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade??" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejivention" ~ "core/rejuvenation piece",
      `Typology` == "retouchedpice (side scraper?)" ~ "tool/side scraper" ,
      `Typology` == "scraper" ~ "tool/side scraper",
      `Typology` == "sickleblade" ~ "tool/sickle shine",
      `Typology` == "coretool (scraper)" ~ "tool/scraper-on core piece",
      `Typology` == "retouchedpiece (scraper)" ~ "tool/side scraper",
      `Typology` == "inversedenticulate?" ~ "tool/denticulate",
      `Typology` == "retouchedpiece (scraper?)" ~ "tool/side scraper",
      `Typology` == "coreside rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "truncation/endscraper" ~ "tool/endscraper on truncated piece",
      `Typology` == "retouchedpice(double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "doublescraper" ~ "tool/double side scraper",
      `Typology` == "dendiculate" ~ "tool/denticulate",
      `Typology` == "double-scraper" ~ "tool/double side scraper",
      `Typology` == "pyramidalcore" ~ "core/pyramid",
      `Typology` == "bulletcore" ~ "core/bullet",
      `Typology` == "corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "pyrmidalcore" ~ "core/pyramid",
      `Typology` == "corepreperation" ~ "core/core preparation piece",
      `Typology` == "crested" ~ "core/crested bladelet",
      `Typology` == "preperationblade" ~ "core/core preperation",
      `Typology` == "burin" ~ "tool/burin",
      `Typology` == "serrateddenticulate" ~ "tool/serrated scraper",
      `Typology` == "saw:serrated denticulate" ~ "tool/serrated scraper",
      `Typology` == "serratedused" ~ "tool/serrated scraper",
      `Typology` == "corepreeration" ~ "tool/core preparation piece",
      `Typology` == "coreprepearation" ~ "tool/core preparation piece",
      `Typology` == "awl" ~ "tool/perforator-awl",
      `Typology` == "notch.inverse" ~ "tool/notch",
      `Typology` == "point.notch" ~ "tool/perforator-point;notch",
      `Typology` == "inversescraper" ~ "tool/side scraper",
      `Typology` == "roundendscraper" ~ "tool/round scraper",
      `Typology` == "alternate.scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper/disc scraper" ~ "tool/scraper-convergent scraper",
      `Typology` == "serratedside scraper" ~ "tool/serrated scraper",
      `Typology` == "debri" ~ "NA/NA",
      `Typology` == "distalpart of a drill" ~ "tool/perforator-drill",
      `Typology` == "corepreperation?" ~ "core/core preparation piece",
      `Typology` == "preformend of a borer" ~ "tool/perforator-borer; preform",
      `Typology` == "used" ~ "tool/used",
      `Typology` == "distalend of a inverse sidescraper" ~ "tool/side scraper",
      `Typology` == "retouchedblade" ~ "tool/retouched piece",
      `Typology` == "drill" ~ "tool/perforator-drill",
      `Typology` == "corerejuvention tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore" ~ "core/pyramid",
      `Typology` == "flake-bladecore" ~ "core/mixed",
      `Typology` == "notch, truncated" ~ "tool/notch;truncated piece",
      `Typology` == "retouchpiece" ~ "tool/retouched piece",
      `Typology` == "endscraper" ~ "tool/end scraper",
      `Typology` == "notch" ~ "tool/notch",
      `Typology` == "sidescraper" ~ "tool/side scraper",
      `Typology` == "doubleside scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper (?)" ~ "tool/convergent scraper",
      `Typology` == "doublenotch on retouched piece" ~ "tool/double notch",
      `Typology` == "retouchpieces" ~ "tool/retouched piece",
      `Typology` == "burin(?)/ notch/scraper" ~ "tool/burin-notch-scraper",
      `Typology` == "borer" ~ "tool/perforator-borer",
      `Typology` == "retouchedpiece" ~ "tool/retouched piece",
      `Typology` == "notch/side scraper" ~ "tool/notch-side scraper",
      `Typology` == "corefragment" ~ "core/core fragment",
      `Typology` == "backed" ~ "tool/backed knife",
      `Typology` == "doubleside scraper ?" ~ "tool/double side scraper",
      `Typology` == "inversedenticulate" ~ "tool/denticulate",
      `Typology` == "retouchpieces/ side scraper?" ~ "tool/side scraper",
      `Typology` == "sidescraper ?" ~ "tool/side scraper",
      `Typology` == "notch-denticulate" ~ "tool/notch-denticulate",
      `Typology` == "borer/drill" ~ "tool/perforator-drill",
      `Typology` == "retouchedpiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchepiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchedpiece (double side scraper)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece (double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchepiece" ~ "tool/retouched piece",
      `Typology` == "inversenotch" ~ "tool/notch",
      `Typology` == "retouchepiece (double side scaraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece( double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "notchon retouche piece" ~ "tool/notch",
      `Typology` == "retouchedpiece (souble side scraper)" ~ "tool/double side scraper",
      `Typology` == "point" ~ "tool/perforator-point",
      `Typology` == "chunk" ~ "NA/NA",
      `Typology` == "coretablet" ~ "core/core tablet",
      `Typology` == "core" ~ "core/shapeless",
      `Typology` == "denticulate" ~ "tool/denticulate",
      `Typology` == "doublenotch (haft)" ~ "tool/double notch",
      `Typology` == "retouchedpiece (fine retouch)" ~ "tool/retouched piece",
      `Typology` == "endscraper on retouched piece" ~ "tool/end scraper",
      `Typology` == "retouchedpiece (serrated scraper)" ~ "tool/serrated scraper",
      `Typology` == "truncation" ~ "tool/truncated piece",
      `Typology` == "rejuventionpiece" ~ "core/rejuvenation piece",
      `Typology` == "notch/denticulate" ~ "tool/notch-denticulate",
      `Typology` == "retoucheduse on breakage" ~ "tool/retouched piece",
      `Typology` == "inversenotch/endscraper" ~ "tool/notch-end scraper",
      `Typology` == "alternatingside scraper" ~ "tool/side scraper",
      `Typology` == "alternatedouble side scraper" ~ "tool/double side scraper",
      `Typology` == "truncation/sidescaraper" ~ "tool/truncated piece",
      `Typology` == "doublenotch" ~ "tool/double notch",
      `Typology` == "borer/awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper?" ~ "tool/round scraper",
      `Typology` == "doublenotch scraper" ~ "tool/double notch scraper",
      `Typology` == "corerejuvention (core tablet?)" ~ "core/core tablet",
      `Typology` == "truncationاریبب" ~ "tool/truncated piece",
      `Typology` == "retouchedpiece (side scraper)" ~ "tool/side scraper",
      `Typology` == "retochedpice" ~ "tool/retouched piece",
      `Typology` == "truncation/notch" ~ "tool/truncated piece-notch",
      `Typology` == "doublealternate scraper" ~ "tool/double side scraper",
      `Typology` == "awl/ inverse notch" ~ "tool/notch",
      `Typology` == "retouchedpiece (alternate scraper?)" ~ "tool/double side scraper",
      `Typology` == "inverseside scraper" ~ "tool/side scraper",
      `Typology` == "usedcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "corerejuvention platform" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade??" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejivention" ~ "core/rejuvenation piece",
      `Typology` == "retouchedpice (side scraper?)" ~ "tool/side scraper" ,
      `Typology` == "scraper" ~ "tool/side scraper",
      `Typology` == "sickleblade" ~ "tool/sickle shine",
      `Typology` == "coretool (scraper)" ~ "tool/scraper-on core piece",
      `Typology` == "retouchedpiece (scraper)" ~ "tool/side scraper",
      `Typology` == "inversedenticulate?" ~ "tool/denticulate",
      `Typology` == "retouchedpiece (scraper?)" ~ "tool/side scraper",
      `Typology` == "coreside rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "truncation/endscraper" ~ "tool/endscraper on truncated piece",
      `Typology` == "retouchedpice(double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "doublescraper" ~ "tool/double side scraper",
      `Typology` == "dendiculate" ~ "tool/denticulate",
      `Typology` == "double-scraper" ~ "tool/double side scraper",
      `Typology` == "pyramidalcore" ~ "core/pyramid",
      `Typology` == "bulletcore" ~ "core/bullet",
      `Typology` == "corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "pyrmidalcore" ~ "core/pyramid",
      `Typology` == "corepreperation" ~ "core/core preparation piece",
      `Typology` == "crested" ~ "core/crested bladelet",
      `Typology` == "preperationblade" ~ "core/core preperation",
      `Typology` == "burin" ~ "tool/burin",
      `Typology` == "serrateddenticulate" ~ "tool/serrated scraper",
      `Typology` == "saw:serrated denticulate" ~ "tool/serrated scraper",
      `Typology` == "serratedused" ~ "tool/serrated scraper",
      `Typology` == "corepreeration" ~ "tool/core preparation piece",
      `Typology` == "coreprepearation" ~ "tool/core preparation piece",
      `Typology` == "awl" ~ "tool/perforator-awl",
      `Typology` == "notch.inverse" ~ "tool/notch",
      `Typology` == "point.notch" ~ "tool/perforator-point;notch",
      `Typology` == "inversescraper" ~ "tool/side scraper",
      `Typology` == "roundendscraper" ~ "tool/round scraper",
      `Typology` == "alternate.scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper/disc scraper" ~ "tool/scraper-convergent scraper",
      `Typology` == "serratedside scraper" ~ "tool/serrated scraper",
      `Typology` == "debri" ~ "NA/NA",
      `Typology` == "distalpart of a drill" ~ "tool/perforator-drill",
      `Typology` == "corepreperation?" ~ "core/core preparation piece",
      `Typology` == "preformend of a borer" ~ "tool/perforator-borer; preform",
      `Typology` == "used" ~ "tool/used",
      `Typology` == "distalend of a inverse sidescraper" ~ "tool/side scraper",
      `Typology` == "retouchedblade" ~ "tool/retouched piece",
      `Typology` == "drill" ~ "tool/perforator-drill",
      `Typology` == "corerejuvention tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore" ~ "core/pyramid",
      `Typology` == "flake-bladecore" ~ "core/mixed",
      `Typology` == "bladelet-corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "bladelet-corefragment" ~ "core/core fragment",
      `Typology` == "pointedused" ~ "tool/perforator-point",
      `Typology` == "microburin" ~ "tool/microburin",
      `Typology` == "bladeletcore fragment" ~ "core/core fragment",
      `Typology` == "rejuventionface flake" ~ "core/rejuvenation piece",
      `Typology` == "corepreperation flake" ~ "core/core preparation piece",
      `Typology` == "corerejuvention flake"  ~ "core/rejuvenation piece",
      `Typology` == "unfinished  pyramid core" ~ "core/pyramid",
      `Typology` == "bladeletcoretablet" ~ "core/core tablet",
      `Typology` == "bladeletcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "debitage" ~ "NA/NA",
      `Typology` == "bladeletcore  rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated scraper" ~ "tool/serrated scraper",
      `Typology` == "backedknife" ~ "tool/backed knife",
      `Typology` == "corepreparation" ~ "core/core preparation piece",
      `Typology` == "primaryflake" ~ "core/primary flake",
      `Typology` == "pointedconvergent scraper on bladelet" ~ "tool/perforator-point",
      `Typology` == "trie" ~ "NA/NA",
      `Typology` == "part of core tablet" ~ "core/core tablet",
      `Typology` == "zaviedar" ~ "NA/NA",
      `Typology` == "convergentscraper" ~ "tool/convergent scraper",
      `Typology` == "preparationflake" ~ "core/core preparation piece",
      `Typology` == "atypicalborer" ~ "tool/perforator-borer",
      `Typology` == "naturalybacked knife" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "pyramidbladelet core" ~ "core/pyramid",
      `Typology` == "flatbladelet core" ~ "core/flat",
      `Typology` == "bladecore" ~ "core/pyramid",
      `Typology` == "bulletbladelet core" ~ "core/bullet",
      `Typology` == "coerejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "finishedbladelet core" ~ "core/bullet",
      `Typology` == "failed pyramid core?" ~ "core/pyramid",
      `Typology` == "pyramidplain bladelet core" ~ "core/pyramid",
      `Typology` == "coeon flake" ~ "core/core on flake",
      `Typology` == "coeejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "bladeletcoe rejuvention tablet?" ~ "core/core tablet",
      `Typology` == "pyramidcore" ~ "core/pyramid",
      `Typology` == "bladeletcore tablet" ~ "core/core tablet",
      `Typology` == "pyramidbladeleet core" ~ "core/pyramid",
      `Typology` == "flake" ~ "NA/NA",
      `Typology` == "serrated" ~ "tool/serrated scraper",
      `Typology` == "curvedretouched piece" ~ "tool/retouched piece",
      `Typology` == "pyramidalbladelet core" ~ "core/pyramid",
      `Typology` == "multidirectionalcore fragment" ~ "core/multidirectional core fragment",
      `Typology` == "unidirectionalblade? core fragent" ~ "core/pyramid",
      `Typology` == "backed/corerejuvention" ~ "tool/backed knife",
      `Typology` == "unidirectionalbladelet core" ~ "core/pyramid",
      `Typology` == "unifacialbalde core fragment?" ~ "core/pyramid",
      `Typology` == "flakedetached from a bladelet core" ~ "core/core preparation piece",
      `Typology` == "corepreparation tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore fragent" ~ "core/core fragment-pyramid",
      `Typology` == "alternatescraper??" ~ "tool/double side scraper",
      `Typology` == "endscraper.onrejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated" ~ "tool/serrated scraper",
      `Typology` == "patination.omitted" ~ "NA/NA",
      `Typology` == "trapze" ~ "tool/geometric-triangle",
      `Typology` == "trapzoid" ~ "tool/geometric-triangle",
      `Typology` == "borer.drill" ~ "tool/perforator-drill",
      `Typology` == "serraed" ~ "tool/serrated scraper",
      `Typology` == "bladeletcore on a flake" ~ "core/core on flake",
      `Typology` == "geofact" ~ "NA/NA",
      `Typology` == "denticulateborer?" ~ "tool/perforator-borer;denticulate",
      `Typology` == "sideborer-denticulate" ~ "tool/perforator-borer;denticulate",
      `Typology` == "borer.awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper" ~ "tool/round scraper",
      `Typology` == "truncationon a bladelet core" ~ "tool/truncated piece",
      `Typology` == "truncation.notch" ~ "tool/truncated piece;notch",
      `Typology` == "backed-denticulate" ~ "tool/backed;denticulate",
      `Typology` == "brokendrill" ~ "tool/perforator-drill",
      `Typology` == "drillbroken" ~ "tool/perforator-drill",
      `Typology` == "point/broken drill" ~ "tool/perforator-point",
      `Typology` == "brokendrill?" ~ "tool/perforator-drill",
      `Typology` == "partialyserrated" ~ "tool/serrated scraper",
      `Typology` == "naturallybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "transversescraper" ~ "tool/transverse scraper",
      `Typology` == "naturalybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "awl-notch" ~ "tool/perforator-awl",
      `Typology` == "psedulevalois" ~ "tool/psedulevalois flake",
      `Typology` == "alternatingdenticulate" ~ "tool/denticulate",
      `Typology` == "usedcore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalmicroflake core" ~ "core/multidirectional flake core",
      `Typology` == "mixedpyramidal core" ~ "core/pyramid",
      `Typology` == "usedbladecore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalbladelet core fragment" ~ "core/multidirectional bladelet core",
      `Typology` == "unidirectionalbladelet core.cylinder" ~ "core/prismatic",
      `Typology` == "pyrymedalmixed core" ~ "core/pyramid",
      `Typology` == "pyramedial/bullet core" ~ "core/bullet",
      `Typology` == "corefragent" ~ "core/core fragment",
      `Typology` == "pyramedialbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unidirectionalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unifacialunidirectional bladelet core" ~ "core/pyramid",
      `Typology` == "unidirectionalblade core" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "corerejuvention?" ~ "core/rejuvenation piece",
      `Typology` == "flatburin? point?" ~ "tool/flat burin",
      `Typology` == "primaryblade" ~ "core/primary blade",
      `Typology` == "trihedralgeometric" ~ "tool/geometric-triangle",
      `Typology` == "lunate" ~ "tool/geometric-lunate",
      `Typology` == "corepaltform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "pressurebladelet core /semi flat" ~ "core/pyramid",
      `Typology` == "failedpercussion blade core/semi pyramedal" ~ "core/pyramid",
      `Typology` == "percussionbladelet core/unidirectional" ~ "core/pyramid-percussion",
      `Typology` == "percussion.unidirectional.blade-flakecore" ~ "core/mixed-percussion",
      `Typology` == "obliqueretouched bladelet" ~ "tool/retouched piece",
      `Typology` == "percussionunidirectional pyramidal  bladelet core" ~ "core/pyramid",
      `Typology` == "lunategeometric" ~ "tool/geometric-lunate",
      `Typology` == "awl on a core rejuvention" ~ "tool/perforator-awl on core rejuvenation piece",
      `Typology` == "truncated" ~ "tool/truncated piece",
      `Typology` == "multiplenotch" ~ "tool/notch",
      `Typology` == "dishedconvergent bladelet" ~ "tool/convergent scraper",
      `Typology` == "bipolarpercussion blade-bladelet core" ~ "core/bipolar-percussion",
      `Typology` == "bladeletbullet core" ~ "core/bullet",
      `Typology` == "alternatingserrated" ~ "tool/serrated scraper",
      `Typology` == "shaplesscore" ~ "core/shapeless",
      `Typology` == "corepreparation?" ~ "core/core preparation piece",
      `Typology` == "obliqueretouched" ~ "tool/retouched piece",
      `Typology` == "corepreparation flake" ~ "core/core preparation piece",
      `Typology` == "backeddenticulate" ~ "tool/backed;denticulate",
      `Typology` == "awlnotch" ~ "tool/perforator-awl;notch",
      `Typology` == "scraperburin?" ~ "tool/burin;scraper",
      `Typology` == "notchround scraper" ~ "tool/round scraper;notch",
      `Typology` == "corerejuvention tablet round scraper" ~ "tool/round scraper on core tablet",
      `Typology` == "alternatescraper" ~ "tool/double side scraper",
      `Typology` == "coreplatform preperation flake" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention tablet" ~ "core/core tablet",
      `Typology` == "brokenpyramidal core bladelet" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core/semi bullet?" ~ "core/bullet",
      `Typology` == "coreside rejuvention flake?" ~ "core/rejuvenation piece-side",
      `Typology` == "unidirectionalpressure bladelet core/unifacila" ~ "core/pyramid",
      `Typology` == "coreplatform rejuvention tablet/or multidirectional core" ~ "core/rejuvenation piece",
      `Typology` == "bulletbladlet core" ~ "core/bullet",
      `Typology` == "cylinderbipolar bladelet core" ~ "core/prismatic",
      `Typology` == "backedknife/core platform rejuvenation" ~ "tool/backed knife",
      `Typology` == "flatunifacial bladelet core" ~ "core/flat",
      `Typology` == "coreprepration flake" ~ "core/core preparation piece",
      `Typology` == "coreplatform preperation" ~ "core/core preparation piece-platform",
      `Typology` == "borerpreform" ~ "tool/perforator-borer",
      `Typology` == "rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "borerpreform?" ~ "tool/perforator-borer",
      `Typology` == "inversnotch" ~ "tool/notch",
      `Typology` == "alternateconvergent" ~ "tool/convergent scraper",
      `Typology` == "unidirectionalbladelet/flake-blade core" ~ "core/mixed",
      `Typology` == "unidirectionalmixed core" ~ "core/mixed",
      `Typology` == "unidirectionalflat bladelet core" ~ "core/flat",
      `Typology` == "" ~ "NA-NA",
      `Typology` == "na" ~ "NA-NA",
      `Typology` == "bladelet-corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "bladelet-corefragment" ~ "core/core fragment",
      `Typology` == "pointedused" ~ "tool/perforator-point",
      `Typology` == "microburin" ~ "tool/microburin",
      `Typology` == "bladeletcore fragment" ~ "core/core fragment",
      `Typology` == "rejuventionface flake" ~ "core/rejuvenation piece",
      `Typology` == "corepreperation flake" ~ "core/core preparation piece",
      `Typology` == "corerejuvention flake"  ~ "core/rejuvenation piece",
      `Typology` == "unfinished  pyramid core" ~ "core/pyramid",
      `Typology` == "bladeletcoretablet" ~ "core/core tablet",
      `Typology` == "bladeletcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "debitage" ~ "NA/NA",
      `Typology` == "bladeletcore  rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated scraper" ~ "tool/serrated scraper",
      `Typology` == "backedknife" ~ "tool/backed knife",
      `Typology` == "corepreparation" ~ "core/core preparation piece",
      `Typology` == "primaryflake" ~ "core/primary flake",
      `Typology` == "pointedconvergent scraper on bladelet" ~ "tool/perforator-point",
      `Typology` == "trie" ~ "NA/NA",
      `Typology` == "part of core tablet" ~ "core/core tablet",
      `Typology` == "zaviedar" ~ "NA/NA",
      `Typology` == "convergentscraper" ~ "tool/convergent scraper",
      `Typology` == "preparationflake" ~ "core/core preparation piece",
      `Typology` == "atypicalborer" ~ "tool/perforator-borer",
      `Typology` == "naturalybacked knife" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "pyramidbladelet core" ~ "core/pyramid",
      `Typology` == "flatbladelet core" ~ "core/flat",
      `Typology` == "bladecore" ~ "core/pyramid",
      `Typology` == "bulletbladelet core" ~ "core/bullet",
      `Typology` == "coerejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "finishedbladelet core" ~ "core/bullet",
      `Typology` == "failed pyramid core?" ~ "core/pyramid",
      `Typology` == "pyramidplain bladelet core" ~ "core/pyramid",
      `Typology` == "coeon flake" ~ "core/core on flake",
      `Typology` == "coeejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "bladeletcoe rejuvention tablet?" ~ "core/core tablet",
      `Typology` == "pyramidcore" ~ "core/pyramid",
      `Typology` == "bladeletcore tablet" ~ "core/core tablet",
      `Typology` == "pyramidbladeleet core" ~ "core/pyramid",
      `Typology` == "flake" ~ "NA/NA",
      `Typology` == "serrated" ~ "tool/serrated scraper",
      `Typology` == "curvedretouched piece" ~ "tool/retouched piece",
      `Typology` == "pyramidalbladelet core" ~ "core/pyramid",
      `Typology` == "multidirectionalcore fragment" ~ "core/multidirectional core fragment",
      `Typology` == "unidirectionalblade? core fragent" ~ "core/pyramid",
      `Typology` == "backed/corerejuvention" ~ "tool/backed knife",
      `Typology` == "unidirectionalbladelet core" ~ "core/pyramid",
      `Typology` == "unifacialbalde core fragment?" ~ "core/pyramid",
      `Typology` == "flakedetached from a bladelet core" ~ "core/core preparation piece",
      `Typology` == "corepreparation tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore fragent" ~ "core/core fragment-pyramid",
      `Typology` == "alternatescraper??" ~ "tool/double side scraper",
      `Typology` == "endscraper.onrejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated" ~ "tool/serrated scraper",
      `Typology` == "patination.omitted" ~ "NA/NA",
      `Typology` == "trapze" ~ "tool/geometric-triangle",
      `Typology` == "trapzoid" ~ "tool/geometric-triangle",
      `Typology` == "borer.drill" ~ "tool/perforator-drill",
      `Typology` == "serraed" ~ "tool/serrated scraper",
      `Typology` == "bladeletcore on a flake" ~ "core/core on flake",
      `Typology` == "geofact" ~ "NA/NA",
      `Typology` == "denticulateborer?" ~ "tool/perforator-borer;denticulate",
      `Typology` == "sideborer-denticulate" ~ "tool/perforator-borer;denticulate",
      `Typology` == "borer.awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper" ~ "tool/round scraper",
      `Typology` == "truncationon a bladelet core" ~ "tool/truncated piece",
      `Typology` == "truncation.notch" ~ "tool/truncated piece;notch",
      `Typology` == "backed-denticulate" ~ "tool/backed;denticulate",
      `Typology` == "brokendrill" ~ "tool/perforator-drill",
      `Typology` == "drillbroken" ~ "tool/perforator-drill",
      `Typology` == "point/broken drill" ~ "tool/perforator-point",
      `Typology` == "brokendrill?" ~ "tool/perforator-drill",
      `Typology` == "partialyserrated" ~ "tool/serrated scraper",
      `Typology` == "naturallybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "transversescraper" ~ "tool/transverse scraper",
      `Typology` == "naturalybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "awl-notch" ~ "tool/perforator-awl",
      `Typology` == "psedulevalois" ~ "tool/psedulevalois flake",
      `Typology` == "alternatingdenticulate" ~ "tool/denticulate",
      `Typology` == "usedcore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalmicroflake core" ~ "core/multidirectional flake core",
      `Typology` == "mixedpyramidal core" ~ "core/pyramid",
      `Typology` == "usedbladecore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalbladelet core fragment" ~ "core/multidirectional bladelet core",
      `Typology` == "unidirectionalbladelet core.cylinder" ~ "core/prismatic",
      `Typology` == "pyrymedalmixed core" ~ "core/pyramid",
      `Typology` == "pyramedial/bullet core" ~ "core/bullet",
      `Typology` == "corefragent" ~ "core/core fragment",
      `Typology` == "pyramedialbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unidirectionalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unifacialunidirectional bladelet core" ~ "core/pyramid",
      `Typology` == "unidirectionalblade core" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "corerejuvention?" ~ "core/rejuvenation piece",
      `Typology` == "flatburin? point?" ~ "tool/flat burin",
      `Typology` == "primaryblade" ~ "core/primary blade",
      `Typology` == "trihedralgeometric" ~ "tool/geometric-triangle",
      `Typology` == "lunate" ~ "tool/geometric-lunate",
      `Typology` == "corepaltform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "pressurebladelet core /semi flat" ~ "core/pyramid",
      `Typology` == "failedpercussion blade core/semi pyramedal" ~ "core/pyramid",
      `Typology` == "percussionbladelet core/unidirectional" ~ "core/pyramid-percussion",
      `Typology` == "percussion.unidirectional.blade-flakecore" ~ "core/mixed-percussion",
      `Typology` == "obliqueretouched bladelet" ~ "tool/retouched piece",
      `Typology` == "percussionunidirectional pyramidal  bladelet core" ~ "core/pyramid",
      `Typology` == "lunategeometric" ~ "tool/geometric-lunate",
      `Typology` == "awl on a core rejuvention" ~ "tool/perforator-awl on core rejuvenation piece",
      `Typology` == "truncated" ~ "tool/truncated piece",
      `Typology` == "multiplenotch" ~ "tool/notch",
      `Typology` == "dishedconvergent bladelet" ~ "tool/convergent scraper",
      `Typology` == "bipolarpercussion blade-bladelet core" ~ "core/bipolar-percussion",
      `Typology` == "bladeletbullet core" ~ "core/bullet",
      `Typology` == "alternatingserrated" ~ "tool/serrated scraper",
      `Typology` == "shaplesscore" ~ "core/shapeless",
      `Typology` == "corepreparation?" ~ "core/core preparation piece",
      `Typology` == "obliqueretouched" ~ "tool/retouched piece",
      `Typology` == "corepreparation flake" ~ "core/core preparation piece",
      `Typology` == "backeddenticulate" ~ "tool/backed;denticulate",
      `Typology` == "awlnotch" ~ "tool/perforator-awl;notch",
      `Typology` == "scraperburin?" ~ "tool/burin;scraper",
      `Typology` == "notchround scraper" ~ "tool/round scraper;notch",
      `Typology` == "corerejuvention tablet round scraper" ~ "tool/round scraper on core tablet",
      `Typology` == "alternatescraper" ~ "tool/double side scraper",
      `Typology` == "coreplatform preperation flake" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention tablet" ~ "core/core tablet",
      `Typology` == "brokenpyramidal core bladelet" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core/semi bullet?" ~ "core/bullet",
      `Typology` == "coreside rejuvention flake?" ~ "core/rejuvenation piece-side",
      `Typology` == "unidirectionalpressure bladelet core/unifacila" ~ "core/pyramid",
      `Typology` == "coreplatform rejuvention tablet/or multidirectional core" ~ "core/rejuvenation piece",
      `Typology` == "bulletbladlet core" ~ "core/bullet",
      `Typology` == "cylinderbipolar bladelet core" ~ "core/prismatic",
      `Typology` == "backedknife/core platform rejuvenation" ~ "tool/backed knife",
      `Typology` == "flatunifacial bladelet core" ~ "core/flat",
      `Typology` == "coreprepration flake" ~ "core/core preparation piece",
      `Typology` == "coreplatform preperation" ~ "core/core preparation piece-platform",
      `Typology` == "borerpreform" ~ "tool/perforator-borer",
      `Typology` == "rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "borerpreform?" ~ "tool/perforator-borer",
      `Typology` == "inversnotch" ~ "tool/notch",
      `Typology` == "alternateconvergent" ~ "tool/convergent scraper",
      `Typology` == "unidirectionalbladelet/flake-blade core" ~ "core/mixed",
      `Typology` == "unidirectionalmixed core" ~ "core/mixed",
      `Typology` == "unidirectionalflat bladelet core" ~ "core/flat",
      `Typology` == "" ~ "NA-NA",
      `Typology` == "na" ~ "NA-NA",
      TRUE ~ `Typology`
    )
  )  %>%
  mutate(
    typology_split = str_split(Typology, "/", simplify = TRUE),
    `typology-tool-core` = typology_split[, 1],
    tool_or_core = ifelse(str_starts(`typology-tool-core`, "tool"), "tool", "core"),
    `tool-typology` = ifelse(tool_or_core == "tool", typology_split[, 2], NA),
    `core-fragment` = ifelse(tool_or_core == "core", typology_split[, 2], NA)
  ) %>%
  # remove temporary columns
  select(-typology_split,-tool_or_core)

# BM: still some typos in there
rev(sort(table((tl4$`tool-typology`))))
rev(sort(table((tl4$`core-fragment`))))
```

```{r}
# Exploring data

tl4 %>% 
  # BM: we need this because of the mutate_all ifelse combo you have above
  mutate(midpoint = parse_number(midpoint)) %>% 
  group_by(midpoint) %>% 
  tally() %>% 
  drop_na(midpoint) %>% 
ggplot() + 
  aes(midpoint, n) +
  geom_col()

```

```{r}

# refine the table

tl_final <- tl4 %>%
  rename(
    blank =                Blank2,
   `sickle shine` =       shine2,
    length =              `Length (mm)` ,
    width =               `Width (mm)`,
    thickness =           `Thickness (mm)`, 
    typology =             Typology,
    utilization =          Utilization,
    retouch =              Retouch,
    blank =                Blank2, 
   `blank part` =         `Blank_part`,
   `bulb thickness` =     `Bulb Thickness (mm)`, 
   `platform thickness` = `Platform Thickness (mm)`, 
   `platform length` =    `platform Length (mm)`, 
    eraillure =            Eraillure , 
    cortex =               `Cortex (%)`)  %>%
  mutate(al = paste0(toupper(area), "_", layer)) %>%
  relocate(al, .after = layer)   %>%
  select(lithic_id, 
         TA, 
         area, 
         layer,
         al,
         cortex,
         depth, upper, lower, midpoint, 
        `length`, `width`, `thickness`, 
         typology, 
         utilization, 
        `sickle shine`,
         retouch, 
        `typology-tool-core`,
        `tool-typology`,
         blank, `blank part`, 
        `core-typology`, `core-fragment`,`core-technology`, 
        cortex)  %>% 
# Update the 'typology-tool-core' column based on 'Retouch'
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",

      TRUE ~ `typology-tool-core`      
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
      utilization == 1 ~ "utilized tool",
      `sickle shine` == 1 ~ "sickle shine",
      TRUE ~ `tool-typology`               
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
     # If Retouch is 1 and tool-typology is NA
       retouch == 1 & is.na(`tool-typology`) ~ "retouched piece",
       `sickle shine` == 1 & is.na(`tool-typology`) ~ "sickle shine",

      # For all other cases 
      TRUE ~ `tool-typology`                                       
    )
  )

# BM: take a look at tool typology
rev(sort(table(tl_final$`tool-typology`)))

```

```{r}
# List of columns that should be numeric
columns_to_convert <- c("cortex", 
                        "upper", 
                        "lower", 
                        "midpoint", 
                        "length", 
                        "width", 
                        "utilization", 
                        "retouch",
                        "sickle shine")

# Convert each column to numeric
for (col_name in columns_to_convert) {
  # Replace non-numeric characters with NA
  tl_final[[col_name]][tl_final[[col_name]] %in% c("", "NA", "N/A")] <- NA
  # Convert to numeric
  tl_final[[col_name]] <- as.numeric(tl_final[[col_name]])
}

# check the structure to see if they have been converted
str(tl_final)
```

```{r}
#set the value of utilization to NA only for the rows where `sickle shine` == 1.

tl_final <- tl_final %>%
  rowwise() %>%
  mutate(utilization = ifelse(is.na(`sickle shine`) | is.na(utilization), utilization,
                              ifelse(`sickle shine` == 1 & utilization == 1, NA_real_, utilization))) %>%
  ungroup()

```

```{r}
#number of `sickle shine` in clumn `sickle shine`=1 and tooltypegroup are not match: 
rev(sort(table(tl_final$tooltypegroup)))
rev(sort(table(tl_final$`sickle shine`)))

#troubleshoot:
tl_final %>% filter(`sickle shine` == 1) %>% 
  select(`tool-typology`, 
         `sickle shine`)

tl_final %>% filter(`tool-typology` == "sickle shine") %>% 
  select(`tool-typology`, `sickle shine`)

tl_final <- tl_final %>%
  mutate(
    `tool-typology` = case_when(
      `sickle shine` == 1 ~ "sickle shine",
      utilization == 1 ~ "utilized tool",
      TRUE ~ `tool-typology`               
    )
  )

rev(sort(table(tl_final$`tool-typology`)))
```

```{r}

# Creating tooltypegroup

tl_final <- tl_final %>%
  mutate(
    tooltypegroup = case_when(
      `tool-typology` %in%
        c(
          "end scraper",
          "side scraper",
          "double side scraper",
          "convergent scraper",
          "round scraper",
          "perforator-convergent scraper",
          "perforator-borer; preform",
          "scraper-on core piece",
          "endscraper on truncated piece",
          "round scraper",
          "transverse scraper",
          "core tool-used core fragment",
          "round scraper on core tablet"
        ) ~ "scraper",
      `tool-typology` %in%
        c(
          "perforator-borer",
          "perforator-drill",
          "perforator-point",
          "perforator-awl",
          "perforator-point;notch",
          "perforator-awl on core rejuvenation piece",
          "perforator-awl;notch",
          "perforator-borer;denticulate"
        ) ~ "Perforator",
      `tool-typology` %in%
        c(
          "backed knife",
          "backed knife-naturally backed knife",
          "backed;denticulate"
        ) ~ "backed pieces",
      `tool-typology` %in% c(
        "double notch",
        "denticulate",
        " denticulate",
        "notch-denticulate",
        "notch"
      ) ~ "denticulate-notch",
      `tool-typology` %in% c(
        "notch-side scraper",
        "notch-end scraper",
        "double notch scraper",
        "round scraper;notch"
      ) ~ "scraper-notch",
      `tool-typology` %in% c("geometric-triangle", "geometric-lunate") ~ "geometric",
      `tool-typology` == "serrated scraper" ~ "serrated scraper",
      `tool-typology` %in% c(
        "truncated piece",
        "truncated piece-notch",
        "truncated piece;notch",
        "notch;truncated piece"
      ) ~ "truncated pieces",

      # Represents NA group
      `tool-typology` %in% c("core preparation piece", "used", "psedulevalois flake") ~ NA_character_,
      `tool-typology` %in% c("burin", "flat burin", "burin;scraper", "burin-notch-scraper") ~ "burin",
      `tool-typology` == "microburin" ~ "microburin",
      `tool-typology` == "retouched piece" ~ "retouched piece",
      `tool-typology` == "utilized tool" ~ "utilized tool",
      `tool-typology` == "sickle shine" ~ "sickle shine",

      #classify any remaining values as NA
      TRUE ~ NA_character_
    )
  )

# Checking the unique values of tooltypegroup
rev(sort(table(tl_final$tooltypegroup)))

```

```{r}
tl_final <- tl_final %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`))  %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`)) %>%
  mutate(
    `core-fragment` = case_when(
      `core-fragment` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-fragment` == "rejuvention piece-side" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece-platform" ~ "core tablet",
      `core-fragment` == "core preparation" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece" ~ "rejuvenation piece",
      `core-fragment` == "core preperation" ~ "rejuvenation piece",
      `core-fragment` == "pyramid" ~ "pyramid",
      `core-fragment` == "pyramid core" ~ "pyramid",
      `core-fragment` == "pyramid-unidirectional" ~ "pyramid",
      `core-fragment` == "core fragment-pyramid" ~ "pyramid",
      `core-fragment` == "cylinder" ~ "cylinder/prismatic",
      `core-fragment` == "prismatic" ~ "cylinder/prismatic",
      `core-fragment` == "shapeless" ~ "cylinder/prismatic",
      `core-fragment` == "multidirectional core fragment" ~ "shapeless",
      `core-fragment` == "multidirectional flake core" ~ "shapeless",
      `core-fragment` == "multidirectional bladelet core" ~ "shapeless",
      `core-fragment` == "mixed core" ~ "shapeless",
      `core-fragment` == "mixed-percussion" ~ "shapeless",
      `core-fragment` == "mixed" ~ "shapeless",
      `core-fragment` == ("bipolar-percussion") ~ "bipolar",
      `core-fragment` == "flat" ~ "flat",
      `core-fragment` == "core on flake" ~ "core on flake",
      `core-fragment` == "bullet" ~ "bullet",
      TRUE ~ `core-fragment`

    )
  )  %>%
  mutate(
    core = case_when(
      `core-fragment` %in% c(
        "pyramid",
        "pyramid core",
        "pyramid-unidirectional",
        "core fragment-pyramid",
        "cylinder/prismatic",
        "shapeless",
        "multidirectional core fragment",
        "multidirectional flake core",
        "multidirectional bladelet core",
        "mixed core",
        "mixed-percussion",
        "mixed",
        "bipolar-percussion",
        "flat",
        "core on flake",
        "bullet"
      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    ),
    corefrag = case_when(
      `core-fragment` %in% c(
        "rejuvenation piece",
        "rejuvention piece-side",
        "core preparation piece-platform",
        "core preparation",
        "core preparation piece",
        "core preperation",
        "crested bladelet" ,
        "primary flake" ,
        "primary blade"
      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    )
  ) %>%
  # Convert empty strings to NA in `core-fragment` column
  mutate(`core-fragment` = if_else(`core-fragment` == "", NA_character_, `core-fragment`)) %>%
  # Convert "pyramid-percussion" to "pyramid" in `core-fragment` column
  mutate(
    `core-fragment` =
      if_else(
        `core-fragment` == "pyramid-percussion",
        "pyramid",
        `core-fragment`
      )
  ) %>%
  # Update existing `core` column with selected values from `core-fragment` column
  mutate(core = if_else(
    `core-fragment` %in%
      c(
        "pyramid",
        "cylinder/prismatic",
        "flat",
        "bullet",
        "core on flake",
        "pyramid"
      ),
    `core-fragment`,
    core
  )) %>%
  # Add 'core tablet' and 'core fragment' from `core-fragment` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-fragment` == "core tablet" ~ "core tablet",
      `core-fragment` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-typology` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology`
  mutate(`core-typology` = if_else(
    `core-typology` %in% c("rejuvenation piece", "core fragment"),
    NA_character_,
    `core-typology`
  ))  %>% 
# Update the 'blank' column based on conditions in the 'core-fragment' column
  mutate(blank = if_else(
    `core-fragment` %in% c(
      "core fragment",
      "core tablet",
      "rejuvenation piece",
      "pyramid",
      "crested bladelet",
      "shapeless",
      "primary flake",
      "flat",
      "bullet",
      "core on flake",
      "cylinder/prismatic",
      "primary blade",
      "rejuvenation piece-side",
      "core platform rejuvention"
    ),
    NA_character_,
    blank
  ))

# take a look
rev(sort(table(tl_final$core)))
rev(sort(table(tl_final$corefrag)))
rev(sort(table(tl_final$`core-fragment`)))
rev(sort(table(tl_final$blank)))
```

```{r}
library(dplyr)

# Filter the data to include only the areas 'ta', 'd1', and 'tp1'
tl_final_area <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1"))

```

Summary tables: 1. AREA

```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
    `Cores (n)` = sum(!is.na(core) & core != ""),
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) &
                                            corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

# BM: looks like this is area and layer, is that right? I see c1, d1, d2, d3, d4 etc.
# SS: those are different excavated areas, we mainly consider ta, tp1, and d1.
knitr::kable(summary_table_area)

```

```{r}
library(kableExtra)

kable(summary_table_area, 
      caption = "Table: Summary by Area", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

2.LAYER

```{r}
# Create a summary table
summary_table_layer <- tl_final %>%
  group_by(layer) %>% 
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```

```{r}
# Generate the table
kable(summary_table_layer, 
      caption = "Table: Summary by Layer", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

AREA-LAYER:

```{r}
# Create a summary table based on the 'al' column
summary_table_al <- tl_final %>%
  group_by(al) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```

```{r}
# Generate the table

kable(summary_table_al, 
      caption = "Table: Summary by Al", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

Techno-Typo:

```{r}

# Prepare data for Summary table-area
summary_table_area <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(area) %>%
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

summary_table_area <- summary_table_area %>%
  add_row(
    area = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  # Explicitly setting the total
  )

# Adding a new row with total counts
summary_table_area <- summary_table_area %>%
  add_row(
    area = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  # Explicitly setting the total
  )

```

```{r}
# Prepare data for Summary table-area_layer(al)
summary_table_al <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>% 
  group_by(al) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
summary_table_al <- summary_table_al %>% 
  add_row(
    al = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  
  )

rev(sort(table(tl_final$tooltypegroup)))
```

```{r}
#explore tool type group

summary_table_tooltype <- tl_final %>%
  filter(tooltypegroup != 'utilized tool') %>%  # Exclude 'utilized tool'
  group_by(tooltypegroup) %>% 
  summarise(

  )

```

```{r}
#generate a table

summary_table_al <- summary_table_al %>%
  filter(!(al %in% c("D1_NA", "TP1_NA")))

kable(
  summary_table_al,
  col.names = c(
    "Area-Layer",
    "Blade",
    "Flake",
    "Bladelet",
    "Tools",
    "Cores",
    "Core Rejuvenations",
    "Retouched Tools",
    "Utilized Tools",
    "Sickle Shine",
    "Total"
  ),
  caption = "Table: Summary ",
  align = 'c'
) %>%
  kable_styling(
    bootstrap_options = c("striped", 
                          "hover", 
                          "condensed"),
    full_width = FALSE,
    font_size = 12
  ) %>%
  row_spec(0, 
           bold = TRUE, 
           font_size = 14)

```

```{r}
# Create a summary table for typology based on area

summary_table_tool_area <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  filter(!is.na(tooltypegroup), tooltypegroup != "") %>%  # Exclude rows where 'tooltypegroup' is NA or empty
  group_by(area, tooltypegroup) %>%  # Group the data by area and tooltypegroup
  tally() %>%  # Count the number of rows in each group
  arrange(area, tooltypegroup) %>%  # Arrange by area and tooltypegroup
  spread(key = area, value = n, fill = 0)  # Spread the 'area' column into multiple columns

# Add a "Total" column that sums each row
summary_table_tool_area <- summary_table_tool_area %>%
  mutate(Total = rowSums(select(.,-tooltypegroup)))

total_row <- summary_table_tool_area %>%
  select(-tooltypegroup) %>%
  summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(tooltypegroup = "Total")

summary_table_tool_area <-
  bind_rows(summary_table_tool_area, total_row)

```

```{r}

kable(
  summary_table_tool_area,
  col.names = c("Tools", "D1", "TA", "TP1", "Total"),
  caption = "Table: Tool Typology in three different main areas.",
  align = 'c'
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = F,
    font_size = 12
  ) %>%
  column_spec(1, extra_css = "text-align:left;") %>%
  row_spec(0, bold = TRUE, font_size = 14)

```

```{r}
# Create the summary table based on layer
summary_table1_layer <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(layer) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    numb_cores = sum(!is.na(`core-typology`), na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

```

```{r}
# Create a summary table based on the 'depth', 'area', and 'layer' columns
tl_final <- tl_final %>%
  mutate(depth = gsub("cm", "", depth))

summary_table_depth_area_layer <- tl_final %>%
  group_by(TA, depth, area, layer, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

# Generate the table
library(kableExtra)

kable(summary_table_depth_area_layer, 
      caption = "Table: Summary by Depth, Area, and Layer", 
      align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F, 
                font_size = 12) %>%
  row_spec(0, bold = TRUE, font_size = 14)
```

```{r}
#Here, I've imported the table 'Summary by Depth, Area, Layer' into Google Sheets for easier management.
#I thought this format would simplify the process of those depths needed to be splitting up.

# BM: looks like you need this data starting from around here in the document

library(tidyverse)
library(googlesheets4)

google_sheet_url <- 
  "https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0"

sum_depth_sheet <- read_sheet(google_sheet_url)

# BM: from my perspective, what would be ideal here is to have the 
# excavation area and volume calculations done here in R. I feel anxious about
# the possibilities of typos when these values are calculated somewhere I can't 
# see and pasted into the google sheet. I prefer for the workflow of calculations
# to be as transparent and traceable as possible. From this point in the qmd I
# think we can use the volume values from this google sheet. Another option is to 
# export the google sheet as a CSV, and put that in the data folder in this project. 
# Then we can use read_csv to import it.

#SS: is that true?: (I have a weird number in row 5, cl Depth and couldn't find out what happened)

# Separate `excavation area` into `length` and `width` and calculate `volume`
sum_depth_sheet <- sum_depth_sheet %>%
  separate(`excavation area`, into = c("length", "width"), sep = "\\*") %>%
  mutate(
    length = as.numeric(length),
    width = as.numeric(width),
    `thickness of deposit` = `thickness of deposit` / 100  # Convert from cm to m
  ) %>%
  mutate(
    volume = length * width * `thickness of deposit`  # Calculate volume in m^3
  )

# Calculate lithic volumetric density 
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    `lithic_dens` = round(total_lithics / volume, 2)
  )

# Calculate retouch frequency 
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    retouch_freq = round(retouch_tool / volume, 2)  
  )

```

```{r}
#WABI plot: here I do not consider those levels with a value 0 of retouch frequency.

library(ggpmisc)

# Filter the data (get out those level with Retouch frequency=0)
filtered_data <- sum_depth_sheet %>% filter(lithic_dens > 0 & retouch_freq > 0)

# Create the plot: WABI_Level_logscale
WABI_Level <- ggplot(filtered_data,  
            aes(x = lithic_dens,
                y = retouch_freq)) +
  geom_point(color = "black", size = 1.5) +
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              linewidth = 0.5) +
  # Add text labels 
  geom_text(aes(label = paste(area, layer, level, sep = "_")), 
            vjust = 1.5, 
            hjust = 0.5, 
            check_overlap = TRUE) +

  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                  stat(rr.label), 
                                  stat(p.value.label), 
                                  sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +
  scale_x_log10() +
  scale_y_log10() +
  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

# Print the plot to display it
print(WABI_Level)

```

```{r}
#PCA_Level: preparing data

#PCA_level: for PCA Bicho and Cascalheira considered these variables: estimated area (as we calculated based on levels, should we not consider it?), core frequency, blank frequency, chip frequency: artefacts smaller than 1 cm (as we work on a neolithic asssemblage, we have so many small artefacts; so, I do not consider it), feature frequency (we have some stone alignments, but we do not know the exact level, we know the area of those features), retouch frequency, tool diversity (diversity of tool types within each assemblage, calculated using Menhinick’s index); 
##core frequency
##blank frequency
##retouch frequency
##tool diversity

#Calculate Core Frequency
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    core_freq = round(cores / volume, 2)  
  )

# calculate Blank Frequency
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    blank_freq = round((flake + blade + bladelet) / volume, 2)  
  )

# calculate Tool Diversity
##calculating tool diversity is a bit tricky here; as I have some spitted depth, which we assign the number of artefacts based on the calculation of splitting, so how should I consider the tool diversity here?
```

PCA_al:

```{r}
#PCA: area_layer (al)
#I think as we are going to calculate PCA to understand other factors that might be affected the duration of the occupation, it would be logical to calculate it for each layer, instead of level. we have 5 stratigraphical layers: e, d, c, b, a
#d1:1.5*1.5/ ta: 4*5/ tp1:2*1.8
#e: excavated area is A/ depth of excavation is 0-75(mean)
#d: excavated area is A/ 0.75m-3m→ 2.25
#c: d1 and TP → 30 cm and 40cm
#b: TP/ 170 cm 
#a: D1: 70cm

# Add 'dimension' columns and relocate it after 'al' (based on Hole (1974))
library(dplyr)

summary_table_al_pca <- summary_table_al %>%
  mutate(
    length = case_when(
      al == "D1_a" ~ 1.5,
      al == "D1_c" ~ 1.5,
      al == "TA_d" ~ 4,
      al == "TP1_c" ~ 2,
      al == "TA_e" ~ 4,
      TRUE ~ NA_real_
    ),
    width = case_when(
      al == "D1_a" ~ 1.5,
      al == "D1_c" ~ 1.5,
      al == "TA_d" ~ 5,
      al == "TP1_c" ~ 1.8,
       al == "TA_e" ~ 5,
      TRUE ~ NA_real_
    ),
    thickness = case_when(
      al == "D1_a" ~ 0.7,
      al == "D1_c" ~ 0.3,
      al == "TA_d" ~ 2.25,
      al == "TP1_c" ~ 0.4,
      al == "TA_e" ~ 0.75,
      TRUE ~ NA_real_)
  ) %>%

  mutate(volume = length * width * thickness) %>%
  relocate(length, width, thickness, volume, .after = al)
# Show the updated data frame
summary_table_al_pca

```

```{r}
# Load required library
library(dplyr)

# Add a new column called 'blank_count'
summary_table_al_pca <- summary_table_al_pca %>%
  mutate(blank_count = blade_count + flake_count + bladelet_count) %>%
  select(al, length, width, thickness, volume, blade_count, flake_count, bladelet_count, blank_count, everything())

# Calculate core frequency, blank frequency, retouch frequency, lithic density 
summary_table_al_pca <- summary_table_al_pca %>%
  mutate(
    lithic_density = round((core_count + blank_count + core_rejuvenation_count) / volume, 2),
    core_freq = round(core_count / volume, 2),
    blank_freq = round(blank_count / volume, 2),
    retouch_freq = round(retouched_tool_count / volume, 2)
  )

# Calculate tool diversity using Menhinick's index
summary_table_al_with_diversity <- tl_final %>% 
  filter(retouch == 1) %>%
  group_by(al, `tool-typology`) %>%  
  summarise(n = n(), .groups = 'drop') %>% 
  group_by(al) %>%  
  summarise(
    num_unique_tool_types = n_distinct(`tool-typology`),
    total_retouched_tools = sum(n),
    tool_diversity = round(num_unique_tool_types / sqrt(total_retouched_tools), 2),
    .groups = 'drop'
  )

# Merge this new information into the existing summary_table_al_pca
summary_table_al_pca <- left_join(summary_table_al_pca, summary_table_al_with_diversity, by = "al")

# Remove the 'Total' row and select only the columns needed for PCA
fin_summary_table_al_pca <- summary_table_al_pca %>%
  filter(al != "Total") %>%
  select(al, core_freq, blank_freq, lithic_density, retouch_freq, tool_diversity)

# Show the final table
print(fin_summary_table_al_pca)

```

```{r}
library(FactoMineR)
# Perform PCA
res.pca <- PCA(fin_summary_table_al_pca[, -1], graph = FALSE)  # Exclude the 'al' column

```

```{r}
## Got error: couldnt figure out the reason of error

# Create a data frame to hold the rounded eigenvalues and the rounded percentage variance
eigen_df <- data.frame(
  Dimension = paste("Dim", 1:nrow(eigenvalues), sep = "."),
  Eigenvalue = round(eigenvalues[, 1], 3),
  `Variance percent` = round(eigenvalues[, 2], 3),
  `Cumulative variance percent` = round(cumsum(eigenvalues[, 2]), 3)
)

# Print the table
print(eigen_df)

```

```{r}
# variables to PC1
fviz_contrib(res.pca,  
             choice = "var",  
             axes = 1,  
             top = 10)
```

```{r}
fviz_contrib(res.pca,  
             choice = "var",  
             axes = 2,  
             top = 10)

```

```{r}
fviz_contrib(res.pca,  
             choice = "var",  
             axes = 3,  
             top = 10)
```

```{r}
fviz_contrib(res.pca,  
             choice = "var",  
             axes = 4,  
             top = 10)
```

```{r}
#Plot: Contribution of variables for each of the four relevant PCA dimensions
library(factoextra)
library(ggplot2)

#save contribution plots into one objects
p1 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 1, 
                   top = 10)
p2 <- fviz_contrib(res.pca, 
                   choice = "var",
                   axes = 2, 
                   top = 10)
p3 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 3, 
                   top = 10)
p4 <- fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = 4, 
                   top = 10)

#combine individual plots into one plot 
library(gridExtra)
grid.arrange(p1, p2, p3, p4, ncol = 2)

```

```{r}
# Extract contributions of variables to each principal component
contributions <- res.pca$var$contrib

# Compute average contribution for each dimension
avg_contrib_dim1 <- mean(contributions[,1], na.rm = TRUE)
avg_contrib_dim2 <- mean(contributions[,2], na.rm = TRUE)
avg_contrib_dim3 <- mean(contributions[,3], na.rm = TRUE)
avg_contrib_dim4 <- mean(contributions[,4], na.rm = TRUE)

# Find variables that contribute more than average to each dimension
high_contrib_dim1 <- contributions[contributions[,1] > avg_contrib_dim1, 1]
high_contrib_dim2 <- contributions[contributions[,2] > avg_contrib_dim2, 2]
high_contrib_dim3 <- contributions[contributions[,3] > avg_contrib_dim3, 3]
high_contrib_dim4 <- contributions[contributions[,4] > avg_contrib_dim4, 4]

# Sort them in descending order
sorted_high_contrib_dim1 <- sort(high_contrib_dim1, decreasing = TRUE)
sorted_high_contrib_dim2 <- sort(high_contrib_dim2, decreasing = TRUE)
sorted_high_contrib_dim3 <- sort(high_contrib_dim3, decreasing = TRUE)
sorted_high_contrib_dim4 <- sort(high_contrib_dim4, decreasing = TRUE)

# Print or use for further analysis
print("Significant contributions to Dim1:")
print(sorted_high_contrib_dim1)

print("Significant contributions to Dim2:")
print(sorted_high_contrib_dim2)

print("Significant contributions to Dim3:")
print(sorted_high_contrib_dim3)

print("Significant contributions to Dim4:")
print(sorted_high_contrib_dim4)

```

```{r}
# screeplot; inspect distribution of PCs
library(factoextra)  
fviz_screeplot(res.pca)
```

```{r}
# Inspect eigenvalues
print(res.pca$eig)

```

```{r}
# Load the required package
library(kableExtra)

# Create a table with kable and kableExtra
eigen_df %>%
  kable("html") %>%
  kable_styling("striped", full_width = F) %>%
  column_spec(1, bold = T)
```

```{r}

# Rounding the numerical columns to 3 decimal places
eigen_df[, -1] <- round(eigen_df[, -1], 3)

# Replacing 'comp' with 'Dim'
eigen_df$Dimension <- gsub("comp", "Dim", eigen_df$Dimension)

# Removing row names
rownames(eigen_df) <- NULL

# Creating the kable output
kable_output <- kable(eigen_df, 
                      "html", 
                      align = 'c', 
                      col.names = c("Dimension", 
                                    "Eigenvalue", 
                                    "Variance percent", 
                                    "Cumulative variance percent")) %>%
  kable_styling("striped", 
                full_width = F) %>%
  add_header_above(c(" " = 1, "Eigenvalues and percentage of variance for each dimension of PCA" = 3))

htmltools::browsable(kable_output)

```

```{r}
# Visualize biplot:1
fviz_pca_biplot(res.pca,
                axes = c(1, 2))
rownames(res.pca$ind$coord) <- fin_summary_table_al_pca$al
# Visualize biplot with row names
fviz_pca_biplot(res.pca)

```

```{r}
# Visualize biplot: 2
fviz_pca_biplot(res.pca,
                axes = c(3, 4))
rownames(res.pca$ind$coord) <- fin_summary_table_al_pca$al
# Visualize biplot with row names
fviz_pca_biplot(res.pca)

print(fviz_pca_biplot(res.pca, axes = c(3, 4)))

```

```{r}
# BM: My preference would be to show the calculation of these 
  # values from the raw data, rather than type them in 
  # here, where we can't see where they came from, we can't easily
  # verify these values as they are currently 
##SS: fixed

# Define the volumes for each area

# add the dimensions for each area
dimensions <- data.frame(
  area = c("ta", "tp1", "d1"),
  length = c(4, 2, 1.5),
  width = c(5, 1.8, 1.5),
  thickness = c(3, 1.6, 0.95)
)

# Calculate volume for each area based on the dimensions
dimensions <- dimensions %>%
  mutate(volume = length * width * thickness)

# Count the number of artifacts by area
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Join artifact_counts with dimensions (which now includes volume) by area
lithic_density_df <- left_join(artifact_counts, dimensions, by = "area")

# Compute lithic density for each area and round it to 2 decimal places
lithic_density_df <- lithic_density_df %>%
  mutate(lithic_density = round(n / volume, 2)) # number of artifacts / volume

# Create the summary table
summary_table2_area <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics), 2)
  )

# Merge the density data with the summary table
summary_table2_area <- 
  left_join(summary_table2_area, 
            lithic_density_df[, c("area", "lithic_density")], 
            by = "area") %>%
  mutate(area = recode(area, 'ta' = 'TA', 
                       'tp1' = 'TP1', 
                       'd1' = 'D1'),
         lithic_volumetric_density = lithic_density)

# Remove the 'lithic_density' column
summary_table2_area <- summary_table2_area %>% 
  select(-lithic_density)

# Print out the resulting data frame
print(summary_table2_area)

```

```{r}
# generate table 2 - area

library(kableExtra)

kable(summary_table2_area, 
      caption = "Table 2. Lithic counts for each of the areas in Tula'i", 
      align = 'c',
      col.names = c("Area", 
                    "Number of Retouched Lithics", 
                    "Total Number of Lithics", 
                    "Retouch frequency", 
                    "Lithic Volumetric Density(pcs/(m^3)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 11) %>%
  column_spec(1, width = "1cm") %>% 
  column_spec(2, width = "1cm") %>%

  column_spec(3, width = "1cm") %>% 
  column_spec(4, width = "1cm") %>% 
  column_spec(5, width = "1cm")

```

------------------------------
Done with the main plots
------------------------------

```{r}
# WABI_LEVEL

# BM: I am confused why we are getting the counts of artefacts 
# and retouch pieces from the google sheet, after you wrote so many lines of
# code above to prepare the data from the CSV file. Where did the lithic data
# in the google sheet come from? My preference is to have 
# only copy of the lithic data in one place, and then join it with 
# other data using code. Currently it seems like we have the 
# lithic data in both the google sheet and the CSV file here. I'm worried that 
# will be confusing and be difficult to update and keep correct. I prefer that we 
# omit the lithic data from the google sheet, then import the google sheet and 
# join it to the lithic data from the CSV file. Then we can have a data frame
# with the area/layer/level data and lithic data to work on for plotting.

#SS: the googlesheet is the last thing I created. I created the summary table in R, and then for splitting up the depth (we have few depth range that was not possible to assign the layer to them, so, we splited them), I found it is easier to do so in Google sheet.

filtered_data <- sum_depth_sheet %>% filter(lithic_dens > 0 & retouch_freq > 0)

WABI_Level <- ggplot(filtered_data, 
            aes(x = lithic_dens,
                y = retouch_freq)) +
  geom_point(color = "black", size = 2) +
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              size = 0.5) +
  geom_text(aes(label = area_layer_level), 
            vjust = 1.5, 
            hjust = 0.5, 
            size = 2,
            check_overlap = TRUE) +
  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                 stat(rr.label), 
                                 stat(p.value.label), 
                                 sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 2) +
  # Add titles and labels
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +

  theme(
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

WABI_Level

```

```{r}
# whole assemblage

  # Summarize data by Area
summary_table_tech <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(area) %>% 
  summarise(
   blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    bullet_core_count = sum(`core-typology` == "bullet", na.rm = TRUE),
    flat_core_count = sum(`core-typology` == "flat", na.rm = TRUE),
    pyramid_core_count = sum(`core-typology` == "pyramid", na.rm = TRUE),
    cylinder_core_count = sum(`core-typology` %in% c("prismatic", "cylinder"), na.rm = TRUE),
    shapeless_core_count = sum(`core-typology` == "shapeless", na.rm = TRUE),

    core_fragment_count = sum(`corefrag` == "core fragment", na.rm = TRUE),
    core_rejuvention_count = sum(`corefrag` == "rejuvenation piece", na.rm = TRUE),
    core_tablet_count = sum(`corefrag` == "core tablet", na.rm = TRUE),
    core_crested_count = sum(`corefrag` == "crested bladelet", na.rm = TRUE),
    core_primarypiece_count = sum(`corefrag`  %in% c("primary flake", "primary blade"), na.rm = TRUE),

   total = n()
  )
```

```{r}
# plot_blank_type

# Data filtering and plotting
tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>% 
  filter(blank %in% c('bladelet', 'blade', 'flake')) %>%
  ggplot(aes(x = area, fill = blank)) +
  geom_bar(position = "dodge") +
  labs(
    x = "Area",
    y = "Number",
    fill = "Blank Type"
  ) +
  theme_minimal()

```

```{r}
#barplot_layer

tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d")) %>%
  filter(!is.na(corefrag)) %>% 
  select(layer, corefrag) %>%
  ggplot() +
  aes(x = corefrag) +
  geom_bar() +  
  labs(x = "Core and Core fragments", y = "Frequency") +
  facet_wrap(~layer, ncol = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#Core-barplot_area

tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  filter(!is.na(corefrag)) %>% # Exclude rows where corefrag is NA
  select(area, corefrag) %>%
  ggplot() +
  aes(x = corefrag) +
  geom_bar() +
  labs(x = "Core and Core fragments",
       y = "Frequency") +
  facet_wrap(~area, ncol = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}

filtered_data <- tl_final %>%
  filter(!is.na(`tool-typology`) & `tool-typology` != "" & `tool-typology` != "NA")
ggplot(data = filtered_data) + 
  aes(y = `tool-typology`) + 
  geom_bar() +
  labs(x = "Frequency",
       y = "Tool Type") +
  theme_minimal() +
  theme(axis.text.y = element_text(angle = 20, hjust = 1, vjust = 0.5))

```

PCA-LAYER

```{r}
# prepare the data for PCA-layer

# Calculating Blank Frequency
blank_freq <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(blank)) %>%
  group_by(layer) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Calculating Core Frequency
core_freq <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(corefrag)) %>%
  group_by(layer) %>%
  summarise(coreFreq = sum(corefrag %in% c("core fragment", "bullet", "flat", "pyramid", "cylinder/prismatic", "core rejuvenation piece", "shapeless", "core on flake", "bipolar")) / sum(!is.na(corefrag)), .groups = 'drop')

# Calculating Tool Diversity using Menhinick's Index
tool_div <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(tooltypegroup)) %>%
  group_by(layer) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(sum(!is.na(tooltypegroup))), .groups = 'drop')

# Combining all the calculated frequencies into one tibble
final_table <- blank_freq %>%
  inner_join(core_freq, by = "layer") %>%
  inner_join(tool_div, by = "layer")

# Printing the final table
final_table

```

```{r}
# Blank Frequency
blank_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d"), !is.na(blank)) %>%
  group_by(layer) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Calculating Core Frequency
core_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "e or d")) %>%
  group_by(layer) %>%
  summarise(coreFreq = sum(corefrag %in% c("core fragment", "bullet", "flat", "pyramid", "cylinder/prismatic", "core rejuvenation piece", "shapeless", "core on flake", "bipolar")) / n(), .groups = 'drop')

# Calculating Tool Diversity using Menhinick's Index
tool_div <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "e or d"), !is.na(tooltypegroup)) %>%
  group_by(layer) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(sum(!is.na(tooltypegroup))), .groups = 'drop')

#------------------------------------------------------------------------------
# BM: I wasn't able to run any code below this line
##SS: fixed
# -----------------------------------------------------------------------------

# Calculating Retouch Frequency
retouch_freq <- tl_final %>%
  filter(layer %in% c("a", "c", "d", "e", "eor d")) %>%
  group_by(layer) %>%
  summarise(retFreq = sum(retouch == 1, na.rm = TRUE) / n(), .groups = 'drop')

# Combining all the calculated frequencies, including retouch frequency, into one tibble
final_table <- blank_freq %>%
  inner_join(core_freq, by = "layer") %>%
  inner_join(tool_div, by = "layer") %>%
  inner_join(retouch_freq, by = "layer")

final_table

```

```{r}
# Joining calculated frequencies
final_table <- blank_freq %>%
  inner_join(core_freq, by = "layer") %>%
  inner_join(tool_div, by = "layer") %>%
  inner_join(retouch_freq, by = "layer")

# Converting layer to row names
final_table_rownames <- final_table %>%
  column_to_rownames(var = "layer")

```

```{r}
# Compute the PCA
library(FactoMineR)

res_pca <-
  PCA(final_table_rownames,
      graph = FALSE)

```

```{r}
library(factoextra)

# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca,
             axes = 1,   
             choice = "var")

```

```{r}
library(factoextra)

# contributions of variables to the second principal component
fviz_contrib(res_pca,
             axes = 2,   
             choice = "var")

```

```{r}
library(factoextra)

#  biplot
fviz_pca_biplot(res_pca)
```

```{r}
library(factoextra)
fviz_screeplot(res_pca)
```

```{r}
fviz_pca_var(res_pca)
```

```{r}

library(FactoMineR)
library(factoextra)

# Run PCA
res_pca <- PCA(final_table_rownames, graph = FALSE)

# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca, axes = 1, choice = "var")

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca, axes = 2, choice = "var")

# Visualize eigenvalues and explained variance
fviz_screeplot(res_pca)

# Create a biplot of the PCA
fviz_pca_biplot(res_pca)

# Visualize the variables on the principal components
fviz_pca_var(res_pca)

```

```{r}
# Run PCA
library(FactoMineR)
library(factoextra)

res_pca <- PCA(final_table_rownames, graph = FALSE)

# Get the coefficients of original variables for each principal component
loadings_matrix <- res_pca$var$coord

print(loadings_matrix)

```

```{r}
#Table

# Eigenvalues
eigenvalues <- res_pca$eig[1:2, 1]

# Variance percent
variance_percent <- res_pca$eig[1:2, 2]

# Cumulative variance percent
cumulative_variance_percent <- res_pca$eig[1:2, 3]

# Create the table
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)

rownames(df) <- c("Dim1", "Dim2")

print(df)

```

```{r}
#Table

library(gt)

# Create data frame
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)
rownames(df) <- c("Dim1", "Dim2")

# Create the table using gt
gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = vars(Eigenvalue, Variance_Percent, Cumulative_Variance_Percent),
    decimals = 2
  )

```

feature frequency: d1 1 ta 1 tp1 0)

```{r}
library(dplyr)
library(FactoMineR)
library(tibble)

# Define estimated area and feature frequency data
area_data <- tibble(
  area = c("d1", "ta", "tp1"),
  feature_freq = c(1, 1, 0)
)

# Step 1: Number of artifacts by area
artifact_counts <- tl_final_area %>%
  group_by(area) %>%
  tally()

# Step 2: Calculate each variable

# Retouch Frequency 
retouch_freq <- tl_final_area %>%
  group_by(area) %>%
  summarise(retouchFreq = mean(!is.na(retouch)), .groups = 'drop')

# Blank Frequency
blank_freq <- tl_final_area %>%
  group_by(area) %>%
  summarise(blankFreq = mean(!is.na(blank)), .groups = 'drop')

# Core Frequency
core_freq <- tl_final_area %>%
  group_by(area) %>%
  summarise(coreFreq = mean(!is.na(`core-typology`)),.groups = 'drop')

# Lithic density (Artifact counts / estimated area)
lithic_density <- artifact_counts %>%
  left_join(area_data, by = "area") %>%
  mutate(lithic_density = n / estimated_area)

# Tool Diversity using Menhinick's Index
tool_div <- tl_final_area %>%
  group_by(area) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(n()), .groups = 'drop')

# Step 3: Combining all variables

pca_final_table_area <- left_join(dimensions, retouch_freq, by = "area") %>%
  left_join(blank_freq, by = "area") %>%
  left_join(core_freq, by = "area") %>%
  left_join(lithic_density, by = "area") %>%
  left_join(tool_div, by = "area") %>%
  select(area, new_dimensions, feature_freq, retouchFreq, blankFreq, coreFreq, lithic_density, toolDiv)

```

```{r}
pca_final_table_area <- pca_final_table_area %>%
    select(
        area, 
        estimated_area = estimated_area.x, 
        feature_freq = feature_freq.x, 
        retouchFreq, 
        blankFreq, 
        coreFreq, 
        n, 
        lithic_density, 
        toolDiv
    )

```

```{r}
# Step 4: Perform PCA
res_pca_area <- PCA(pca_final_table_area %>% select(-area, -n), graph = FALSE)

# Show the summary of PCA
print(res_pca_area)

print(res_pca_area$eig)
print(res_pca_area$var$coord)
```

```{r}
# TABLE: Eigenvalues and percentage of variance for each dimension of PCA 
library(kableExtra)

# Define the values manually (you would get these from your PCA results)
# BM: I prefer not to write values in the code like this. I think
# this is bad practice and leads to mistakes
eigenvalues <- c(4.411234, 2.588766)
prop_var_explained <- c(63.02, 36.98)
cumulative_var_explained <- c(63.02, 100.00)

# Create the data frame
results_df <- data.frame(
  Dimension = c("Dim. 1", "Dim. 2"),  # Only keep the dimensions you want to display
  Eigenvalue = eigenvalues,
  `Variance / percent` = prop_var_explained,
  `Cumulative / variance percent` = cumulative_var_explained
)

# Generate the table
kable(results_df,
      caption = "Table 3. Eigenvalues and percentage of variance for each dimension of PCA",
      align = 'c',
      col.names = c("Dimension", "Eigenvalue", "Variance (%)", "Cumulative Variance (%)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 11)

```

```{r}
#plotting Dim-1 and Dim-2
library(FactoMineR)
library(factoextra)

# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca_area, axes = 1, choice = "var")

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca_area, axes = 2, choice = "var")

```

```{r}

# having Dim1 and Dim2 plots next to each other

library(gridExtra)

# Create the first plot and assign it to p1
p1 <- fviz_contrib(res_pca_area, axes = 1, choice = "var")

# Create the second plot and assign it to p2
p2 <- fviz_contrib(res_pca_area, axes = 2, choice = "var")

# Use grid.arrange to place them next to each other
grid.arrange(p1, p2, ncol = 2)

```

```{r}

loadings <- res_pca_area$var$coord

# 
eigenvalues <- res_pca_area$eig

#  Dim.1
contrib_dim1 <- (loadings[,1]^2 / eigenvalues[1, 1]) * 100

#  Dim.2
contrib_dim2 <- (loadings[,2]^2 / eigenvalues[2, 1]) * 100

contrib_df <- data.frame(
  Variable = rownames(loadings),
  Contrib_Dim1 = contrib_dim1,
  Contrib_Dim2 = contrib_dim2
)

print(contrib_df)

```

```{r}
# screeplot

library(FactoMineR)
library(factoextra)

fviz_eig(res_pca_area)

```

```{r}
# Correlation plots of variables for each of the tow relevant PCA dimensions

fviz_pca_var(res_pca_area,

             repel = TRUE) + 
  theme(text = element_text(size = 10))

```

```{r}
# Eigenvalues and percentage of variance
eig.val <- res_pca_area$eig
print(eig.val)

# Variable coordinates (correlations with principal components)
var.coord <- res_pca_area$var$coord
print(var.coord)

```

```{r}
#PCA biplot for Dimensions 1 and 2

fviz_pca_biplot(res_pca_area, labelsize = 3, repel = TRUE)

```

```{r}
# biplot 
p <- fviz_pca_biplot(res_pca_area, label = "none", 
                     repel = TRUE, labelsize = 4)

p + geom_text(aes(x = res_pca_area$ind$coord[, 1], 
                  y = res_pca_area$ind$coord[, 2], 
                  label = pca_final_table_area$area), 
             size = 3, nudge_x = 0.1, nudge_y = 0.1)

```

```{r}
individual_coordinates = res_pca_area$ind$coord
variable_coordinates = res_pca_area$var$coord

```

```{r}
# For variables
fviz_cos2(res_pca_area, choice = "var")

# For individuals
fviz_cos2(res_pca_area, choice = "ind")

```

DONE WITH PCA

```{r}

individual_coordinates <- res_pca_area$ind$coord

individual_coordinates_df <- as.data.frame(individual_coordinates)
individual_coordinates_df$area <- unique(area_data$area)

print(individual_coordinates_df)

```

```{r}
library(kableExtra)

# Extract individual coordinates
individual_coordinates <- res_pca_area$ind$coord

# Convert to a data frame for better visibility and add the 'area' column back for context
individual_coordinates_df <- as.data.frame(individual_coordinates)
individual_coordinates_df$Area <- unique(area_data$area)  # Make sure the order matches

# Use kable to create a beautiful table
kable(individual_coordinates_df, 
      col.names = c("Area", "Principal Component 1", "Principal Component 2"), 
      caption = "PCA Coordinates for the Archaeological Areas excavation",
      align = c('l', 'c', 'c'),
      digits = 3) %>%
  kable_styling("striped", full_width = F)

```

```{r}
# Get the loadings (coefficients) of original variables for each principal component
loadings_matrix <- res_pca$var$coord

```

```{r}
# Eigenvalues
eigenvalues <- res_pca$eig[1:2, 1]
variance_percent <- res_pca$eig[1:2, 2]
cumulative_variance_percent <- res_pca$eig[1:2, 3]

# Create the table
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)

rownames(df) <- c("Dim1", "Dim2")

# Print the table
print(df)

```

```{r}
library(gt)

gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

```

```{r}

artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Print the resulting table
print(artifact_counts)
```

```{r}

blank_freq <- tl_final_area %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(blank)) %>%
  group_by(area) %>%
  summarise(blankFreq = sum(blank %in% c("bladelet", "blade", "flake")) / sum(!is.na(blank)), .groups = 'drop')

# Calculating Core Frequency
core_freq <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(corefrag)) %>%
  group_by(area) %>%
  summarise(coreFreq = sum(corefrag %in% c("core fragment", "bullet", "flat", "pyramid", "cylinder/prismatic", "core rejuvenation piece", "shapeless", "core on flake", "bipolar")) / sum(!is.na(corefrag)), .groups = 'drop')

# Calculating Tool Diversity using Menhinick's Index
tool_div <- tl_final_area %>%
  filter(area %in% c("ta", "d1", "tp1"), !is.na(tooltypegroup)) %>%
  group_by(area) %>%
  summarise(toolDiv = length(unique(tooltypegroup)) / sqrt(sum(!is.na(tooltypegroup))), .groups = 'drop')

# Combining all the calculated frequencies into one tibble
final_table <- blank_freq %>%
  inner_join(core_freq, by = "area") %>%
  inner_join(tool_div, by = "area")

# Printing the final table
print(final_table)

```

```{r}
# Joining calculated frequencies
final_table <- blank_freq %>%
  inner_join(core_freq, by = "area") %>%
  inner_join(tool_div, by = "area") %>%
  inner_join(retouch_freq, by = "area")

# Converting area to row names
final_table_rownames <- final_table %>%
  column_to_rownames(var = "area")

```

```{r}

library(FactoMineR)
library(factoextra)

# Run PCA
res_pca <- PCA(final_table_rownames, graph = FALSE)

# Visualize the contributions of variables to the second principal component
fviz_contrib(res_pca, axes = 2, choice = "var")

```

```{r}
# Visualize the contributions of variables to the first principal component
fviz_contrib(res_pca, axes = 1, choice = "var") 
```

```{r}
# Visualize eigenvalues and explained variance
fviz_screeplot(res_pca)
```

```{r}
fviz_pca_biplot(res_pca, labelsize = 3, repel = TRUE)

```

```{r}
# Visualize the variables on the principal components
fviz_pca_var(res_pca)
```

```{r}
# Get the loadings (coefficients) of original variables for each principal component
loadings_matrix <- res_pca$var$coord

# Print the loadings matrix
print(loadings_matrix)

```

```{r}
# Eigenvalues
eigenvalues <- res_pca$eig[1:2, 1]
variance_percent <- res_pca$eig[1:2, 2]
cumulative_variance_percent <- res_pca$eig[1:2, 3]

# Create the table
df <- data.frame(Eigenvalue = eigenvalues,
                 Variance_Percent = variance_percent,
                 Cumulative_Variance_Percent = cumulative_variance_percent)

rownames(df) <- c("Dim1", "Dim2")

# Print the table
print(df)

```

```{r}
library(gt)

# Create the table 
gt(df) %>%
  tab_header(
    title = "",
    subtitle = "Eigenvalues and percentage of variance for each dimension of PCA"
  ) %>%
  cols_label(
    Eigenvalue = "Eigenvalue",
    Variance_Percent = "Variance (%)",
    Cumulative_Variance_Percent = "Cumulative Variance (%)"
  ) %>%
  fmt_number(
    columns = c("Eigenvalue", "Variance_Percent", "Cumulative_Variance_Percent"),
    decimals = 2
  )

```

```{r}
# Count the number of artifacts by area
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Print the resulting table
print(artifact_counts)

```

```{r}

volumes <- data.frame(
  area = c("ta", "tp1", "d1"),
  volume = c(52, 5.76, 2.1375)
)

# Count the number of artifacts by area
artifact_counts <- tl_final %>%
  filter(area %in% c("ta", "tp1", "d1")) %>%
  group_by(area) %>%
  tally()

# Join artifact_counts with volumes by area
lithic_density_df <- left_join(artifact_counts, volumes, by = "area")

# Compute lithic density for each area
lithic_density_df <- lithic_density_df %>%
  mutate(lithic_density = n / volume)

# Print the resulting table
print(lithic_density_df[, c("area", "lithic_density")])

```

```{r}
# Counting the number of cores by area
core_counts <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1") & !is.na(`core and core fragments`)) %>%
  group_by(area) %>%
  tally()

print(core_counts)

```