### 2023-09-21 paper/paper_tulai.qmd
added grouped_level column. edited depths and chronological units.

---
title: ""
author: ""
date: now
format: 
  html:
    self-contained: true
execute: 
  warning: false
  fig-dpi: 600
title-block-published: "last updated"
editor: visual
---

```{r}
# load libraries
library(tidyverse)
library(here)
library(ggpmisc)
library(ggrepel)
library(ggbeeswarm)
library(cowplot)
library(knitr)
```

Cleaning Data:

```{r}
# Read the data
tl <- read_csv(here("data/Tulai_Lithic_Assamblage.csv"))

# Create a new data frame with renamed columns
tl1 <- tl
names(tl1)[1:2] <- c("lithic_id", 
                     "TA")

# modify TA column
tl1 <- tl1 %>%
  # Remove everything within brackets
  mutate(TA = str_remove_all(TA, 
                             "\\(.*\\)")) %>%

  # Add slash between TA and numbers
  mutate(TA = str_replace_all(TA, 
                              "TA([0-9]+)", 
                              "TA/\\1")) %>% 
  # Add slash between numbers
  mutate(TA = str_replace_all(TA, 
                              "([0-9]+)/([0-9]+)", "\\1/\\2")) %>% 
 # fill NA values in TA column
  tidyr::fill(TA, .direction = "down") %>%
  # Prepend 'TA/' to numbers
  mutate(TA = if_else(str_detect(TA, 
                                 "^[0-9]+$"), 
                      paste0("TA/",
                             TA), 
                      TA)) %>% 
  # Remove all non-alphanumeric characters
  mutate(TA = str_replace_all(TA, 
                              "[^[:alnum:] /]", 
                              "")) %>% 
  # Make all text lowercase
  mutate(TA = tolower(TA)) %>% 
  # Remove white space
  mutate(TA = str_replace_all(TA, " ", ""))  %>%
  # Separate 'TA' column into 'area' and 'depth' columns
  separate(TA, into = c("area", 
                        "depth"), 
           sep = "/",
           remove = FALSE,
           convert = TRUE) %>%
  # Convert 'depth' to numeric
  mutate(depth = as.numeric(depth)) %>% 
  # Create depth ranges
  mutate(depth = case_when(
             depth == 3 ~ "0-30 cm",
             depth == 34 ~ "30-40 cm",
             depth == 45 ~ "40-50 cm",
             depth == 56 ~ "50-60 cm",
             depth == 67 ~ "60-70 cm",
             depth == 72 ~ "70-120 cm",
             depth == 125 ~ "120-150 cm",
             depth == 158 ~ "150-180 cm",
             depth == 189 ~ "180-190 cm",
             depth == 190 ~ "190-200 cm",
             depth == 201 ~ "200-210 cm",
             depth == 212 ~ "210-220 cm",
             depth == 223 ~ "220-230 cm",
             depth == 234 ~ "230-240 cm",
             depth == 235 ~ "230-250 cm",
             depth == 256 ~ "250-260 cm",
             depth == 267 ~ "260-270 cm",
             depth == 72 ~ "70-120 cm",
             depth == 1 ~ "0-10 cm",
             depth == 2 ~ "0-20 cm", 
             depth == 13 ~ "10-30 cm",
             depth == 19 ~ "10-90 cm",
             depth == 23 ~ "20-30 cm",
             depth == 34 ~ "30-40 cm",
             depth == 45 ~ "40-50 cm",
             depth == 51 ~ "50-100 cm",
             depth == 56 ~ "50-60 cm",
             depth == 57 ~ "50-70 cm",
             depth == 89 ~ "80-90 cm",
             depth == 91 ~ "90-100 cm",
             depth == 101 ~ "100-110 cm",
             depth == 112 ~ "110-120 cm",
             depth == 123 ~ "120-130 cm",
             depth == 124 ~ "120-140 cm",
             depth == 23 ~ "20-30 cm",
             depth == 46 ~ "40-60 cm",
             depth == 78 ~ "70-80 cm",
             depth == 235 ~ "230-250 cm",
             depth == 12 ~ "10-20 cm",
             depth == 90 ~ "90-100 cm",

             TRUE ~ as.character(depth)
  )) %>%

  separate(depth, 
           into = c("upper",
                    "lower"),
           sep = "-",
           remove = FALSE,
           convert = TRUE) %>% 
  mutate(lower = parse_number(lower)) %>% 
  rowwise() %>% 
  mutate(midpoint = mean(c(upper, lower))) %>%
# create 'chronological unit' column
  mutate(chronological_unit = case_when(
      area == "tp1" & depth == "0-10 cm" ~ "c",
      area == "tp1" & depth == "0-20 cm" ~ "c",
      area == "tp1" & depth == "10-30 cm" ~ "c",
      area == "tp1" & depth == "10-90 cm" ~ "c",
      area == "tp1" & depth == "20-30 cm" ~ "c",
      area == "tp1" & depth == "30-40 cm" ~ "c",
      area == "tp1" & depth == "40-50 cm" ~ "c",
      area == "tp1" & depth == "50-100 cm" ~ "c",
      area == "tp1" & depth == "50-60 cm" ~ "b",
      area == "tp1" & depth == "50-70 cm" ~ "b",
      area == "tp1" & depth == "80-90 cm" ~ "b",
      area == "tp1" & depth == "90-100 cm" ~ "b",
      area == "tp1" & depth == "100-110 cm" ~ "b",
      area == "tp1" & depth == "110-120 cm" ~ "b",
      area == "tp1" & depth == "120-130 cm" ~ "b",
      area == "tp1" & depth == "120-140 cm" ~ "b",
      area == "d1" & depth == "0-20 cm" ~ "c",
      area == "d1" & depth == "20-30 cm" ~ "c",
      area == "d1" & depth == "40-60 cm" ~ "a",
      area == "d1" & depth == "50-60 cm" ~ "a",
      area == "d1" & depth == "60-70 cm" ~ "a",
      area == "d1" & depth == "70-80 cm" ~ "a",
      area == "ta" & depth == "0-30 cm" ~ "e",
      area == "ta" & depth == "30-40 cm" ~ "e",
      area == "ta" & depth == "40-50 cm" ~ "e",
      area == "ta" & depth == "50-60 cm" ~ "e",
      area == "ta" & depth == "60-70 cm" ~ "e",
      area == "ta" & depth == "70-120 cm" ~ "d",
      area == "ta" & depth == "120-150 cm" ~ "d",
      area == "ta" & depth == "150-180 cm" ~ "d",
      area == "ta" & depth == "180-190 cm" ~ "d",
      area == "ta" & depth == "190-200 cm" ~ "d",
      area == "ta" & depth == "200-210 cm" ~ "d",
      area == "ta" & depth == "210-220 cm" ~ "d",
      area == "ta" & depth == "220-230 cm" ~ "d",
      area == "ta" & depth == "230-240 cm" ~ "d",
      area == "ta" & depth == "230-250 cm" ~ "d",
      area == "ta" & depth == "250-260 cm" ~ "d",
      area == "ta" & depth == "260-270 cm" ~ "d",

      TRUE ~ NA_character_
  )) %>%

  mutate(
 grouped_level = case_when(
    area == "tp1" & depth == "0-10 cm" ~ "3",
      area == "tp1" & depth == "0-20 cm" ~ "3",
      area == "tp1" & depth == "10-30 cm" ~ "3",
      area == "tp1" & depth == "10-90 cm" ~ "3",
      area == "tp1" & depth == "20-30 cm" ~ "3",
      area == "tp1" & depth == "30-40 cm" ~ "3",
      area == "tp1" & depth == "40-50 cm" ~ "3",
      area == "tp1" & depth == "50-100 cm" ~ "3",
      area == "tp1" & depth == "50-60 cm" ~ "2",
      area == "tp1" & depth == "50-70 cm" ~ "2",
      area == "tp1" & depth == "80-90 cm" ~ "2",
      area == "tp1" & depth == "90-100 cm" ~ "2",
      area == "tp1" & depth == "100-110 cm" ~ "1",
      area == "tp1" & depth == "110-120 cm" ~ "1",
      area == "tp1" & depth == "120-130 cm" ~ "1",
      area == "tp1" & depth == "120-140 cm" ~ "1",
      area == "d1" & depth == "0-20 cm" ~ "3",
      area == "d1" & depth == "20-30 cm" ~ "3",
      area == "d1" & depth == "40-60 cm" ~ "2",
      area == "d1" & depth == "50-60 cm" ~ "2",
      area == "d1" & depth == "60-70 cm" ~ "1",
      area == "d1" & depth == "70-80 cm" ~ "1",
      area == "ta" & depth == "0-30 cm" ~ "6",
      area == "ta" & depth == "30-40 cm" ~ "6",
      area == "ta" & depth == "40-50 cm" ~ "5",
      area == "ta" & depth == "50-60 cm" ~ "5",
      area == "ta" & depth == "60-70 cm" ~ "5",
      area == "ta" & depth == "70-120 cm" ~ "4",
      area == "ta" & depth == "120-150 cm" ~ "4",
      area == "ta" & depth == "150-180 cm" ~ "3",
      area == "ta" & depth == "180-190 cm" ~ "3",
      area == "ta" & depth == "190-200 cm" ~ "2",
      area == "ta" & depth == "200-210 cm" ~ "2",
      area == "ta" & depth == "210-220 cm" ~ "2",
      area == "ta" & depth == "220-230 cm" ~ "2",
      area == "ta" & depth == "230-240 cm" ~ "2",
      area == "ta" & depth == "230-250 cm" ~ "1",
      area == "ta" & depth == "250-260 cm" ~ "1",
      area == "ta" & depth == "260-270 cm" ~ "1",

TRUE ~ NA_character_
  )
)

# Remove specific columns
tl1 %>% 
  select(-c("AREA", 
            "DEPTH (cm)", 
            "Raw Material", 
            "Colour/Grain/Opacity", 
            "Pattern"))
```

```{r}
# exploring Data

p1 <- 
tl1 %>% 
  group_by(midpoint) %>% 
  tally() %>% 
  drop_na(midpoint) %>% 
ggplot() + 
  aes(midpoint, n) +
  geom_col()

p2 <- 
tl1 %>% 
  group_by(chronological_unit) %>% 
  tally() %>% 
  drop_na(chronological_unit) %>% 
ggplot() + 
  aes(chronological_unit, n) +
  geom_col()

plot_grid(p1, p2, ncol = 1)
```

```{r}
# exploring Data

tl2 <- tl1 %>%
  mutate(
    Blank = case_when(
      Breakage %in% c("0", "1", NA_character_) ~ Blank, 
      TRUE ~ Breakage  
    )
  )
```

```{r}
# Cleaning Blank column

tl2 <- tl2 %>%
  mutate(
    Blank = str_to_lower(Blank),
    Blank = str_trim(Blank),
    Blank = str_replace_all(Blank, "[./]", "-"),
    Blank = str_replace_all(Blank, "\\s*-\\s*", "-")
  ) %>%
  mutate(Blank = case_when(
             Blank == "peo" ~                      "bladelet-pro",
             Blank == "pro" ~                      "bladelet-pro",
             Blank == "bladelet-dis" ~             "bladelet-dis",
             Blank == "microblade" ~               "bladelet-complete",
             Blank == "microblade-pro" ~           "bladelet-pro",
             Blank == "microblade-med" ~           "bladelet-med",
             Blank == "micrpblade-dis" ~           "bladelet-dis",
             Blank == "dis" ~                      "bladelet-dis",
             Blank == "bladelrt-med" ~             "bladelet-med",
             Blank == "bladelet-mes" ~             "bladelet-med",
             Blank == "bladelet-,ed" ~             "bladelet-med",
             Blank == "bladelet-bladelet-med" ~    "bladelet-med",
             Blank == "medial" ~                   "bladelet-med",
             Blank == "bladelet-bladelet-pro" ~    "bladelet-pro",
             Blank == "bladelert-pro" ~            "bladelet-pro",
             Blank == "microblade-medial" ~        "bladelet-med",
             Blank == "nicroblade-pro" ~           "bladelet-pro",
             Blank == "microblade-?" ~             "bladelet-complete",
             Blank == "med" ~                      "bladelet-med",
             Blank == "bladeler-pro" ~             "bladelet-pro",
             Blank == "microblade-dis" ~           "bladelet-dis",
             Blank == "microblde-dis" ~            "bladelet-dis",
             Blank == "bladlet-pro" ~              "bladelet-pro",
             Blank == "indistinct" ~               "bladelet-med",
             Blank == "flake" ~                    "flake-complete",
             Blank == "dladelet-pro" ~             "bladelet-pro",
             Blank == "dladelet-pro" ~             "bladelet-pro",
             Blank == "bldelet-bladelet-pro" ~     "bladelet-pro",
             Blank == "bldelet-pro" ~              "bladelet-pro",
             Blank == "nicroblade-pro" ~           "bladelet-pro",
             Blank == "microblade-?" ~             "bladelet-med",
             Blank == "bladelt-pro" ~              "bladelet-pro",
             Blank == "bladeket-med" ~             "bladelet-med",
             Blank == "nicroblade-pro" ~           "bladelet-pro",
             Blank == "blaelet" ~                  "bladelet-complete",
             Blank == "bladele-med" ~              "bladelet-med",
             Blank == "bladelet-nearly complete" ~ "bladelet-complete",
             Blank == "blade-nearly complete" ~    "blade-complete",
             Blank == "bladelet-?" ~               "bladelet-complete",
             Blank == "flake?" ~                   "flake-complete",
             Blank == "micrpblade-dis" ~           "bladelet-dis",
             Blank == "blaelet-med" ~              "bladelet-med",
             Blank == "blaedlet-pro" ~             "bladelet-pro",  
             Blank == "bladelt-pro" ~              "bladelet-pro",   
             Blank == "bladelet-nearly complete"~  "bladelet-complete", 
             Blank == "microblade-med" ~           "bladelet-med", 
             Blank == "bldelet-bladelet-pro" ~     "bladelet-pro",
             Blank == "bldelet-flake" ~            "blade-complete",
             Blank == "mwdial" ~                   "bladelet-med",
             Blank == "blade" ~                    "blade-complete",
             Blank == "bladelt" ~                  "bladelet-complete",
             Blank == "bladlet" ~                  "bladelet-complete",
             Blank == "bladlet-med" ~              "bladelet-med",
             Blank == "bladelt-med" ~              "bladelet-med",
             Blank == "flaje" ~                    "flake",
             Blank == "flke" ~                     "flake",
             Blank == "bladelet" ~                 "bladelet-complete",
             Blank == "NA" ~                       "bladelet-",
             Blank == "thick flake" ~              "flake",
             Blank == "flke-mid" ~                 "flake-med",
             Blank == "flake-blade" ~              "flake",
             Blank == "angular flake?" ~           "flake",
             Blank == "flake-prox" ~               "flake-pro",
             Blank == "fklae-dis" ~                "flake-dis",
             Blank == "bladelete" ~                "bladelet",
             Blank == "NA" ~                    NA_character_,
    TRUE ~ Blank
  )) %>%
  mutate(Blank = na_if(Blank, ""))

# rev(sort(table(tl2$Blank)))

# BM: still many typos and unusual categories in here, why is that? Can they be fixed? This might increase our sample size.

tl3 <- tl2 %>%
  separate(Blank, 
           into = c("Blank2", "Blank_part"),
           sep = "-",
           remove = FALSE,
           convert = TRUE)

na_rows <- which(is.na(tl2$Blank))

zero_rows <- which(tl2$Blank == "0")
one_rows <- which(tl2$Blank == "1")

# rev(sort(table(tl3$Blank2)))

```

```{r}
# BM: this is an unusual coding pattern, I'm curious about where you got it from?

tl3 <- tl3 %>%
  mutate(
    Blank2 = case_when(
      Blank2 == "flake" ~ "flake",  # Keep "flake" as it is
      (`Core Typology` %in% c("0", "NA", "na", "-", ".") | 
       is.na(`Core Typology`) | 
       `Core Typology` == "") & `Width (mm)` < 12 ~ "bladelet",
      (`Core Typology` %in% c("0", "NA", "na", "-", ".") | 
       is.na(`Core Typology`) | 
       `Core Typology` == "") & `Width (mm)` >= 12 ~ "blade",
      TRUE ~ Blank2
    )
  )  %>%
  mutate(
    Blank2 = case_when(
      Blank2 %in% c("blade", "bladelet", "flake") ~ Blank2,
      TRUE ~ NA_character_
    )
  )

```

```{r}
# SS: they are natural rocks, and don't want to count them as lithic
# BM: better to remove them by their ID than by row number, since row number can easily change, but ID cannot

tl3 <- tl3 %>% 
  filter(!lithic_id %in% c("3154", "1741", "1680", "طبیعی"))
```

```{r}
tl4 <- tl3 %>%
  mutate(
    Utilization = if_else(Retouch == "1", 
                          NA_character_, 
                          as.character(Utilization))
  )

```

```{r}

tl4 <- tl4 %>%
  # BM: I updated this to use across
  mutate(across(everything(), 
                ~ifelse(is.na(.x)|.x == ""|.x == "-", NA, .x))) %>% 
  #SS:  remove any asterisks, whitespace, or *, ?, or !. is that wrong?
   mutate(across(everything(),
                 ~str_replace(., "(?<=\\D)\\*(?=\\D)|\\s+|[*?!]+", ""))) %>%
  ##SS: we meant sickle shine, this is one of the criteria based on which Hole claimed the site is not related to agriculture (similar to ALi Kosh located in Dehluran)
  mutate(
    SHINE = str_trim(SHINE) %>% 
            str_to_title(),
    shine2 = SHINE
  ) %>%
  mutate(
    shine2 = case_when(
      SHINE %in% c("Subparallel", 
                   "Parallel", 
                   "Sub-Parallel", 
                   "Parallel-Subprarallel",
                   "Scaled", 
                   "Sub-Paralel", 
                   "Sub-Parallell", 
                   "Sub-Paallel",
                   "Sub-Paralell", 
                   "Sub-Prallel", 
                   "L:semiparallel/R:scaled",
                   "L:semiparallel/R: Scaled",
                   "Semiparallel", 
                   "Semi-Parallel", 
                   "Scaled/Sub-Parallel",
                   "Seb-Parallel", 
                   "Semi-Abrupt") ~ NA_character_,
      TRUE ~ SHINE
    ),
    Utilization = if_else(Utilization == "0", NA_character_, Utilization),
    Retouch =     if_else(Retouch == "0", NA_character_, Retouch),
    SHINE =       if_else(SHINE == "0", NA_character_, SHINE),
    shine2 =      if_else(shine2 == "0", NA_character_, shine2),
    Blank2 =      if_else(Blank2 == "0", NA_character_, Blank2)
  ) %>% 
  mutate(
    SHINE = na_if(SHINE, "")
  )

# BM: I think this is a better way to inspect:
# rev(sort(table(tl4$SHINE)))
# rev(sort(table(tl4$shine2)))

```

```{r}

tl4 <- tl4 %>%
  mutate(
    `Core Typology` = str_to_lower(`Core Typology`),
    `Core Typology` = str_trim(`Core Typology`),
    # BM: what does this regex do in the next line?
    `Core Typology` = str_replace_all(`Core Typology`, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(`Core Typology` = case_when(
    `Core Typology` == "flatcore" ~ "flat-pressure",
             `Core Typology` == "rejuvention" ~ "rejuvenation piece-NA",
             `Core Typology` == "pressureprymidal core" ~ "pyramid-pressure",
             `Core Typology` == "pressureflat core" ~ "flat-pressure",
             `Core Typology` == "pressurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prussurebullet core" ~ "bullet-pressure",
             `Core Typology` == "prismaticcore" ~ "prismatic-percussion",
             `Core Typology` == "pyramidalpressure core" ~ "pyramid-pressure",
             `Core Typology` == "flatcore with one debitage surface/pressure" ~ "flat-pressure",
             `Core Typology` == "corefragment" ~ "core fragment-NA",
             `Core Typology` == "cilandrical/pressure" ~ "cylinder-pressure",
             `Core Typology` == "bullet/pressure" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/pressure" ~ "pyramid-pressure",
             `Core Typology` == "percussion/pyramidal" ~ "pyramid-pressure",
             `Core Typology` == "flat/pressure" ~ "flat-pressure",
             `Core Typology` == "pyramidal/ percussion??" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ peressure" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/ pressure" ~ "pyramid-pressure",
             `Core Typology` == "pressure" ~ "pyramid-pressure",
             `Core Typology` == "multidirectional/percussion" ~ "shapeless-percussion",
             `Core Typology` == "prismatic/ percussion" ~ "pyramid-pressure",
             `Core Typology` == "flatcore/ unidirectional/pressure" ~ "flat-pressure",
             `Core Typology` == "heavilyused" ~ "NA",
             `Core Typology` == "burnt" ~ "NA",
             `Core Typology` == "patinated" ~ "NA",
             `Core Typology` == "pyramidal(bullet)/pressure" ~ "bullet-pressure",
             `Core Typology` == "cylandrycal/ bidirectional pressure core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure bullet core" ~ "bullet-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "pyramidal/unidirectional pressure  core" ~ "pyramid-pressure",
             `Core Typology` == "bladelet" ~ "bullet-pressure",
             `Core Typology` == "lip" ~ "NA",
             `Core Typology` == "bladlet" ~ "bullet-pressure",
             `Core Typology` == "bladelet(30.56)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(28.55.6)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(16.28.1)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(40 5.7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(23.2 2.3)" ~ "bullet-pressure",
             `Core Typology` == "bladelet-flakeblade(41 16)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(28 6.7)" ~ "bullet-pressure",
             `Core Typology` == "bladelet(3512.6/7)" ~ "pyramid-pressure",
             `Core Typology` == "bladelet(27.307.5)" ~ "bullet-pressure",
    `Core Typology` == "pressure/pyramidal" ~ "pyramid-pressure",
    `Core Typology` == "bladeletcore" ~ "bullet-pressure",
    `Core Typology` == "pyramidal? / pressure" ~ "pyramid-pressure",
    `Core Typology` == "prismatic" ~ "pyramid-pressure",

    `Core Typology` == "bullet" ~ "bullet-pressure",

    TRUE ~ `Core Typology`
  )) %>%
# Separate 'Core Typology' into 'core-typology' and 'core-technology'
  separate(`Core Typology`, 
           into = c("core-typology", "core-technology"),
           sep = "-",
           remove = FALSE,
           convert = TRUE)

# Check the unique values again
# rev(sort(table((tl4$`core-typology`))))
# rev(sort(table((tl4$`core-technology`))))

```

```{r}

tl4 <- tl4 %>%
  mutate(
    Typology = str_to_lower(Typology),
    Typology = str_trim(Typology),
    Typology = str_replace_all(Typology, "^[-._/NA\\s]*$|^\\s*$", "0")
  ) %>%
  mutate(
    # BM: wow, seems like almost each artefact is its own type! 
    # BM: That is not ideal for statistical analysis
    # BM: This looks like it was very time-consuming to simplify. 
    ##SS: this column was so messy with too many mistakes; I needed to have all typos correct then transfer them in their corresponding columns.  
    Typology = case_when(
      `Typology` == "notch, truncated" ~ "tool/notch;truncated piece",
      `Typology` == "retouchpiece" ~ "tool/retouched piece",
      `Typology` == "endscraper" ~ "tool/end scraper",
      `Typology` == "notch" ~ "tool/notch",
      `Typology` == "sidescraper" ~ "tool/side scraper",
      `Typology` == "doubleside scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper (?)" ~ "tool/convergent scraper",
      `Typology` == "doublenotch on retouched piece" ~ "tool/double notch",
      `Typology` == "retouchpieces" ~ "tool/retouched piece",
      `Typology` == "burin(?)/ notch/scraper" ~ "tool/burin-notch-scraper",
      `Typology` == "borer" ~ "tool/perforator-borer",
      `Typology` == "retouchedpiece" ~ "tool/retouched piece",
      `Typology` == "notch/side scraper" ~ "tool/notch-side scraper",
      `Typology` == "corefragment" ~ "core/core fragment",
      `Typology` == "backed" ~ "tool/backed knife",
      `Typology` == "doubleside scraper ?" ~ "tool/double side scraper",
      `Typology` == "inversedenticulate" ~ "tool/denticulate",
      `Typology` == "retouchpieces/ side scraper?" ~ "tool/side scraper",
      `Typology` == "sidescraper ?" ~ "tool/side scraper",
      `Typology` == "notch-denticulate" ~ "tool/notch-denticulate",
      `Typology` == "borer/drill" ~ "tool/perforator-drill",
      `Typology` == "retouchedpiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchepiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchedpiece (double side scraper)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece (double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchepiece" ~ "tool/retouched piece",
      `Typology` == "inversenotch" ~ "tool/notch",
      `Typology` == "retouchepiece (double side scaraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece( double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "notchon retouche piece" ~ "tool/notch",
      `Typology` == "retouchedpiece (souble side scraper)" ~ "tool/double side scraper",
      `Typology` == "point" ~ "tool/perforator-point",
      `Typology` == "chunk" ~ "NA/NA",
      `Typology` == "coretablet" ~ "core/core tablet",
      `Typology` == "core" ~ "core/shapeless",
      `Typology` == "denticulate" ~ "tool/denticulate",
      `Typology` == "doublenotch (haft)" ~ "tool/double notch",
      `Typology` == "retouchedpiece (fine retouch)" ~ "tool/retouched piece",
      `Typology` == "endscraper on retouched piece" ~ "tool/end scraper",
      `Typology` == "retouchedpiece (serrated scraper)" ~ "tool/serrated scraper",
      `Typology` == "truncation" ~ "tool/truncated piece",
      `Typology` == "rejuventionpiece" ~ "core/rejuvenation piece",
      `Typology` == "notch/denticulate" ~ "tool/notch-denticulate",
      `Typology` == "retoucheduse on breakage" ~ "tool/retouched piece",
      `Typology` == "inversenotch/endscraper" ~ "tool/notch-end scraper",
      `Typology` == "alternatingside scraper" ~ "tool/side scraper",
      `Typology` == "alternatedouble side scraper" ~ "tool/double side scraper",
      `Typology` == "truncation/sidescaraper" ~ "tool/truncated piece",
      `Typology` == "doublenotch" ~ "tool/double notch",
      `Typology` == "borer/awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper?" ~ "tool/round scraper",
      `Typology` == "doublenotch scraper" ~ "tool/double notch scraper",
      `Typology` == "corerejuvention (core tablet?)" ~ "core/core tablet",
      `Typology` == "truncationاریبب" ~ "tool/truncated piece",
      `Typology` == "retouchedpiece (side scraper)" ~ "tool/side scraper",
      `Typology` == "retochedpice" ~ "tool/retouched piece",
      `Typology` == "truncation/notch" ~ "tool/truncated piece-notch",
      `Typology` == "doublealternate scraper" ~ "tool/double side scraper",
      `Typology` == "awl/ inverse notch" ~ "tool/notch",
      `Typology` == "retouchedpiece (alternate scraper?)" ~ "tool/double side scraper",
      `Typology` == "inverseside scraper" ~ "tool/side scraper",
      `Typology` == "usedcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "corerejuvention platform" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade??" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejivention" ~ "core/rejuvenation piece",
      `Typology` == "retouchedpice (side scraper?)" ~ "tool/side scraper" ,
      `Typology` == "scraper" ~ "tool/side scraper",
      `Typology` == "sickleblade" ~ "tool/sickle shine",
      `Typology` == "coretool (scraper)" ~ "tool/scraper-on core piece",
      `Typology` == "retouchedpiece (scraper)" ~ "tool/side scraper",
      `Typology` == "inversedenticulate?" ~ "tool/denticulate",
      `Typology` == "retouchedpiece (scraper?)" ~ "tool/side scraper",
      `Typology` == "coreside rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "truncation/endscraper" ~ "tool/endscraper on truncated piece",
      `Typology` == "retouchedpice(double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "doublescraper" ~ "tool/double side scraper",
      `Typology` == "dendiculate" ~ "tool/denticulate",
      `Typology` == "double-scraper" ~ "tool/double side scraper",
      `Typology` == "pyramidalcore" ~ "core/pyramid",
      `Typology` == "bulletcore" ~ "core/bullet",
      `Typology` == "corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "pyrmidalcore" ~ "core/pyramid",
      `Typology` == "corepreperation" ~ "core/core preparation piece",
      `Typology` == "crested" ~ "core/crested bladelet",
      `Typology` == "preperationblade" ~ "core/core preperation",
      `Typology` == "burin" ~ "tool/burin",
      `Typology` == "serrateddenticulate" ~ "tool/serrated scraper",
      `Typology` == "saw:serrated denticulate" ~ "tool/serrated scraper",
      `Typology` == "serratedused" ~ "tool/serrated scraper",
      `Typology` == "corepreeration" ~ "tool/core preparation piece",
      `Typology` == "coreprepearation" ~ "tool/core preparation piece",
      `Typology` == "awl" ~ "tool/perforator-awl",
      `Typology` == "notch.inverse" ~ "tool/notch",
      `Typology` == "point.notch" ~ "tool/perforator-point;notch",
      `Typology` == "inversescraper" ~ "tool/side scraper",
      `Typology` == "roundendscraper" ~ "tool/round scraper",
      `Typology` == "alternate.scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper/disc scraper" ~ "tool/scraper-convergent scraper",
      `Typology` == "serratedside scraper" ~ "tool/serrated scraper",
      `Typology` == "debri" ~ "NA/NA",
      `Typology` == "distalpart of a drill" ~ "tool/perforator-drill",
      `Typology` == "corepreperation?" ~ "core/core preparation piece",
      `Typology` == "preformend of a borer" ~ "tool/perforator-borer; preform",
      `Typology` == "used" ~ "tool/used",
      `Typology` == "distalend of a inverse sidescraper" ~ "tool/side scraper",
      `Typology` == "retouchedblade" ~ "tool/retouched piece",
      `Typology` == "drill" ~ "tool/perforator-drill",
      `Typology` == "corerejuvention tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore" ~ "core/pyramid",
      `Typology` == "flake-bladecore" ~ "core/mixed",
      `Typology` == "notch, truncated" ~ "tool/notch;truncated piece",
      `Typology` == "retouchpiece" ~ "tool/retouched piece",
      `Typology` == "endscraper" ~ "tool/end scraper",
      `Typology` == "notch" ~ "tool/notch",
      `Typology` == "sidescraper" ~ "tool/side scraper",
      `Typology` == "doubleside scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper (?)" ~ "tool/convergent scraper",
      `Typology` == "doublenotch on retouched piece" ~ "tool/double notch",
      `Typology` == "retouchpieces" ~ "tool/retouched piece",
      `Typology` == "burin(?)/ notch/scraper" ~ "tool/burin-notch-scraper",
      `Typology` == "borer" ~ "tool/perforator-borer",
      `Typology` == "retouchedpiece" ~ "tool/retouched piece",
      `Typology` == "notch/side scraper" ~ "tool/notch-side scraper",
      `Typology` == "corefragment" ~ "core/core fragment",
      `Typology` == "backed" ~ "tool/backed knife",
      `Typology` == "doubleside scraper ?" ~ "tool/double side scraper",
      `Typology` == "inversedenticulate" ~ "tool/denticulate",
      `Typology` == "retouchpieces/ side scraper?" ~ "tool/side scraper",
      `Typology` == "sidescraper ?" ~ "tool/side scraper",
      `Typology` == "notch-denticulate" ~ "tool/notch-denticulate",
      `Typology` == "borer/drill" ~ "tool/perforator-drill",
      `Typology` == "retouchedpiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchepiece (side scraper?)" ~ "tool/side scraper",
      `Typology` == "retouchedpiece (double side scraper)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece (double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchepiece" ~ "tool/retouched piece",
      `Typology` == "inversenotch" ~ "tool/notch",
      `Typology` == "retouchepiece (double side scaraper?)" ~ "tool/double side scraper",
      `Typology` == "retouchedpiece( double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "notchon retouche piece" ~ "tool/notch",
      `Typology` == "retouchedpiece (souble side scraper)" ~ "tool/double side scraper",
      `Typology` == "point" ~ "tool/perforator-point",
      `Typology` == "chunk" ~ "NA/NA",
      `Typology` == "coretablet" ~ "core/core tablet",
      `Typology` == "core" ~ "core/shapeless",
      `Typology` == "denticulate" ~ "tool/denticulate",
      `Typology` == "doublenotch (haft)" ~ "tool/double notch",
      `Typology` == "retouchedpiece (fine retouch)" ~ "tool/retouched piece",
      `Typology` == "endscraper on retouched piece" ~ "tool/end scraper",
      `Typology` == "retouchedpiece (serrated scraper)" ~ "tool/serrated scraper",
      `Typology` == "truncation" ~ "tool/truncated piece",
      `Typology` == "rejuventionpiece" ~ "core/rejuvenation piece",
      `Typology` == "notch/denticulate" ~ "tool/notch-denticulate",
      `Typology` == "retoucheduse on breakage" ~ "tool/retouched piece",
      `Typology` == "inversenotch/endscraper" ~ "tool/notch-end scraper",
      `Typology` == "alternatingside scraper" ~ "tool/side scraper",
      `Typology` == "alternatedouble side scraper" ~ "tool/double side scraper",
      `Typology` == "truncation/sidescaraper" ~ "tool/truncated piece",
      `Typology` == "doublenotch" ~ "tool/double notch",
      `Typology` == "borer/awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper?" ~ "tool/round scraper",
      `Typology` == "doublenotch scraper" ~ "tool/double notch scraper",
      `Typology` == "corerejuvention (core tablet?)" ~ "core/core tablet",
      `Typology` == "truncationاریبب" ~ "tool/truncated piece",
      `Typology` == "retouchedpiece (side scraper)" ~ "tool/side scraper",
      `Typology` == "retochedpice" ~ "tool/retouched piece",
      `Typology` == "truncation/notch" ~ "tool/truncated piece-notch",
      `Typology` == "doublealternate scraper" ~ "tool/double side scraper",
      `Typology` == "awl/ inverse notch" ~ "tool/notch",
      `Typology` == "retouchedpiece (alternate scraper?)" ~ "tool/double side scraper",
      `Typology` == "inverseside scraper" ~ "tool/side scraper",
      `Typology` == "usedcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "corerejuvention platform" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade" ~ "core/rejuvenation piece",
      `Typology` == "coreside rejuvention blade??" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "coreedge rejivention" ~ "core/rejuvenation piece",
      `Typology` == "retouchedpice (side scraper?)" ~ "tool/side scraper" ,
      `Typology` == "scraper" ~ "tool/side scraper",
      `Typology` == "sickleblade" ~ "tool/sickle shine",
      `Typology` == "coretool (scraper)" ~ "tool/scraper-on core piece",
      `Typology` == "retouchedpiece (scraper)" ~ "tool/side scraper",
      `Typology` == "inversedenticulate?" ~ "tool/denticulate",
      `Typology` == "retouchedpiece (scraper?)" ~ "tool/side scraper",
      `Typology` == "coreside rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "truncation/endscraper" ~ "tool/endscraper on truncated piece",
      `Typology` == "retouchedpice(double side scraper?)" ~ "tool/double side scraper",
      `Typology` == "doublescraper" ~ "tool/double side scraper",
      `Typology` == "dendiculate" ~ "tool/denticulate",
      `Typology` == "double-scraper" ~ "tool/double side scraper",
      `Typology` == "pyramidalcore" ~ "core/pyramid",
      `Typology` == "bulletcore" ~ "core/bullet",
      `Typology` == "corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "pyrmidalcore" ~ "core/pyramid",
      `Typology` == "corepreperation" ~ "core/core preparation piece",
      `Typology` == "crested" ~ "core/crested bladelet",
      `Typology` == "preperationblade" ~ "core/core preperation",
      `Typology` == "burin" ~ "tool/burin",
      `Typology` == "serrateddenticulate" ~ "tool/serrated scraper",
      `Typology` == "saw:serrated denticulate" ~ "tool/serrated scraper",
      `Typology` == "serratedused" ~ "tool/serrated scraper",
      `Typology` == "corepreeration" ~ "tool/core preparation piece",
      `Typology` == "coreprepearation" ~ "tool/core preparation piece",
      `Typology` == "awl" ~ "tool/perforator-awl",
      `Typology` == "notch.inverse" ~ "tool/notch",
      `Typology` == "point.notch" ~ "tool/perforator-point;notch",
      `Typology` == "inversescraper" ~ "tool/side scraper",
      `Typology` == "roundendscraper" ~ "tool/round scraper",
      `Typology` == "alternate.scraper" ~ "tool/double side scraper",
      `Typology` == "convergentscraper/disc scraper" ~ "tool/scraper-convergent scraper",
      `Typology` == "serratedside scraper" ~ "tool/serrated scraper",
      `Typology` == "debri" ~ "NA/NA",
      `Typology` == "distalpart of a drill" ~ "tool/perforator-drill",
      `Typology` == "corepreperation?" ~ "core/core preparation piece",
      `Typology` == "preformend of a borer" ~ "tool/perforator-borer; preform",
      `Typology` == "used" ~ "tool/used",
      `Typology` == "distalend of a inverse sidescraper" ~ "tool/side scraper",
      `Typology` == "retouchedblade" ~ "tool/retouched piece",
      `Typology` == "drill" ~ "tool/perforator-drill",
      `Typology` == "corerejuvention tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore" ~ "core/pyramid",
      `Typology` == "flake-bladecore" ~ "core/mixed",
      `Typology` == "bladelet-corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "bladelet-corefragment" ~ "core/core fragment",
      `Typology` == "pointedused" ~ "tool/perforator-point",
      `Typology` == "microburin" ~ "tool/microburin",
      `Typology` == "bladeletcore fragment" ~ "core/core fragment",
      `Typology` == "rejuventionface flake" ~ "core/rejuvenation piece",
      `Typology` == "corepreperation flake" ~ "core/core preparation piece",
      `Typology` == "corerejuvention flake"  ~ "core/rejuvenation piece",
      `Typology` == "unfinished  pyramid core" ~ "core/pyramid",
      `Typology` == "bladeletcoretablet" ~ "core/core tablet",
      `Typology` == "bladeletcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "debitage" ~ "NA/NA",
      `Typology` == "bladeletcore  rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated scraper" ~ "tool/serrated scraper",
      `Typology` == "backedknife" ~ "tool/backed knife",
      `Typology` == "corepreparation" ~ "core/core preparation piece",
      `Typology` == "primaryflake" ~ "core/primary flake",
      `Typology` == "pointedconvergent scraper on bladelet" ~ "tool/perforator-point",
      `Typology` == "trie" ~ "NA/NA",
      `Typology` == "part of core tablet" ~ "core/core tablet",
      `Typology` == "zaviedar" ~ "NA/NA",
      `Typology` == "convergentscraper" ~ "tool/convergent scraper",
      `Typology` == "preparationflake" ~ "core/core preparation piece",
      `Typology` == "atypicalborer" ~ "tool/perforator-borer",
      `Typology` == "naturalybacked knife" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "pyramidbladelet core" ~ "core/pyramid",
      `Typology` == "flatbladelet core" ~ "core/flat",
      `Typology` == "bladecore" ~ "core/pyramid",
      `Typology` == "bulletbladelet core" ~ "core/bullet",
      `Typology` == "coerejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "finishedbladelet core" ~ "core/bullet",
      `Typology` == "failed pyramid core?" ~ "core/pyramid",
      `Typology` == "pyramidplain bladelet core" ~ "core/pyramid",
      `Typology` == "coeon flake" ~ "core/core on flake",
      `Typology` == "coeejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "bladeletcoe rejuvention tablet?" ~ "core/core tablet",
      `Typology` == "pyramidcore" ~ "core/pyramid",
      `Typology` == "bladeletcore tablet" ~ "core/core tablet",
      `Typology` == "pyramidbladeleet core" ~ "core/pyramid",
      `Typology` == "flake" ~ "NA/NA",
      `Typology` == "serrated" ~ "tool/serrated scraper",
      `Typology` == "curvedretouched piece" ~ "tool/retouched piece",
      `Typology` == "pyramidalbladelet core" ~ "core/pyramid",
      `Typology` == "multidirectionalcore fragment" ~ "core/multidirectional core fragment",
      `Typology` == "unidirectionalblade? core fragent" ~ "core/pyramid",
      `Typology` == "backed/corerejuvention" ~ "tool/backed knife",
      `Typology` == "unidirectionalbladelet core" ~ "core/pyramid",
      `Typology` == "unifacialbalde core fragment?" ~ "core/pyramid",
      `Typology` == "flakedetached from a bladelet core" ~ "core/core preparation piece",
      `Typology` == "corepreparation tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore fragent" ~ "core/core fragment-pyramid",
      `Typology` == "alternatescraper??" ~ "tool/double side scraper",
      `Typology` == "endscraper.onrejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated" ~ "tool/serrated scraper",
      `Typology` == "patination.omitted" ~ "NA/NA",
      `Typology` == "trapze" ~ "tool/geometric-triangle",
      `Typology` == "trapzoid" ~ "tool/geometric-triangle",
      `Typology` == "borer.drill" ~ "tool/perforator-drill",
      `Typology` == "serraed" ~ "tool/serrated scraper",
      `Typology` == "bladeletcore on a flake" ~ "core/core on flake",
      `Typology` == "geofact" ~ "NA/NA",
      `Typology` == "denticulateborer?" ~ "tool/perforator-borer;denticulate",
      `Typology` == "sideborer-denticulate" ~ "tool/perforator-borer;denticulate",
      `Typology` == "borer.awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper" ~ "tool/round scraper",
      `Typology` == "truncationon a bladelet core" ~ "tool/truncated piece",
      `Typology` == "truncation.notch" ~ "tool/truncated piece;notch",
      `Typology` == "backed-denticulate" ~ "tool/backed;denticulate",
      `Typology` == "brokendrill" ~ "tool/perforator-drill",
      `Typology` == "drillbroken" ~ "tool/perforator-drill",
      `Typology` == "point/broken drill" ~ "tool/perforator-point",
      `Typology` == "brokendrill?" ~ "tool/perforator-drill",
      `Typology` == "partialyserrated" ~ "tool/serrated scraper",
      `Typology` == "naturallybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "transversescraper" ~ "tool/transverse scraper",
      `Typology` == "naturalybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "awl-notch" ~ "tool/perforator-awl",
      `Typology` == "psedulevalois" ~ "tool/psedulevalois flake",
      `Typology` == "alternatingdenticulate" ~ "tool/denticulate",
      `Typology` == "usedcore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalmicroflake core" ~ "core/multidirectional flake core",
      `Typology` == "mixedpyramidal core" ~ "core/pyramid",
      `Typology` == "usedbladecore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalbladelet core fragment" ~ "core/multidirectional bladelet core",
      `Typology` == "unidirectionalbladelet core.cylinder" ~ "core/prismatic",
      `Typology` == "pyrymedalmixed core" ~ "core/pyramid",
      `Typology` == "pyramedial/bullet core" ~ "core/bullet",
      `Typology` == "corefragent" ~ "core/core fragment",
      `Typology` == "pyramedialbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unidirectionalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unifacialunidirectional bladelet core" ~ "core/pyramid",
      `Typology` == "unidirectionalblade core" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "corerejuvention?" ~ "core/rejuvenation piece",
      `Typology` == "flatburin? point?" ~ "tool/flat burin",
      `Typology` == "primaryblade" ~ "core/primary blade",
      `Typology` == "trihedralgeometric" ~ "tool/geometric-triangle",
      `Typology` == "lunate" ~ "tool/geometric-lunate",
      `Typology` == "corepaltform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "pressurebladelet core /semi flat" ~ "core/pyramid",
      `Typology` == "failedpercussion blade core/semi pyramedal" ~ "core/pyramid",
      `Typology` == "percussionbladelet core/unidirectional" ~ "core/pyramid-percussion",
      `Typology` == "percussion.unidirectional.blade-flakecore" ~ "core/mixed-percussion",
      `Typology` == "obliqueretouched bladelet" ~ "tool/retouched piece",
      `Typology` == "percussionunidirectional pyramidal  bladelet core" ~ "core/pyramid",
      `Typology` == "lunategeometric" ~ "tool/geometric-lunate",
      `Typology` == "awl on a core rejuvention" ~ "tool/perforator-awl on core rejuvenation piece",
      `Typology` == "truncated" ~ "tool/truncated piece",
      `Typology` == "multiplenotch" ~ "tool/notch",
      `Typology` == "dishedconvergent bladelet" ~ "tool/convergent scraper",
      `Typology` == "bipolarpercussion blade-bladelet core" ~ "core/bipolar-percussion",
      `Typology` == "bladeletbullet core" ~ "core/bullet",
      `Typology` == "alternatingserrated" ~ "tool/serrated scraper",
      `Typology` == "shaplesscore" ~ "core/shapeless",
      `Typology` == "corepreparation?" ~ "core/core preparation piece",
      `Typology` == "obliqueretouched" ~ "tool/retouched piece",
      `Typology` == "corepreparation flake" ~ "core/core preparation piece",
      `Typology` == "backeddenticulate" ~ "tool/backed;denticulate",
      `Typology` == "awlnotch" ~ "tool/perforator-awl;notch",
      `Typology` == "scraperburin?" ~ "tool/burin;scraper",
      `Typology` == "notchround scraper" ~ "tool/round scraper;notch",
      `Typology` == "corerejuvention tablet round scraper" ~ "tool/round scraper on core tablet",
      `Typology` == "alternatescraper" ~ "tool/double side scraper",
      `Typology` == "coreplatform preperation flake" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention tablet" ~ "core/core tablet",
      `Typology` == "brokenpyramidal core bladelet" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core/semi bullet?" ~ "core/bullet",
      `Typology` == "coreside rejuvention flake?" ~ "core/rejuvenation piece-side",
      `Typology` == "unidirectionalpressure bladelet core/unifacila" ~ "core/pyramid",
      `Typology` == "coreplatform rejuvention tablet/or multidirectional core" ~ "core/rejuvenation piece",
      `Typology` == "bulletbladlet core" ~ "core/bullet",
      `Typology` == "cylinderbipolar bladelet core" ~ "core/prismatic",
      `Typology` == "backedknife/core platform rejuvenation" ~ "tool/backed knife",
      `Typology` == "flatunifacial bladelet core" ~ "core/flat",
      `Typology` == "coreprepration flake" ~ "core/core preparation piece",
      `Typology` == "coreplatform preperation" ~ "core/core preparation piece-platform",
      `Typology` == "borerpreform" ~ "tool/perforator-borer",
      `Typology` == "rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "borerpreform?" ~ "tool/perforator-borer",
      `Typology` == "inversnotch" ~ "tool/notch",
      `Typology` == "alternateconvergent" ~ "tool/convergent scraper",
      `Typology` == "unidirectionalbladelet/flake-blade core" ~ "core/mixed",
      `Typology` == "unidirectionalmixed core" ~ "core/mixed",
      `Typology` == "unidirectionalflat bladelet core" ~ "core/flat",
      `Typology` == "" ~ "NA-NA",
      `Typology` == "na" ~ "NA-NA",
      `Typology` == "bladelet-corerejuvention" ~ "core/rejuvenation piece",
      `Typology` == "bladelet-corefragment" ~ "core/core fragment",
      `Typology` == "pointedused" ~ "tool/perforator-point",
      `Typology` == "microburin" ~ "tool/microburin",
      `Typology` == "bladeletcore fragment" ~ "core/core fragment",
      `Typology` == "rejuventionface flake" ~ "core/rejuvenation piece",
      `Typology` == "corepreperation flake" ~ "core/core preparation piece",
      `Typology` == "corerejuvention flake"  ~ "core/rejuvenation piece",
      `Typology` == "unfinished  pyramid core" ~ "core/pyramid",
      `Typology` == "bladeletcoretablet" ~ "core/core tablet",
      `Typology` == "bladeletcore rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "debitage" ~ "NA/NA",
      `Typology` == "bladeletcore  rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated scraper" ~ "tool/serrated scraper",
      `Typology` == "backedknife" ~ "tool/backed knife",
      `Typology` == "corepreparation" ~ "core/core preparation piece",
      `Typology` == "primaryflake" ~ "core/primary flake",
      `Typology` == "pointedconvergent scraper on bladelet" ~ "tool/perforator-point",
      `Typology` == "trie" ~ "NA/NA",
      `Typology` == "part of core tablet" ~ "core/core tablet",
      `Typology` == "zaviedar" ~ "NA/NA",
      `Typology` == "convergentscraper" ~ "tool/convergent scraper",
      `Typology` == "preparationflake" ~ "core/core preparation piece",
      `Typology` == "atypicalborer" ~ "tool/perforator-borer",
      `Typology` == "naturalybacked knife" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "pyramidbladelet core" ~ "core/pyramid",
      `Typology` == "flatbladelet core" ~ "core/flat",
      `Typology` == "bladecore" ~ "core/pyramid",
      `Typology` == "bulletbladelet core" ~ "core/bullet",
      `Typology` == "coerejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "finishedbladelet core" ~ "core/bullet",
      `Typology` == "failed pyramid core?" ~ "core/pyramid",
      `Typology` == "pyramidplain bladelet core" ~ "core/pyramid",
      `Typology` == "coeon flake" ~ "core/core on flake",
      `Typology` == "coeejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "bladeletcoe rejuvention tablet?" ~ "core/core tablet",
      `Typology` == "pyramidcore" ~ "core/pyramid",
      `Typology` == "bladeletcore tablet" ~ "core/core tablet",
      `Typology` == "pyramidbladeleet core" ~ "core/pyramid",
      `Typology` == "flake" ~ "NA/NA",
      `Typology` == "serrated" ~ "tool/serrated scraper",
      `Typology` == "curvedretouched piece" ~ "tool/retouched piece",
      `Typology` == "pyramidalbladelet core" ~ "core/pyramid",
      `Typology` == "multidirectionalcore fragment" ~ "core/multidirectional core fragment",
      `Typology` == "unidirectionalblade? core fragent" ~ "core/pyramid",
      `Typology` == "backed/corerejuvention" ~ "tool/backed knife",
      `Typology` == "unidirectionalbladelet core" ~ "core/pyramid",
      `Typology` == "unifacialbalde core fragment?" ~ "core/pyramid",
      `Typology` == "flakedetached from a bladelet core" ~ "core/core preparation piece",
      `Typology` == "corepreparation tablet" ~ "core/core tablet",
      `Typology` == "bladeletcore fragent" ~ "core/core fragment-pyramid",
      `Typology` == "alternatescraper??" ~ "tool/double side scraper",
      `Typology` == "endscraper.onrejuvention" ~ "core/rejuvenation piece",
      `Typology` == "inverseserrated" ~ "tool/serrated scraper",
      `Typology` == "patination.omitted" ~ "NA/NA",
      `Typology` == "trapze" ~ "tool/geometric-triangle",
      `Typology` == "trapzoid" ~ "tool/geometric-triangle",
      `Typology` == "borer.drill" ~ "tool/perforator-drill",
      `Typology` == "serraed" ~ "tool/serrated scraper",
      `Typology` == "bladeletcore on a flake" ~ "core/core on flake",
      `Typology` == "geofact" ~ "NA/NA",
      `Typology` == "denticulateborer?" ~ "tool/perforator-borer;denticulate",
      `Typology` == "sideborer-denticulate" ~ "tool/perforator-borer;denticulate",
      `Typology` == "borer.awl" ~ "tool/perforator-awl",
      `Typology` == "roundscraper" ~ "tool/round scraper",
      `Typology` == "truncationon a bladelet core" ~ "tool/truncated piece",
      `Typology` == "truncation.notch" ~ "tool/truncated piece;notch",
      `Typology` == "backed-denticulate" ~ "tool/backed;denticulate",
      `Typology` == "brokendrill" ~ "tool/perforator-drill",
      `Typology` == "drillbroken" ~ "tool/perforator-drill",
      `Typology` == "point/broken drill" ~ "tool/perforator-point",
      `Typology` == "brokendrill?" ~ "tool/perforator-drill",
      `Typology` == "partialyserrated" ~ "tool/serrated scraper",
      `Typology` == "naturallybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "transversescraper" ~ "tool/transverse scraper",
      `Typology` == "naturalybacked" ~ "tool/backed knife-naturally backed knife",
      `Typology` == "awl-notch" ~ "tool/perforator-awl",
      `Typology` == "psedulevalois" ~ "tool/psedulevalois flake",
      `Typology` == "alternatingdenticulate" ~ "tool/denticulate",
      `Typology` == "usedcore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalmicroflake core" ~ "core/multidirectional flake core",
      `Typology` == "mixedpyramidal core" ~ "core/pyramid",
      `Typology` == "usedbladecore fragment" ~ "tool/core tool-used core fragment",
      `Typology` == "multidirectionalbladelet core fragment" ~ "core/multidirectional bladelet core",
      `Typology` == "unidirectionalbladelet core.cylinder" ~ "core/prismatic",
      `Typology` == "pyrymedalmixed core" ~ "core/pyramid",
      `Typology` == "pyramedial/bullet core" ~ "core/bullet",
      `Typology` == "corefragent" ~ "core/core fragment",
      `Typology` == "pyramedialbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unidirectionalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "unifacialunidirectional bladelet core" ~ "core/pyramid",
      `Typology` == "unidirectionalblade core" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core fragment" ~ "core/pyramid",
      `Typology` == "corerejuvention?" ~ "core/rejuvenation piece",
      `Typology` == "flatburin? point?" ~ "tool/flat burin",
      `Typology` == "primaryblade" ~ "core/primary blade",
      `Typology` == "trihedralgeometric" ~ "tool/geometric-triangle",
      `Typology` == "lunate" ~ "tool/geometric-lunate",
      `Typology` == "corepaltform rejuvention flake" ~ "core/rejuvenation piece",
      `Typology` == "pressurebladelet core /semi flat" ~ "core/pyramid",
      `Typology` == "failedpercussion blade core/semi pyramedal" ~ "core/pyramid",
      `Typology` == "percussionbladelet core/unidirectional" ~ "core/pyramid-percussion",
      `Typology` == "percussion.unidirectional.blade-flakecore" ~ "core/mixed-percussion",
      `Typology` == "obliqueretouched bladelet" ~ "tool/retouched piece",
      `Typology` == "percussionunidirectional pyramidal  bladelet core" ~ "core/pyramid",
      `Typology` == "lunategeometric" ~ "tool/geometric-lunate",
      `Typology` == "awl on a core rejuvention" ~ "tool/perforator-awl on core rejuvenation piece",
      `Typology` == "truncated" ~ "tool/truncated piece",
      `Typology` == "multiplenotch" ~ "tool/notch",
      `Typology` == "dishedconvergent bladelet" ~ "tool/convergent scraper",
      `Typology` == "bipolarpercussion blade-bladelet core" ~ "core/bipolar-percussion",
      `Typology` == "bladeletbullet core" ~ "core/bullet",
      `Typology` == "alternatingserrated" ~ "tool/serrated scraper",
      `Typology` == "shaplesscore" ~ "core/shapeless",
      `Typology` == "corepreparation?" ~ "core/core preparation piece",
      `Typology` == "obliqueretouched" ~ "tool/retouched piece",
      `Typology` == "corepreparation flake" ~ "core/core preparation piece",
      `Typology` == "backeddenticulate" ~ "tool/backed;denticulate",
      `Typology` == "awlnotch" ~ "tool/perforator-awl;notch",
      `Typology` == "scraperburin?" ~ "tool/burin;scraper",
      `Typology` == "notchround scraper" ~ "tool/round scraper;notch",
      `Typology` == "corerejuvention tablet round scraper" ~ "tool/round scraper on core tablet",
      `Typology` == "alternatescraper" ~ "tool/double side scraper",
      `Typology` == "coreplatform preperation flake" ~ "core/rejuvenation piece",
      `Typology` == "coreplatform rejuvention tablet" ~ "core/core tablet",
      `Typology` == "brokenpyramidal core bladelet" ~ "core/pyramid",
      `Typology` == "pyramidalbladelet core/semi bullet?" ~ "core/bullet",
      `Typology` == "coreside rejuvention flake?" ~ "core/rejuvenation piece-side",
      `Typology` == "unidirectionalpressure bladelet core/unifacila" ~ "core/pyramid",
      `Typology` == "coreplatform rejuvention tablet/or multidirectional core" ~ "core/rejuvenation piece",
      `Typology` == "bulletbladlet core" ~ "core/bullet",
      `Typology` == "cylinderbipolar bladelet core" ~ "core/prismatic",
      `Typology` == "backedknife/core platform rejuvenation" ~ "tool/backed knife",
      `Typology` == "flatunifacial bladelet core" ~ "core/flat",
      `Typology` == "coreprepration flake" ~ "core/core preparation piece",
      `Typology` == "coreplatform preperation" ~ "core/core preparation piece-platform",
      `Typology` == "borerpreform" ~ "tool/perforator-borer",
      `Typology` == "rejuvention" ~ "core/rejuvenation piece",
      `Typology` == "borerpreform?" ~ "tool/perforator-borer",
      `Typology` == "inversnotch" ~ "tool/notch",
      `Typology` == "alternateconvergent" ~ "tool/convergent scraper",
      `Typology` == "unidirectionalbladelet/flake-blade core" ~ "core/mixed",
      `Typology` == "unidirectionalmixed core" ~ "core/mixed",
      `Typology` == "unidirectionalflat bladelet core" ~ "core/flat",
      `Typology` == "" ~ "NA-NA",
      `Typology` == "na" ~ "NA-NA",
      TRUE ~ `Typology`
    )
  )  %>%
  mutate(
    typology_split = str_split(Typology, "/", simplify = TRUE),
    `typology-tool-core` = typology_split[, 1],
    tool_or_core = ifelse(str_starts(`typology-tool-core`, "tool"), "tool", "core"),
    `tool-typology` = ifelse(tool_or_core == "tool", typology_split[, 2], NA),
    `core-fragment` = ifelse(tool_or_core == "core", typology_split[, 2], NA)
  ) %>%
  # remove temporary columns
  select(-typology_split,-tool_or_core)

# BM: still some typos in there
# rev(sort(table((tl4$`tool-typology`))))
# rev(sort(table((tl4$`core-fragment`))))
```

```{r}

# refine the table

tl_final <- tl4 %>%
  rename(
    blank =                Blank2,
   `sickle shine` =       shine2,
    length =              `Length (mm)` ,
    width =               `Width (mm)`,
    thickness =           `Thickness (mm)`, 
    typology =             Typology,
    utilization =          Utilization,
    retouch =              Retouch,
    blank =                Blank2, 
   `blank part` =         `Blank_part`,
   `bulb thickness` =     `Bulb Thickness (mm)`, 
   `platform thickness` = `Platform Thickness (mm)`, 
   `platform length` =    `platform Length (mm)`, 
    eraillure =            Eraillure , 
    cortex =               `Cortex (%)`)  %>%
  mutate(area_unit = paste0(toupper(area), "_", chronological_unit)) %>%
  relocate(area_unit, .after = chronological_unit)   %>%
  relocate(grouped_level, .after = area_unit)   %>%
  select(lithic_id, 
         TA, 
         area, 
         chronological_unit,
         area_unit,
         grouped_level,
         cortex,
         depth, upper, lower, midpoint, 
        `length`, `width`, `thickness`, 
         typology, 
         utilization, 
        `sickle shine`,
         retouch, 
        `typology-tool-core`,
        `tool-typology`,
         blank, `blank part`, 
        `core-typology`, `core-fragment`,`core-technology`, 
        cortex)  %>% 
# Update the 'typology-tool-core' column based on 'Retouch'
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",
      TRUE ~ `typology-tool-core`      
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
      utilization == 1 ~ "utilized tool",
      `sickle shine` == 1 ~ "sickle shine",
      TRUE ~ `tool-typology`               
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
     # If Retouch is 1 and tool-typology is NA
       retouch == 1 & is.na(`tool-typology`) ~ "retouched piece",
       `sickle shine` == 1 & is.na(`tool-typology`) ~ "sickle shine",

      # For all other cases 
      TRUE ~ `tool-typology`                                       
    )
  )

# BM: take a look at tool typology
# rev(sort(table(tl_final$`tool-typology`)))

```

```{r}
# List of columns that should be numeric
columns_to_convert <- c("cortex", 
                        "upper", 
                        "lower", 
                        "midpoint", 
                        "length", 
                        "width", 
                        "utilization", 
                        "retouch",
                        "sickle shine")

# Convert each column to numeric

# BM: this is how I would convert those columns to numeric, 
# I think it's easier to read and requires less typing
tl_final <- 
tl_final %>% 
 mutate(across(all_of(columns_to_convert), 
               parse_number))

# check the structure to see if they have been converted
# str(tl_final)
```

```{r}
#set the value of utilization to NA only for the rows where `sickle shine` == 1.

tl_final <- tl_final %>%
  rowwise() %>%
  mutate(utilization = ifelse(is.na(`sickle shine`) | is.na(utilization),
                              utilization,
                              ifelse(`sickle shine` == 1 & utilization == 1, 
                                     NA_real_, 
                                     utilization))) %>%
  ungroup() %>%
  # make the number of `sickle shine` in column `sickle shine` match:
  mutate(
    `tool-typology` = case_when(
      `sickle shine` == 1 ~ "sickle shine",
      utilization == 1 ~ "utilized tool",
      TRUE ~ `tool-typology`               
    )
  )

```

```{r}

# Creating tooltypegroup

tl_final <- tl_final %>%
  mutate(
    tooltypegroup = case_when(
      `tool-typology` %in%
        c(
          "end scraper",
          "side scraper",
          "double side scraper",
          "convergent scraper",
          "round scraper",
          "perforator-convergent scraper",
          "perforator-borer; preform",
          "scraper-on core piece",
          "endscraper on truncated piece",
          "round scraper",
          "transverse scraper",
          "core tool-used core fragment",
          "round scraper on core tablet"
        ) ~ "scraper",
      `tool-typology` %in%
        c(
          "perforator-borer",
          "perforator-drill",
          "perforator-point",
          "perforator-awl",
          "perforator-point;notch",
          "perforator-awl on core rejuvenation piece",
          "perforator-awl;notch",
          "perforator-borer;denticulate"
        ) ~ "Perforator",
      `tool-typology` %in%
        c(
          "backed knife",
          "backed knife-naturally backed knife",
          "backed;denticulate"
        ) ~ "backed pieces",
      `tool-typology` %in% c(
        "double notch",
        "denticulate",
        " denticulate",
        "notch-denticulate",
        "notch"
      ) ~ "denticulate-notch",
      `tool-typology` %in% c(
        "notch-side scraper",
        "notch-end scraper",
        "double notch scraper",
        "round scraper;notch"
      ) ~ "scraper-notch",
      `tool-typology` %in% c("geometric-triangle", "geometric-lunate") ~ "geometric",
      `tool-typology` == "serrated scraper" ~ "serrated scraper",
      `tool-typology` %in% c(
        "truncated piece",
        "truncated piece-notch",
        "truncated piece;notch",
        "notch;truncated piece"
      ) ~ "truncated pieces",

      # Represents NA group
      `tool-typology` %in% c("core preparation piece", "used", "psedulevalois flake") ~ NA_character_,
      `tool-typology` %in% c("burin", "flat burin", "burin;scraper", "burin-notch-scraper") ~ "burin",
      `tool-typology` == "microburin" ~ "microburin",
      `tool-typology` == "retouched piece" ~ "retouched piece",
      `tool-typology` == "utilized tool" ~ "utilized tool",
      `tool-typology` == "sickle shine" ~ "sickle shine",

      #classify any remaining values as NA
      TRUE ~ NA_character_
    )
  )

# Checking the unique values of tooltypegroup
# rev(sort(table(tl_final$tooltypegroup)))

```

```{r}

# Define the patterns to search for
patterns <- c("denticulate-notch", "retouched piece", "scraper",
              "Perforator", "serrated scraper", "backed pieces",
              "truncated pieces", "burin", "geometric", "scraper-notch",
              "microburin")

# Update the `retouch` column based on the `tooltypegroup` column
tl_final <- tl_final %>%
  mutate(retouch = ifelse(str_detect(tooltypegroup, 
                                     str_c(patterns, 
                                           collapse = "|")), 
                          1, retouch))

# View the first few rows to confirm
# head(tl_final)
# rev(sort(table(tl_final$tooltypegroup)))
# rev(sort(table(tl_final$retouch)))
```

```{r}
tl_final <- tl_final %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`))  %>%
  mutate(`core-fragment` = if_else(!is.na(`core-typology`),
                                   NA_character_,
                                   `core-fragment`)) %>%
  mutate(
    `core-fragment` = case_when(
      `core-fragment` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-fragment` == "rejuvention piece-side" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece-platform" ~ "core tablet",
      `core-fragment` == "core preparation" ~ "rejuvenation piece",
      `core-fragment` == "core preparation piece" ~ "rejuvenation piece",
      `core-fragment` == "core preperation" ~ "rejuvenation piece",
      `core-fragment` == "pyramid" ~ "pyramid",
      `core-fragment` == "pyramid core" ~ "pyramid",
      `core-fragment` == "pyramid-unidirectional" ~ "pyramid",
      `core-fragment` == "core fragment-pyramid" ~ "pyramid",
      `core-fragment` == "cylinder" ~ "cylinder/prismatic",
      `core-fragment` == "prismatic" ~ "cylinder/prismatic",
      `core-fragment` == "shapeless" ~ "cylinder/prismatic",
      `core-fragment` == "multidirectional core fragment" ~ "shapeless",
      `core-fragment` == "multidirectional flake core" ~ "shapeless",
      `core-fragment` == "multidirectional bladelet core" ~ "shapeless",
      `core-fragment` == "mixed core" ~ "shapeless",
      `core-fragment` == "mixed-percussion" ~ "shapeless",
      `core-fragment` == "mixed" ~ "shapeless",
      `core-fragment` == ("bipolar-percussion") ~ "bipolar",
      `core-fragment` == "flat" ~ "flat",
      `core-fragment` == "core on flake" ~ "core on flake",
      `core-fragment` == "bullet" ~ "bullet",
      TRUE ~ `core-fragment`

    )
  )  %>%
  mutate(
    core = case_when(
      `core-fragment` %in% c(
        "pyramid",
        "pyramid core",
        "pyramid-unidirectional",
        "core fragment-pyramid",
        "cylinder/prismatic",
        "shapeless",
        "multidirectional core fragment",
        "multidirectional flake core",
        "multidirectional bladelet core",
        "mixed core",
        "mixed-percussion",
        "mixed",
        "bipolar-percussion",
        "flat",
        "core on flake",
        "bullet"
      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    ),
    corefrag = case_when(
      `core-fragment` %in% c(
        "rejuvenation piece",
        "rejuvention piece-side",
        "core preparation piece-platform",
        "core preparation",
        "core preparation piece",
        "core preperation",
        "crested bladelet" ,
        "primary flake" ,
        "primary blade"
      ) ~ `core-fragment`,
      TRUE ~ NA_character_
    )
  ) %>%
  # Convert empty strings to NA in `core-fragment` column
  mutate(`core-fragment` = if_else(`core-fragment` == "", NA_character_, `core-fragment`)) %>%
  # Convert "pyramid-percussion" to "pyramid" in `core-fragment` column
  mutate(
    `core-fragment` =
      if_else(
        `core-fragment` == "pyramid-percussion",
        "pyramid",
        `core-fragment`
      )
  ) %>%
  # Update existing `core` column with selected values from `core-fragment` column
  mutate(core = if_else(
    `core-fragment` %in%
      c(
        "pyramid",
        "cylinder/prismatic",
        "flat",
        "bullet",
        "core on flake",
        "pyramid"
      ),
    `core-fragment`,
    core
  )) %>%
  # Add 'core tablet' and 'core fragment' from `core-fragment` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-fragment` == "core tablet" ~ "core tablet",
      `core-fragment` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Transfer 'rejuvenation piece' and 'core fragment' from `core-typology` to `corefrag`
  mutate(
    corefrag = case_when(
      `core-typology` == "rejuvenation piece" ~ "rejuvenation piece",
      `core-typology` == "core fragment" ~ "core fragment",
      TRUE ~ corefrag
    )
  ) %>%
  # Remove 'rejuvenation piece' and 'core fragment' from `core-typology`
  mutate(`core-typology` = if_else(
    `core-typology` %in% c("rejuvenation piece", "core fragment"),
    NA_character_,
    `core-typology`
  ))  %>% 
# Update the 'blank' column based on conditions in the 'core-fragment' column
  mutate(blank = if_else(
    `core-fragment` %in% c(
      "core fragment",
      "core tablet",
      "rejuvenation piece",
      "pyramid",
      "crested bladelet",
      "shapeless",
      "primary flake",
      "flat",
      "bullet",
      "core on flake",
      "cylinder/prismatic",
      "primary blade",
      "rejuvenation piece-side",
      "core platform rejuvention"
    ),
    NA_character_,
    blank
  ))

# take a look
# rev(sort(table(tl_final$core)))
# rev(sort(table(tl_final$corefrag)))
# rev(sort(table(tl_final$`core-fragment`)))
# rev(sort(table(tl_final$blank)))
```

### Summary tables: AREA

```{r}
# Create a summary table
summary_table_area <- tl_final %>%
  group_by(area) %>%
  summarise(
    `Cores (n)` = sum(!is.na(core) & core != ""),
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag) &
                                            corefrag != ""),
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

# SS: we mainly consider ta, tp1, and d1.
knitr::kable(summary_table_area,
             caption = "Summary of lithics in all areas")
```

```{r}

# Filter the data to include only the areas 'ta', 'd1', and 'tp1'
tl_final_area <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1"))

summary_table_area_filtered <- summary_table_area %>%
  filter(area %in% c("ta", "d1", "tp1"))

knitr::kable(summary_table_area_filtered,
             caption = "Summary of lithics in the main areas")

```

### Summary tables: chronological unit

```{r}
# Create a summary table
summary_table_unit <- tl_final %>%
  group_by(chronological_unit) %>% 
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```

```{r}
# Generate the table
kable(summary_table_unit, 
      caption = "Summary of lithics by Chronological Units", 
      align = 'c')

```

### Summary tables: AREA-UNIT:

```{r}
# Create a summary table based on the 'area_unit' column
summary_table_au <- tl_final %>%
  group_by(area_unit) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

```

```{r}
# Generate the table

kable(summary_table_al, 
      caption = "Summary of lithics by Area and chronological unit")

```

### Summary tables: Techno-Typology

```{r}

# Prepare data for Summary table-area
summary_table_area <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(area) %>%
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

summary_table_area <- summary_table_area %>%
  add_row(
    area = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),

  )

# Adding a new row with total counts
summary_table_area <- summary_table_area %>%
  add_row(
    area = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  # Explicitly setting the total
  )

```

```{r}
# Prepare data for Summary table-area_unit(au)
summary_table_au <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>% 
  group_by(area_unit) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
summary_table_al <- summary_table_al %>% 
  add_row(
    area_unit = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  
  )

# rev(sort(table(tl_final$tooltypegroup)))
```

```{r}
#explore tool type group

summary_table_tooltype <- tl_final %>%
  filter(tooltypegroup != 'utilized tool') %>%  # Exclude 'utilized tool'
  group_by(tooltypegroup) %>% 
  summarise(
  )
```

```{r}
# generate a table

summary_table_al <- summary_table_al %>%
  filter(!(area_unit %in% c("D1_NA", "TP1_NA")))

kable(
  summary_table_al,
  col.names = c(
    "Area-unit",
    "Blade",
    "Flake",
    "Bladelet",
    "Tools",
    "Cores",
    "Core Rejuvenations",
    "Retouched Tools",
    "Utilized Tools",
    "Sickle Shine",
    "Total"
  ),
  caption = "Summary of lithics by techno-typological classification")

```

```{r}
# Create a summary table for typology based on area

summary_table_tool_area <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  filter(!is.na(tooltypegroup), tooltypegroup != "") %>%  # Exclude rows where 'tooltypegroup' is NA or empty
  group_by(area, tooltypegroup) %>%  # Group the data by area and tooltypegroup
  tally() %>%  # Count the number of rows in each group
  arrange(area, tooltypegroup) %>%  # Arrange by area and tooltypegroup
  spread(key = area, value = n, fill = 0)  # Spread the 'area' column into multiple columns

# Add a "Total" column that sums each row
summary_table_tool_area <- summary_table_tool_area %>%
  mutate(Total = rowSums(select(.,-tooltypegroup)))

total_row <- summary_table_tool_area %>%
  select(-tooltypegroup) %>%
  summarise(across(everything(), \(x) sum(x, na.rm = TRUE))) %>%
  mutate(tooltypegroup = "Total")

summary_table_tool_area <-
  bind_rows(summary_table_tool_area, total_row)

```

```{r}

kable(
  summary_table_tool_area,
  col.names = c("Tools", "D1", "TA", "TP1", "Total"),
  caption = "Tool Typology in the three main areas")

```

```{r}
# Create the summary table based on chronological_unit
summary_table1_unit <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(chronological_unit) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    numb_cores = sum(!is.na(`core-typology`), na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

```

```{r}
# Create a summary table based on the 'depth', 'area', and 'chronological unit' columns
tl_final <- tl_final %>%
  mutate(depth = gsub("cm", "", depth))

summary_table_depth_area_unit <- tl_final %>%
  group_by(TA, depth, area, chronological_unit, grouped_level, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

summary_table_depth_area_unit_to_join <- 
  summary_table_depth_area_unit

# Generate the table
kable(summary_table_depth_area_unit, 
      caption = "Summary of lithics by Depth, Area, and Chronological Unit")
```

```{r}
# Generate a summary table by grouping by 'depth' and 'area'
summary_table_depth_tool_type <- tl_final %>%
  group_by(area, depth) %>%
  summarise(tool_type = n_distinct(tooltypegroup, na.rm = TRUE), .groups = 'drop')

# Create a full summary table by joining it back to the original summary table based on 'depth' and 'area'
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  left_join(summary_table_depth_tool_type, by = c("area", "depth"))

# Remove the existing 'tool_type' if it exists
if ("tool_type" %in% names(summary_table_depth_area_unit)) {
  summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
    select(-tool_type)
}

# Join again based on 'depth', this time being specific about what we want
summary_table_depth_area_unit <- 
  summary_table_depth_area_unit %>%
  left_join(summary_table_depth_tool_type %>% select(depth, tool_type), 
            by = "depth")

# Generate the table

kable(summary_table_depth_area_unit, 
      caption = "Summary of lithics by Depth, Area, and chronological unit")

```

```{r}
# Specify the exact columns want to keep
desired_columns <-
  c(
    "TA",
    "depth",
    "area",
    "chronological_unit",
    "grouped_level",
    "upper",
    "lower",
    "Cores (n)",
    "Core rejuvenations pieces (n)",
    "Blade (n)",
    "Flake (n)",
    "Bladelet (n)",
    "Retouched Tools (n)",
    "Utilized (n)",
    "Sickle Shine (n)",
    "tool_type"
  )

# Filter the tl_final dataframe
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  filter(area %in% c('ta', 'd1', 'tp1'))

# Generate the table
kable(summary_table_depth_area_unit, 
      caption = "Summary of lithics by Depth, Area, and chronological_unit")
```

```{r}

# import table with depths that we have subdivided
depths_divided_tbl <- 
  read_csv(here("data/Lookup Table - Sheet1.csv")) %>% 
  mutate(division_id = rep(c("a", "b"),
                           nrow(.)/2))

# google sheet has 42 rows, and cols for 
# area, chronological_unit, level, depth. What is the equivalent above? tp1_c_3_a

# This 'depths_divided_tbl' needs more information, for example, some depth ranges

# join with lithic data that 
summary_table_depth_area_unit_to_join %>% 
  left_join(depths_divided_tbl) %>% 
  mutate(label = paste(area, 
                       chronological_unit, 
                       # level, # need this
                       division_id,
                       sep = "_")) %>% View

```

```{r}
#| eval: false

# Here, I've imported the table 'Summary by Depth, Area, chronological_unit' into Google Sheets for easier management.
#I thought this format would simplify the process of those depths needed to be splitting up.

# BM: looks like you need this data starting from around here in the document

library(googlesheets4)

google_sheet_url <- 
  "https://docs.google.com/spreadsheets/d/1Q0QZESk412ZQLE24yPs6Rg-7Y9OMX63DHoXeIPNAIFM/edit#gid=0"
sum_depth_sheet <- read_sheet(google_sheet_url)

# Create the 'label' column
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(label = paste(area, chronological_unit, grouped_level, sep = "_"))

# Add a suffix only to duplicates
sum_depth_sheet <- sum_depth_sheet %>%
  group_by(label) %>%
  mutate(suffix = case_when(
    n() == 1 ~ NA_character_,
    TRUE ~ letters[row_number()]
  )) %>%
  ungroup()

# Append the suffix to 'label'
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(label = if_else(is.na(suffix), label, paste0(label, "_", suffix))) %>%
  select(-suffix)

# Relocate 'label' column after 'level'
sum_depth_sheet <- sum_depth_sheet %>%
  relocate(label, .after = level)

# BM: from my perspective, what would be ideal here is to have the 
# excavation area and volume calculations done here in R. I feel anxious about
# the possibilities of typos when these values are calculated somewhere I can't 
# see and pasted into the google sheet. I prefer for the workflow of calculations
# to be as transparent and traceable as possible. From this point in the qmd I
# think we can use the volume values from this google sheet. Another option is to 
# export the google sheet as a CSV, and put that in the data folder in this project. 
# Then we can use read_csv to import it.

#SS: is that true?: (I have a weird number in row 5, cl Depth and couldn't find out what happened)

#### previously I had a column'excavation area'; I seprated them in 2 columns in order to calculate the colume; but now when I wanted to run the code it said "! object 'excavation area' not found"; would it be OK to delete these lines?

# Separate `excavation area` into `length` and `width` and calculate `volume`
sum_depth_sheet <- sum_depth_sheet %>%
  separate(`excavation area`, into = c("length", "width"), sep = "\\*") %>%
  mutate(
    length = as.numeric(length),
    width = as.numeric(width),
    `thickness` = `thickness` / 100  # Convert from cm to m
  ) %>%
  mutate(
    volume = length * width * `thickness`  # Calculate volume in m^3
  )

# Calculate lithic volumetric density 
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    `lithic_dens` = round(total_lithics / volume, 2)
  )

# Calculate retouch frequency 
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    retouch_freq = round(retouch_tool / volume, 2)  
  )

```

#try with summary_table_depth_area_unit
```{r}
# Create the 'label' column
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(label = paste(area, chronological_unit, grouped_level, sep = "_"))

# Add a suffix only to duplicates
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  group_by(label) %>%
  mutate(suffix = case_when(
    n() == 1 ~ NA_character_,
    TRUE ~ letters[row_number()]
  )) %>%
  ungroup()

# Append the suffix to 'label'
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(label = if_else(is.na(suffix), label, paste0(label, "_", suffix))) %>%
  select(-suffix)

# Relocate 'label' column after 'level'
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  relocate(label, .after = grouped_level)

------------------------------------------
# working up to here. I need to add length, width, and thickness columns
------------------------------------------

# Separate `excavation area` into `length` and `width` and calculate `volume`
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  separate(`excavation area`, into = c("length", "width"), sep = "\\*") %>%
  mutate(
    length = as.numeric(length),
    width = as.numeric(width),
    `thickness` = `thickness` / 100  # Convert from cm to m
  ) %>%
  mutate(
    volume = length * width * `thickness`  # Calculate volume in m^3
  )

# Calculate lithic volumetric density 
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(
    `lithic_dens` = round(total_lithics / volume, 2)
  )

# Calculate retouch frequency 
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(
    retouch_freq = round(retouch_tool / volume, 2)  
  )
```

```{r}
library(ggpmisc)

# Create the plot: WABI_Level_logscale
# here I do not consider those levels with a value 0 of retouch frequency. 
filtered_data <- sum_depth_sheet %>% 
  filter(lithic_dens > 0 & retouch_freq > 0)

WABI_Level <- ggplot(filtered_data,  
            aes(x = lithic_dens,
                y = retouch_freq)) +
  geom_point(color = "black", size = 1.5) +
  geom_smooth(method = "lm", 
              se = TRUE, 
              color = "blue", 
              linewidth = 0.5) +
  # Add text labels based on 'label' column
  geom_text(aes(label = label), 
            vjust = 1.5, 
            hjust = 0.5, 
            check_overlap = TRUE) +

  stat_poly_eq(aes(label = paste(stat(eq.label), 
                                  stat(rr.label), 
                                  stat(p.value.label), 
                                  sep = "~~~")),
               formula = y ~ x, 
               parse = TRUE, 
               size = 4) +
  labs(
    title = "Relationship between Retouch Frequency and Lithic Volumetric Density",
    x = "Lithic Volumetric Density",
    y = "Retouch Frequency"
  ) +
  scale_x_log10() +
  scale_y_log10() +
  theme(
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12)
  )

# Print the plot to display it
print(WABI_Level)

```

```{r}
#PCA_Level: preparing data

#PCA_level: for PCA Bicho and Cascalheira considered these variables: estimated area (as we calculated based on levels, should I consider the thickness?), core frequency, blank frequency, chip frequency: artefacts smaller than 1 cm (as we work on a neolithic asssemblage, we have so many small artefacts; so, I do not consider it), feature frequency (we have some stone alignments, but we do not know the exact level, we know the area of those features), retouch frequency, tool diversity (diversity of tool types within each assemblage, calculated using Menhinick’s index); 
  ##core frequency
  ##blank frequency
  ##retouch frequency
  ##tool diversity
  ##thickness of deposit (???)
  ##lithic density

#Calculate Core Frequency
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    core_freq = round(cores / volume, 2)  
  )

# calculate Blank Frequency
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    blank_freq = round((flake + blade + bladelet) / volume, 2)  
  )

# Calculate tool diversity using Menhinick's index
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(
    total_tools = retouch_tool + utilized + sickle_shine,
    tool_diversity = ifelse(total_tools > 0, n_distinct(tool_type) / sqrt(total_tools), NA)
  )

```

```{r}
# performin PCA

library(FactoMineR)
library(factoextra)

sum_depth_sheet_subset <- 
 sum_depth_sheet %>%
  drop_na(tool_diversity) #%>% 
  #filter(!label %in% c("tp1_c_3_a",
 #                      "d1_a_2_b",
  #                     "d1_c_3_b"))

pca_data <- 
sum_depth_sheet_subset %>% 
  select(lithic_dens, 
         core_freq, 
         blank_freq, 
         retouch_freq, 
         tool_diversity)

# Perform PCA
res.pca <- PCA(pca_data, graph = FALSE)

# View the summary results
summary(res.pca)

```

```{r}
#Plot: Contribution of variables for each of the four relevant PCA dimensions

#save contribution plots into one objects

contribution_plots <- 
map(1:5, ~fviz_contrib(res.pca, 
                   choice = "var", 
                   axes = .x, 
                   top = 10))

#combine individual plots into one plot 
library(cowplot)
plot_grid(plotlist = contribution_plots,
                   ncol = 2)

```

```{r}
# Extract contributions of variables to each principal component
contributions <- res.pca$var$contrib

# Compute average contribution for each dimension
avg_contrib_dim1 <- mean(contributions[,1], na.rm = TRUE)
avg_contrib_dim2 <- mean(contributions[,2], na.rm = TRUE)
avg_contrib_dim3 <- mean(contributions[,3], na.rm = TRUE)
avg_contrib_dim4 <- mean(contributions[,4], na.rm = TRUE)
avg_contrib_dim5 <- mean(contributions[,5], na.rm = TRUE)

# Find variables that contribute more than average to each dimension
high_contrib_dim1 <- contributions[contributions[,1] > avg_contrib_dim1, 1]
high_contrib_dim2 <- contributions[contributions[,2] > avg_contrib_dim2, 2]
high_contrib_dim3 <- contributions[contributions[,3] > avg_contrib_dim3, 3]
high_contrib_dim4 <- contributions[contributions[,4] > avg_contrib_dim4, 4]
high_contrib_dim5 <- contributions[contributions[,5] > avg_contrib_dim4, 5]

# Sort them in descending order
sorted_high_contrib_dim1 <- sort(high_contrib_dim1, decreasing = TRUE)
sorted_high_contrib_dim2 <- sort(high_contrib_dim2, decreasing = TRUE)
sorted_high_contrib_dim3 <- sort(high_contrib_dim3, decreasing = TRUE)
sorted_high_contrib_dim4 <- sort(high_contrib_dim4, decreasing = TRUE)
sorted_high_contrib_dim5 <- sort(high_contrib_dim5, decreasing = TRUE)

# Print or use for further analysis
print("Significant contributions to Dim1:")
print(sorted_high_contrib_dim1)

print("Significant contributions to Dim2:")
print(sorted_high_contrib_dim2)

print("Significant contributions to Dim3:")
print(sorted_high_contrib_dim3)

print("Significant contributions to Dim4:")
print(sorted_high_contrib_dim4)

print("Significant contributions to Dim5:")
print(sorted_high_contrib_dim5)
```

```{r}
# screeplot; inspect distribution of PCs
fviz_screeplot(res.pca)
```

```{r}
# Inspect eigenvalues
res.pca$eig
```

```{r}
eigen_df <- data.frame(
  eigenvalue = res.pca$eig[,1],
  `percentage of variance` = res.pca$eig[,2],
  `cumulative percentage of variance` = res.pca$eig[,3]
)

```

```{r}

library(kableExtra)
library(htmltools)

# Create a new column named "Dimension"
eigen_df$Dimension <- 
  paste("Dim", seq_len(nrow(eigen_df)), sep = "")

# Re-order the columns based on the corrected names
eigen_df <- eigen_df[, c("Dimension", "eigenvalue", "percentage.of.variance", "cumulative.percentage.of.variance")]

# Round the numerical columns to 3 decimal places
eigen_df[, 2:4] <- round(eigen_df[, 2:4], 3)

# Removing row names
rownames(eigen_df) <- NULL

# Creating the kable output
kable_output <- kable(eigen_df, 
                      "html", 
                      align = 'c', 
                      col.names = c("Dimension", 
                                    "Eigenvalue", 
                                    "Variance Percent", 
                                    "Cumulative Variance Percent")) %>%
  kable_styling("striped", 
                full_width = F) %>%
  add_header_above(c(" " = 1, "Eigenvalues and percentage of variance for each dimension of PCA" = 3))

# Display the table
kable_output

```

```{r}
library(ggrepel)

rownames(res.pca$ind$coord) <- sum_depth_sheet_subset$label

# Generate the base biplot
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               labelsize = 2,
                               label = "var",
                               addEllipses = TRUE,
                               col.ind = sum_depth_sheet_subset$area
                              )

# Modify row names of coordinates
rownames(res.pca$ind$coord) <- sum_depth_sheet_subset$label

pca_label_text <- 
res.pca$ind$coord %>% 
  as_tibble() %>% 
  mutate(labels = sum_depth_sheet_subset$label,
         area =   sum_depth_sheet_subset$area)

# Add geom_text_repel() to the base plot
biplot_modified <- biplot_base +
                   geom_text_repel(data = pca_label_text,
                                   aes(x = Dim.1, 
                                       y = Dim.2, 
                                       label = labels,
                                       colour = area),
                                   segment.color = "grey80",
                                   size = 2,
                                   force = 10
                                  )

# Display the plot
biplot_modified

```

```{r}
# Base biplot without points
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               geom = "none",  
                               labelsize = 2
                               )

# Add points and text labels manually
biplot_modified <- biplot_base +
  geom_point(data = coord_df, 
             aes(x = Dim.1, 
                 y = Dim.2, 
                 color = group), size = 1) +
  geom_text_repel(data = coord_df,
                  aes(x = Dim.1,
                      y = Dim.2,
                      label = label),
                  box.padding = 0.5,
                  point.padding = 0.5,
                  size = 2)

# Display the plot
biplot_modified

```

```{r}
# Visualize biplot: 2
fviz_pca_biplot(res.pca,
                axes = c(3, 4),
                labelsize = 2
                )
rownames(res.pca$ind$coord) <- sum_depth_sheet$label
```